---
title: '04 Invulnerable Types - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: '../../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "..",
    "results", unlist(strsplit(getwd(), "/"))[6],
    unlist(strsplit(getwd(), "/"))[7], "04-invulnerable-types.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

In this document we probe for vulnerable cell types.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r}
#    This file is part of tau-vulnerability.
#    Copyright (C) 2019-2020  Emir Turkes, UK DRI at UCL, Columbia University Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

options(stringsAsFactors = FALSE)
packages <- c(
  "conflicted", "Seurat", "future", "parallel", "biomaRt", "dplyr", "ggplot2", "ggrepel", "viridis"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "..", "utils.R"))

protocol <- c("human", "droplet", "single-nuc", "umis") # See `cluster_pipeline` in `utils.R`.
subset_names <- "EC_neuronal" # Subsetted Seurats to loop over.

analysis_no <- 4
data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
assets_dir <- file.path(getwd(), "..", "..", "..", "assets") # Backed up data.
results_dir <- file.path(getwd(), "..", "..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path(
  "..", "..", "..", "cache", data_name, unlist(strsplit(getwd(), "/"))[7], paste0("0", analysis_no)
)
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7)
```

We use data saved from previous analyses.

```{r}
subset_list <- c(readRDS(file.path(cache_dir, "..", "02", paste0(subset_names[1], "_seurat.rds"))))

for (i in 1:length(subset_list)) {
  print(subset_names[i])
  print(red_dim_plot(subset_list[[i]], "umap1", "umap2", "seurat_clusters", "cat"))

  if (protocol[1] == "human") {
    plots <- FeaturePlot(
      subset_list[[i]], c("SLC17A7", "GAD1"), order = TRUE, blend = TRUE,
      combine = FALSE, blend.threshold = 0, max.cutoff = "q10"
    )
  } else if (protocol[1] == "mouse") {
    plots <- FeaturePlot(
      subset_list[[i]], c("Slc17a7", "Gad1"), order = TRUE, blend = TRUE,
      combine = FALSE, blend.threshold = 0, max.cutoff = "q10"
    )
  }

  print(CombinePlots(plots[3:4], legend = "none"))
}
```

# IN

```{r}
if (protocol[1] == "human") {
  genes <- c("PVALB", "SST", "CALB2", "GAD1")
} else if (protocol[1] == "mouse") {
  genes <- c("Pvalb", "Sst", "Calb2", "Gad1")
}

gene_plot(subset_list, subset_names, genes)
```

# L4-5 EX

```{r}
if (protocol[1] == "human") {
  genes <- c("RORB", "GABGR1")
} else if (protocol[1] == "mouse") {
  genes <- c("Rorb", "Gabgr1")
}

gene_plot(subset_list, subset_names, genes)
```

# L5-6 EX

```{r}
if (protocol[1] == "human") {
  genes <- c("NR4A2", "NTNG2")
} else if (protocol[1] == "mouse") {
  genes <- c("Nr4a2", "Ntng2")
}

gene_plot(subset_list, subset_names, genes)
```

# HIP DG EX

```{r}
if (protocol[1] == "human") {
  genes <- c("SSTR2", "PROX1")
} else if (protocol[1] == "mouse") {
  genes <- c("Sstr4", "Prox1")
}

gene_plot(subset_list, subset_names, genes)
```

# HIP CA3 EX

```{r}
if (protocol[1] == "human") {
  genes <- c("GPRIN3", "BOK")
} else if (protocol[1] == "mouse") {
  genes <- c("Gprin3", "Bok")
}

gene_plot(subset_list, subset_names, genes)
```

# Hip CA2 EX

```{r}
if (protocol[1] == "human") {
  genes <- c("CACNG5", "PTPN5")
} else if (protocol[1] == "mouse") {
  genes <- c("Cacng5", "Ptpn5")
}

gene_plot(subset_list, subset_names, genes)
```

# S1 L2-3 EX

```{r}
if (protocol[1] == "human") {
  genes <- c("CARTPT", "LAMP5")
} else if (protocol[1] == "mouse") {
  genes <- c("Cartpt", "Lamp5")
}

gene_plot(subset_list, subset_names, genes)
```

# V1 L4 EX

```{r}
if (protocol[1] == "human") {
  genes <- c("CALCA", "WHRN")
} else if (protocol[1] == "mouse") {
  genes <- c("Calca", "Whrn")
}

gene_plot(subset_list, subset_names, genes)
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
