---
title: '01 Prep - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "01-prep.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

In this document we prepare the gene count matrix for downstream analysis.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r}
#    This file is part of tau-vulnerability.
#    Copyright (C) 2019-2020  Emir Turkes, UK DRI at UCL, Columbia University Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# This section should be checked per document.
# Blank variables should be set to NULL.
# --------------------------------------------
options(stringsAsFactors = FALSE)
packages <- c(
  "conflicted", "BiocFileCache", "scrattch.io", "SingleCellExperiment", "scater", "DT",
  "data.table", "DropletUtils", "Seurat", "future", "Biostrings", "zinbwave",
  "biomaRt", "dplyr", "ggplot2", "ggrepel", "viridis"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
conflict_prefer("strsplit", "Biostrings", quiet = TRUE)
source(file.path(getwd(), "..", "utils.R"))

protocol <- c("human", "droplet", "single-nuc", "umis") # See `cluster_pipeline` in `utils.R`.
controls <- NULL # Extract controls. First term is the SCE column, second is label.
vars_to_regress <- NULL # Used for sctransform in `cluster_pipeline` in `utils.R`
parallel_override <- NULL # See `parallel_plan` in `utils.R`.
download_name <- "GTEx_droncseq_hip_pcf" # Name of download file corresponding to dataset.

# Metadata to plot after dimensionality reduction and clustering.
# Values in list can include "no_legend and/or "no_label" to exclude those.
# -------------------------------------------------------------------------
metadata_to_plot <- vector("list", 4)
names(metadata_to_plot) <- c("seurat_clusters", "Cluster.Name", "sample_stem", "Phase")
metadata_to_plot$sample_stem <- "no_label"
metadata_to_plot$Phase <- "no_label"
# -------------------------------------------------------------------------
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
analysis_no <- 1
data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
assets_dir <- file.path(getwd(), "..", "..", "assets") # Backed up data.

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path(getwd(), "..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 10, fig.height = 7)
# ----------------------------------------------------------
```

# Cleaning

We start by gathering together the data and forming a SingleCellExperiment (SCE) object.

```{r}
# This is generally the only chunk that requires code specific to each dataset.
# -----------------------------------------------------------------------------

# Download data using `BiocFileCache`.
# ------------------------------------
data <- bfcrpath(
  BiocFileCache(cache_dir, FALSE),
  file.path(
    "https://storage.googleapis.com/gtex_additional_datasets/single_cell_data",
    paste0(download_name, ".tar")
  )
)
untar(data, exdir = tempdir())
# ------------------------------------

counts <- fread(
  file.path(tempdir(), download_name, paste0(download_name, ".umi_counts.txt.gz")),
  data.table = FALSE
)
rownames(counts) <- counts[ , 1]
counts <- as.matrix(counts[ , -1])
counts <- Matrix::Matrix(counts, sparse = TRUE)

sample_stem <- sapply(colnames(counts), function(x) strsplit(x, "_")[[1]][1], USE.NAMES = FALSE)
col_data <- data.frame(sample = colnames(counts), sample_stem = sample_stem, check.names = FALSE)

tsne <- fread(
  file.path(tempdir(), download_name, paste0(download_name, ".tsne.txt.gz")),
  data.table = FALSE
)

# The original clusters file, `GTEx_droncseq_hip_pcf.clusters.txt`, has several inconsistencies.
# Specifically in clusters 15-18, which do not match those in the original publication.
# Therefore we use Supplementary Table 7 (nmeth.4407-S10.xlsx), which does not have those issues.
# However, because cluster 11 is incorrectly named in this table, we must correct that.
# Also, as this file contained non-standard formatting, some manual editing was done for use in R.
# ------------------------------------------------------------------------------------------------
clusters <- fread(
  file.path(assets_dir, data_name, "nmeth.4407-S10-edited.txt"),
  data.table = FALSE
)
# ------------------------------------------------------------------------------------------------

# Fix mislabeled clusters and give unique names for unclassified clusters.
# ------------------------------------------------------------------------
for (i in seq_len(nrow(clusters))) {
  if (clusters[i, 4] == 11) {
    clusters[i, 5] <- "ODC2"
  } else if (clusters[i, 4] == 16) {
    clusters[i, 5] <- "Unlabeled1"
  } else if (clusters[i, 4] == 17) {
    clusters[i, 5] <- "Unlabeled2"
  } else if (clusters[i, 4] == 18) {
    clusters[i, 5] <- "Unlabeled3"
  }
}
# ------------------------------------------------------------------------

clusters <- clusters[match(col_data$sample, clusters[["Cell ID"]]), ]
rownames(clusters) <- NULL
col_data <- cbind(
  col_data, tsne[ , names(tsne) != "V1"], clusters[ , names(clusters) != "Cell ID"]
)

sce <- SingleCellExperiment(list(counts = counts), colData = col_data)
rm(counts)
sce

if (is.null(controls)) {
  print("All control data.")
} else {
  print("Subset to controls.")
  sce <- sce[ , grepl(controls[2], sce[[controls[1]]])]
  sce
}

batch <- NULL
# -----------------------------------------------------------------------------
```

# QC

First, we add cell QC metric metadata to the SCE object.

```{r}
mito <- grep("^MT-", rownames(sce), TRUE)

if (length(mito) != 0) {
  sce <- addPerCellQC(sce, list(mito = mito), BPPARAM = MulticoreParam())
  if (is.null(batch)) {
    remove <- quickPerCellQC(sce, percent_subsets = "subsets_mito_percent")
  } else {
    remove <- quickPerCellQC(sce, percent_subsets = "subsets_mito_percent", batch = batch)
  }

  if (protocol[3] == "single-nuc") {
    print(paste0("Removed ", length(mito), " mitochondrial genes."))
    sce <- sce[-mito, ]
    dim(sce)
  }

} else {
  sce <- addPerCellQC(sce, BPPARAM = MulticoreParam())
  if (is.null(batch)) {
    remove <- quickPerCellQC(sce)
  } else {
    remove <- quickPerCellQC(sce, batch = batch)
  }
}

names(colData(sce))[(length(names(colData(sce))) - 6):length(names(colData(sce)))]
```

We use adaptive thresholds to remove cells that are outliers by more than 3 MADs.

```{r}
if (is.null(batch)) {
  remove <- quickPerCellQC(sce)
} else {
  remove <- quickPerCellQC(sce, batch = batch)
}
sce$discard <- remove$discard
datatable_download(t(colSums(as.matrix(remove))))

sce <- sce[ , !sce$discard]
dim(sce)
```

Here we carry out any protocol specific techniques, such as identification of empty droplets in a droplet-based experiment.

```{r}
if (protocol[2] == "droplet") {
  # Identify empty droplets as those with a low UMI count.
  # ------------------------------------------------------
  bcrank <- barcodeRanks(counts(sce))
  uniq <- !duplicated(bcrank$rank) # Only show unique points for plotting speed.
  plot(
    bcrank$rank[uniq], bcrank$total[uniq], log = "xy",
    xlab = "Rank", ylab = "Total UMI count", cex.lab = 1.2
  )
  abline(h = metadata(bcrank)$inflection, col = "darkgreen", lty = 2)
  abline(h = metadata(bcrank)$knee, col = "dodgerblue", lty = 2)
  legend(
    "bottomleft", legend = c("Inflection", "Knee"), col = c("darkgreen", "dodgerblue"),
    lty = 2, cex = 1.2
  )
  # ------------------------------------------------------

  # TODO: Don't stop on "no counts available to estimate the ambient profile" error.
  # --------------------------------------------------------------------------------
  # set.seed(1)
  # e_out <- emptyDrops(counts(sce))
  # --------------------------------------------------------------------------------
}
```

Finally we add feature QC metrics and remove features not expressed in any cell.

```{r, fig.height = 5}
sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())
names(rowData(sce))
dim(sce)

par(mfrow = c(1, 3), mar = c(5, 4, 1, 1))
hist(
 log10(rowData(sce)$mean + 1e-6), col = "grey80",  main = "",
 breaks = 40, xlab = "log10(Mean Counts Per Gene + 1e-6)")
hist(
 log10((rowData(sce)$detected * dim(sce)[2]) + 1), col = "grey80", main = "",
 breaks = 40, xlab = "log10(Number of Cells Expressing Gene + 1)")
plot(
 log10(rowData(sce)$mean + 1e-6), pch = 16,
 col = rgb(0, 0, 0, 0.4), log10((rowData(sce)$detected * dim(sce)[2]) + 1),
 xlab = "log10(Mean Counts Per Gene + 1e-6)", ylab = "log10(Number of Cells Expressing Gene + 1)")

sce <- sce[rowSums(counts(sce) > 0) > 0, ]
rowData(sce) <- NULL
sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())
dim(sce)

par(mfrow = c(1, 3), mar = c(5, 4, 1, 1))
hist(
 log10(rowData(sce)$mean + 1e-6), col = "grey80",  main = "",
 breaks = 40, xlab = "log10(Mean Counts Per Gene + 1e-6)")
hist(
 log10((rowData(sce)$detected * dim(sce)[2]) + 1), col = "grey80", main = "",
 breaks = 40, xlab = "log10(Number of Cells Expressing Gene + 1)")
plot(
 log10(rowData(sce)$mean + 1e-6), pch = 16,
 col = rgb(0, 0, 0, 0.4), log10((rowData(sce)$detected * dim(sce)[2]) + 1),
 xlab = "log10(Mean Counts Per Gene + 1e-6)", ylab = "log10(Number of Cells Expressing Gene + 1)")
```

# Cluster Pipeline

We run a pipeline that applies normalization and scaling, dimensionality reduction, and clustering.

```{r}
# Remove redundant QC metrics.
# ----------------------------
if (length(mito) != 0) {
  colData(sce)[length(names(colData(sce)))] <- NULL
  colData(sce)[(length(names(colData(sce))) - 8):(length(names(colData(sce))) - 3)] <- NULL
} else {
  colData(sce)[(length(names(colData(sce))) - 6):length(names(colData(sce)))] <- NULL
}
rowData(sce) <- NULL
# ----------------------------

sub_name <- "all"
seurat <- as.Seurat(sce, data = NULL)
rm(sce)
seurat <- cluster_pipeline(
  seurat, cache_dir, sub_name, protocol, vars_to_regress, parallel_override
)
seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend())
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}

if (protocol[4] == "umis") {
  print("nFeature_SCT")
  red_dim_plot(seurat, "umap1", "umap2", "nFeature_SCT", "cont")
  print("nCount_SCT")
  red_dim_plot(seurat, "umap1", "umap2", "nCount_SCT", "cont")
} else if (protocol[4] == "reads") {
  print("nFeature_RNA")
  red_dim_plot(seurat, "umap1", "umap2", "nFeature_RNA", "cont")
  print("nCount_RNA")
  red_dim_plot(seurat, "umap1", "umap2", "nCount_RNA", "cont")
}

if (length(mito) != 0) {
  print("subsets_mito_percent")
  red_dim_plot(seurat, "umap1", "umap2", "subsets_mito_percent", "cont")
}
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
