---
title: '02 Subsets - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "02-subsets.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).

In this particular document we create useful subsets of the original data.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

#    This file is part of tau-vulnerability.
#    Copyright (C) 2019  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# ------------------------------------------

packages <- c(
  "conflicted", "Seurat", "future", "scater", "dplyr", "ggplot2", "ggrepel", "viridis", "DT",
  "data.table"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 2
protocol_type <- "droplet"
download_name <- "GTEx_droncseq_hip_pcf" # Name of download file corresponding to dataset.

knitr::opts_chunk$set(fig.width = 15, fig.height = 13)
options(stringsAsFactors = FALSE)
plan("multiprocess") # Parallel processing for Seurat.

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
data_name_stem <- unlist(strsplit(data_name, "-"))[1] # `data_name` up to the first hyphen.
data_dir <- file.path(getwd(), "..", "..", "data", data_name) # Publicly available data.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Misc binaries and temp files.
results_dir <- file.path(getwd(), "..", "..", "results", data_name)

# Unique cache and results directory for each analysis number.
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}
if (!dir.exists(file.path(results_dir, "data", paste0("0", analysis_no)))) {
  dir.create(file.path(results_dir, "data", paste0("0", analysis_no)), recursive = TRUE)
}
```

We use data saved from previous analyses.

```{r}
seurat_orig <- readRDS(file.path(results_dir, "data", paste0("0", analysis_no - 1), "seurat.rds"))
seurat_orig
```

# BA9

```{r}
sub_name <- "BA9"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  sub <- "PFC"
  keep <- grepl(sub, seurat_orig$batch)
  seurat <- seurat_orig[ , keep]

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
BA9_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

BA9_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 5, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:4], ncol = 2)
VlnPlot(seurat, features = top$gene[5:8], ncol = 2)
VlnPlot(seurat, features = top$gene[9:12], ncol = 2)

FeaturePlot(
  seurat, features = top$gene[1:4], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[5:8], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[9:12], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 45, wt = avg_logFC))
```

## Neuronal

```{r}
sub_name <- "BA9_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat, idents = c(5, 6, 0, 2, 4))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"),
    verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default optimized for 3K cells.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # # Default optimized for 3K cells.
print(paste0(sub_name, " Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# View canonical EX/IN marker genes.
genes <- c("SLC17A7", "GAD1")
if (all(genes %in% rownames(seurat))) { # `blend` requires all features present.
  plots <- FeaturePlot(
    seurat, features = genes, order = TRUE,
    reduction = "umap", blend = TRUE, combine = FALSE
  )
  CombinePlots(plots[3:4], legend = "none")
} else {
  FeaturePlot(
    seurat, features = genes, order = TRUE, reduction = "umap",
    cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
  )
}
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 1)

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # Recommended to use the `RNA` assay for DEGs.
BA9_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

BA9_neuronal_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:4], ncol = 2)
VlnPlot(seurat, features = top$gene[5:8], ncol = 2)

FeaturePlot(
  seurat, features = top$gene[1:4], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[5:8], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 50, wt = avg_logFC))
```

### EX

```{r}
sub_name <- "BA9_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(BA9_neuronal_seurat, idents = c(0, 1, 2, 6, 7))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"),
    verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default optimized for 3K cells.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # # Default optimized for 3K cells.
print(paste0(sub_name, " Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # Recommended to use the `RNA` assay for DEGs.
BA9_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

BA9_neuronal_EX_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:4], ncol = 2)

FeaturePlot(
  seurat, features = top$gene[1:4], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 100, wt = avg_logFC))
```

### IN

```{r}
sub_name <- "BA9_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(BA9_neuronal_seurat, idents = c(3, 4, 5))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"),
    verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- 0.8 # We use the default because optimized value produced no clusters.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # # Default optimized for 3K cells.
print(paste0(sub_name, " Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # Recommended to use the `RNA` assay for DEGs.
BA9_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

BA9_neuronal_IN_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:4], ncol = 2)

FeaturePlot(
  seurat, features = top$gene[1:4], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 100, wt = avg_logFC))
```

# Hippocampus

```{r}
sub_name <- "hip"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  sub <- "hC|hH|H"
  keep <- grepl(sub, seurat_orig$batch)
  seurat <- seurat_orig[ , keep]

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
hip_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

hip_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 3, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 2) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:9], ncol = 3)
VlnPlot(seurat, features = top$gene[10:17], ncol = 3)

FeaturePlot(
  seurat, features = top$gene[1:9], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 3
)
FeaturePlot(
  seurat, features = top$gene[10:17], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 3
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 30, wt = avg_logFC))
```

## Neuronal

```{r}
sub_name <- "hip_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat, idents = c(13, 7, 12, 1, 0, 3, 6))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"),
    verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default optimized for 3K cells.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default optimized for 3K cells.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# View canonical EX/IN marker genes.
genes <- c("SLC17A7", "GAD1")
if (all(genes %in% rownames(seurat))) { # `blend` requires all features present.
  plots <- FeaturePlot(
    seurat, features = genes, order = TRUE,
    reduction = "umap", blend = TRUE, combine = FALSE
  )
  CombinePlots(plots[3:4], legend = "none")
} else {
  FeaturePlot(
    seurat, features = genes, order = TRUE, reduction = "umap",
    cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
  )
}
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 1)

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # Recommended to use the `RNA` assay for DEGs.
hip_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

hip_neuronal_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 5, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:5], ncol = 3)
VlnPlot(seurat, features = top$gene[6:10], ncol = 3)

FeaturePlot(
  seurat, features = top$gene[1:5], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 3
)
FeaturePlot(
  seurat, features = top$gene[6:10], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 3
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 50, wt = avg_logFC))
```

### EX

```{r}
sub_name <- "hip_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(hip_neuronal_seurat, idents = c(0, 5, 3, 9, 8, 7, 2, 1))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"),
    verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default optimized for 3K cells.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # # Default optimized for 3K cells.
print(paste0(sub_name, " Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # Recommended to use the `RNA` assay for DEGs.
hip_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

hip_neuronal_EX_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:4], ncol = 2)
VlnPlot(seurat, features = top$gene[5:8], ncol = 2)

FeaturePlot(
  seurat, features = top$gene[1:4], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[5:8], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 50, wt = avg_logFC))
```

### IN

```{r}
sub_name <- "hip_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(hip_neuronal_seurat, idents = c(4, 10, 6))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"),
    verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default optimized for 3K cells.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # # Default optimized for 3K cells.
print(paste0(sub_name, " Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # Recommended to use the `RNA` assay for DEGs.
hip_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

hip_neuronal_IN_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 40, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:2], ncol = 1)

if (all(genes %in% rownames(seurat))) { # `blend` requires all features present.
  plots <- FeaturePlot(
    seurat, features = top$gene, order = TRUE,
    reduction = "umap", blend = TRUE, combine = FALSE
  )
  CombinePlots(plots[3:4], legend = "none")
} else {
  FeaturePlot(
    seurat, features = genes, order = TRUE, reduction = "umap",
    cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
  )
}

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 250, wt = avg_logFC))
```

# Combined Neuronal

```{r}
sub_name <- "comb_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- merge(BA9_neuronal_seurat, hip_neuronal_seurat)

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"),
    verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(
    seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE
  )
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default optimized for 3K cells.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(
    seurat, resolution = resolution, verbose = FALSE
  )

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Fix formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default optimized for 3K cells.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# View canonical EX/IN marker genes.
genes <- c("SLC17A7", "GAD1")
if (all(genes %in% rownames(seurat))) { # `blend` requires all features present.
  plots <- FeaturePlot(
    seurat, features = genes, order = TRUE,
    reduction = "umap", blend = TRUE, combine = FALSE
  )
  CombinePlots(plots[3:4], legend = "none")
} else {
  FeaturePlot(
    seurat, features = genes, order = TRUE, reduction = "umap",
    cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
  )
}
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 1)

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # Recommended to use the `RNA` assay for DEGs.
comb_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

comb_neuronal_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 5, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:4], ncol = 2)
VlnPlot(seurat, features = top$gene[5:8], ncol = 2)
VlnPlot(seurat, features = top$gene[9:12], ncol = 2)
VlnPlot(seurat, features = top$gene[13:16], ncol = 2)

FeaturePlot(
  seurat, features = top$gene[1:4], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[5:8], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[9:12], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[13:16], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 30, wt = avg_logFC))
```

## EX

```{r}
sub_name <- "comb_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(
    comb_neuronal_seurat, idents = c(3, 1, 10, 19, 17, 2, 18, 6, 5, 0, 11, 16, 14, 12, 4)
  )

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"),
    verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default optimized for 3K cells.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # # Default optimized for 3K cells.
print(paste0(sub_name, " Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # Recommended to use the `RNA` assay for DEGs.
comb_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

comb_neuronal_EX_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 3, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 1.5) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:4], ncol = 2)
VlnPlot(seurat, features = top$gene[5:8], ncol = 2)
VlnPlot(seurat, features = top$gene[6:12], ncol = 2)
VlnPlot(seurat, features = top$gene[13:16], ncol = 2)
VlnPlot(seurat, features = top$gene[17:20], ncol = 2)

FeaturePlot(
  seurat, features = top$gene[1:4], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[5:8], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[9:12], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[13:16], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
FeaturePlot(
  seurat, features = top$gene[17:20], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 25, wt = avg_logFC))
```

## IN

```{r}
sub_name <- "comb_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(comb_neuronal_seurat, idents = c(9, 7, 20, 8, 13, 15))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- ScaleData(seurat, verbose = FALSE)

  # Saves into new assay `SCT` while previous data is retained in `RNA`.
  # Note that this function produces many iterations of the following benign warning:
  # Warning in theta.ml(y = y, mu = fit$fitted): iteration limit reached
  seurat <- SCTransform(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"),
    verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2

  seurat <- RunUMAP(seurat, dims = 1:30, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default optimized for 3K cells.
  seurat <- FindNeighbors(seurat, dims = 1:30, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:15, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # # Default optimized for 3K cells.
print(paste0(sub_name, " Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # Recommended to use the `RNA` assay for DEGs.
comb_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

comb_neuronal_IN_markers <- markers

top <- markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:4], ncol = 2)

FeaturePlot(
  seurat, features = top$gene[1:4], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 100, wt = avg_logFC))
```

# References

```{r}
saveRDS(BA9_seurat, file.path(results_dir, "data", paste0("0", analysis_no), "BA9_seurat.rds"))
saveRDS(BA9_markers, file.path(results_dir, "data", paste0("0", analysis_no), "BA9_markers.rds"))
saveRDS(
  BA9_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "BA9_neuronal_seurat.rds")
)
saveRDS(
  BA9_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "BA9_neuronal_markers.rds")
)
saveRDS(
  BA9_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "BA9_neuronal_EX_seurat.rds")
)
saveRDS(
  BA9_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "BA9_neuronal_EX_markers.rds")
)
saveRDS(
  BA9_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "BA9_neuronal_IN_seurat.rds")
)
saveRDS(
  BA9_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "BA9_neuronal_IN_markers.rds")
)

saveRDS(hip_seurat, file.path(results_dir, "data", paste0("0", analysis_no), "hip_seurat.rds"))
saveRDS(hip_markers, file.path(results_dir, "data", paste0("0", analysis_no), "hip_markers.rds"))
saveRDS(
  hip_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "hip_neuronal_seurat.rds")
)
saveRDS(
  hip_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "hip_neuronal_markers.rds")
)
saveRDS(
  hip_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "hip_neuronal_EX_seurat.rds")
)
saveRDS(
  hip_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "hip_neuronal_EX_markers.rds")
)
saveRDS(
  hip_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "hip_neuronal_IN_seurat.rds")
)
saveRDS(
  hip_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "hip_neuronal_IN_markers.rds")
)

saveRDS(
  comb_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "comb_neuronal_seurat.rds")
)
saveRDS(
  comb_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "comb_neuronal_markers.rds")
)
saveRDS(
  comb_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "comb_neuronal_EX_seurat.rds")
)
saveRDS(
  comb_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "comb_neuronal_EX_markers.rds")
)
saveRDS(
  comb_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "comb_neuronal_IN_seurat.rds")
)
saveRDS(
  comb_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "comb_neuronal_IN_markers.rds")
)

sessionInfo()
```
