---
title: '02 Subsets - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "02-subsets.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

In this document we create several subsets of the data and rerun the clustering pipeline in `01_prep.Rmd`.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r, boilerplate}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

#    This file is part of tau-vulnerability.
#    Copyright (C) 2019-2020  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# -------------------------------------

packages <- c(
  "conflicted", "Seurat", "future", "parallel", "biomaRt", "dplyr", "ggplot2", "ggrepel", "viridis"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 2
organism <- "mouse"
batch_metadata <- "donor_id"
metadata_to_plot <- c(
  "seurat_clusters", "class_label", "region_label", "cluster_label", batch_metadata, "Phase"
)
no_legend <- 4
no_label <- c(3, 5, 6)
vars_to_regress <- batch_metadata
parallel_override <- NULL # See function `parallel_plan` in `utils.R`.
subsets <- c("ENT", "ENT_neuronal")
subset_counter <- 1
options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 10)

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Backed up caches and data.
tmp_dir <- file.path(getwd(), "..", "..", "tmp", data_name)
results_dir <- file.path(getwd(), "..", "..", "results", data_name)

if (!dir.exists(tmp_dir)) {
  dir.create(tmp_dir, recursive = TRUE)
}

# Unique cache directory for each analysis number.
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}
```

We start by reading in the original data.

```{r, prep}
seurat_orig <- readRDS(
  file.path(assets_dir, "cache", paste0("0", analysis_no - 1), "all_seurat.rds")
)
seurat_orig

if ("subsets_mito_percent" %in% names(seurat_orig[[]])) {
  mito <- 1
} else {
  mito <- NULL
}

if (is.null(subsets)) {
  print("No further subsets necessary.")
  saveRDS(seurat_orig, file.path(assets_dir, "cache", paste0("0", analysis_no), "all_seurat.rds"))

  for (i in 1:length(metadata_to_plot)) {
    print(metadata_to_plot[i])
    if (i %in% no_legend) {
      print(
        red_dim_plot(
          seurat_orig, x = "umap1", y = "umap2", color = metadata_to_plot[i], type = "cat"
        ) +
          NoLegend()
      )
    } else if (i %in% no_label) {
      print(red_dim_plot(seurat_orig, x = "umap1", y = "umap2", color = metadata_to_plot[i]))
    } else {
      print(
        red_dim_plot(
          seurat_orig, x = "umap1", y = "umap2", color = metadata_to_plot[i], type = "cat"
        )
      )
    }
  }
  print("nFeature_SCT")
  red_dim_plot(seurat_orig, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont")
  print("nCount_SCT")
  red_dim_plot(seurat_orig, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont")
  if (length(mito) != 0) {
    print("subsets_mito_percent")
    red_dim_plot(
      seurat_orig, x = "umap1", y = "umap2", color = "subsets_mito_percent", type = "cont"
    )
  }
}
```

# `r subsets[subset_counter]`

We rerun a pipeline on the subset that applies sctransform normalization and scaling, PCA, UMAP reduction, and Louvain clustering.
The following nuisance variables are regressed out: `r vars_to_regress`.

```{r}
sub_name <- subsets[subset_counter]
subset_counter <- subset_counter + 1

if ("neuronal" %in% unlist(strsplit(sub_name, "_"))) {
  if (organism == "human") {
    genes <- "SNAP25"
  } else if (organism == "mouse") {
    genes <- "Snap25"
  }
  if (length(unlist(strsplit(sub_name, "_"))) == 1) {
    FeaturePlot(
      seurat_orig, features = genes, order = TRUE, cols = c("lightgrey", "red"), max.cutoff = "q10"
    )
  } else {
    FeaturePlot(
      seurat, features = genes, order = TRUE, cols = c("lightgrey", "red"), max.cutoff = "q10"
    )
  }
}

# TODO: Remove hardcoding and move to `boilerplate`.
seurat <- subset(seurat_orig, subset = joint_region_label == "ENT")
seurat

if ("neuronal" %in% unlist(strsplit(sub_name, "_"))) {
  if (organism == "human") {
    genes <- "SNAP25"
  } else if (organism == "mouse") {
    genes <- "Snap25"
  }
  FeaturePlot(
    seurat, features = genes, order = TRUE, cols = c("lightgrey", "red"), max.cutoff = "q10"
  )
}

print("Remove genes not expressed in any cell.")
DefaultAssay(seurat) <- "RNA"
dim(seurat)
seurat <- seurat[rowSums(as.matrix(GetAssayData(seurat, slot = "counts")) > 0) > 0, ]
dim(seurat)

seurat <- cluster_pipeline(
  seurat, assets_dir = assets_dir, analysis_no = analysis_no, sub_name = sub_name,
  organism = organism, vars_to_regress = vars_to_regress, parallel_override = parallel_override
)
seurat

for (i in 1:length(metadata_to_plot)) {
  print(metadata_to_plot[i])
  if (i %in% large_plot) {
    print(
      red_dim_plot(seurat, x = "umap1", y = "umap2", color = metadata_to_plot[i], type = "cat") +
        NoLegend()
    )
  } else {
    print(red_dim_plot(seurat, x = "umap1", y = "umap2", color = metadata_to_plot[i], type = "cat"))
  }
}
print("nFeature_SCT")
red_dim_plot(seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont")
print("nCount_SCT")
red_dim_plot(seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont")
if (length(mito) != 0) {
  print("subsets_mito_percent")
  red_dim_plot(seurat, x = "umap1", y = "umap2", color = "subsets_mito_percent", type = "cont")
}
```

# `r subsets[subset_counter]`

We rerun a pipeline on the subset that applies sctransform normalization and scaling, PCA, UMAP reduction, and Louvain clustering.
The following nuisance variables are regressed out: `r vars_to_regress`.

```{r}
sub_name <- subsets[subset_counter]
subset_counter <- subset_counter + 1

if ("neuronal" %in% unlist(strsplit(sub_name, "_"))) {
  if (organism == "human") {
    genes <- "SNAP25"
  } else if (organism == "mouse") {
    genes <- "Snap25"
  }
  if (length(unlist(strsplit(sub_name, "_"))) == 1) {
    FeaturePlot(
      seurat_orig, features = genes, order = TRUE, cols = c("lightgrey", "red"), max.cutoff = "q10"
    )
  } else {
    FeaturePlot(
      seurat, features = genes, order = TRUE, cols = c("lightgrey", "red"), max.cutoff = "q10"
    )
  }
}

# TODO: Remove hardcoding and move to `boilerplate`.
seurat <- subset(seurat, idents = setdiff(1:length(unique(seurat$seurat_clusters)) - 1, 25))
seurat

if ("neuronal" %in% unlist(strsplit(sub_name, "_"))) {
  if (organism == "human") {
    genes <- "SNAP25"
  } else if (organism == "mouse") {
    genes <- "Snap25"
  }
  FeaturePlot(
    seurat, features = genes, order = TRUE, cols = c("lightgrey", "red"), max.cutoff = "q10"
  )
}

print("Remove genes not expressed in any cell.")
DefaultAssay(seurat) <- "RNA"
dim(seurat)
seurat <- seurat[rowSums(as.matrix(GetAssayData(seurat, slot = "counts")) > 0) > 0, ]
dim(seurat)

seurat <- cluster_pipeline(
  seurat, assets_dir = assets_dir, analysis_no = analysis_no, sub_name = sub_name,
  organism = organism, vars_to_regress = vars_to_regress, parallel_override = parallel_override
)
seurat

for (i in 1:length(metadata_to_plot)) {
  print(metadata_to_plot[i])
  if (i %in% large_plot) {
    print(
      red_dim_plot(seurat, x = "umap1", y = "umap2", color = metadata_to_plot[i], type = "cat") +
        NoLegend()
    )
  } else {
    print(red_dim_plot(seurat, x = "umap1", y = "umap2", color = metadata_to_plot[i], type = "cat"))
  }
}
print("nFeature_SCT")
red_dim_plot(seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont")
print("nCount_SCT")
red_dim_plot(seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont")
if (length(mito) != 0) {
  print("subsets_mito_percent")
  red_dim_plot(seurat, x = "umap1", y = "umap2", color = "subsets_mito_percent", type = "cont")
}
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r, references}
sessionInfo()
```
