---
title: '03 Vulnerable Types - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "03-vulnerable-types.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

In this document we probe for vulnerable cell types.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r, boilerplate}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

#    This file is part of tau-vulnerability.
#    Copyright (C) 2019-2020  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# -------------------------------------

packages <- c("conflicted", "Seurat", "dplyr", "ggplot2", "ggrepel")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 3
organism <- "mouse"
subset_names <- "ENT_neuronal"
options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(fig.width = 15, fig.height = 10)

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Backed up caches and data.
tmp_dir <- file.path(getwd(), "..", "..", "tmp", data_name)
results_dir <- file.path(getwd(), "..", "..", "results", data_name)

if (!dir.exists(tmp_dir)) {
  dir.create(tmp_dir, recursive = TRUE)
}

# Unique cache directory for each analysis number.
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}
```

We use data saved from previous analyses.

```{r, prep}
# TODO: Don't hardcode size of list.
subset_list <- c(
  readRDS(
    file.path(
      assets_dir, "cache", paste0("0", analysis_no - 1), paste0(subset_names, "_seurat.rds")
    )
  )
)

for (i in 1:length(subset_list)) {
  print(subset_names[i])
  print(
    red_dim_plot(
      subset_list[[i]], x = "umap1", y = "umap2", color = "seurat_clusters", type = "cat"
    )
  )
  if (organism == "human") {
    plots <- FeaturePlot(
      subset_list[[i]], features = c("SLC17A7", "GAD1"), order = TRUE, blend = TRUE,
      combine = FALSE, blend.threshold = 0, max.cutoff = "q10"
    )
    print(CombinePlots(plots[3:4], legend = "none"))
  } else if (organism == "mouse") {
    plots <- FeaturePlot(
      subset_list[[i]], features = c("Slc17a7", "Gad1"), order = TRUE, blend = TRUE,
      combine = FALSE, blend.threshold = 0, max.cutoff = "q10"
    )
    print(CombinePlots(plots[3:4], legend = "none"))
  }
}
```

# Neocortex L2-3 EX

```{r, neocortex-L23-EX}
if (organism == "human") {
  genes <- c("RELN", "NEFH")
} else if (organism == "mouse") {
  genes <- c("Reln", "Nefh")
}

gene_plot(subset_list, subset_names = subset_names, genes = genes)
```

# BA9 L2-3 EX

## Superficial

```{r, BA9-L23-EX-superficial}
if (organism == "human") {
  genes <- c("TBR1", "SATB2", "WFS1", "CUX2", "LAMP5", "SERPINE2")
} else if (organism == "mouse") {
  genes <- c("Tbr1", "Satb2", "Wfs1", "Cux2", "Lamp5", "Serpine2")
}

gene_plot(subset_list, subset_names = subset_names, genes = genes)
```

## Deep

```{r, BA9-L23-EX-deep}
if (organism == "human") {
  genes <- c("CUX2", "COL5A2")
} else if (organism == "mouse") {
  genes <- c("Cux2", "Col5a2")
}

gene_plot(subset_list, subset_names = subset_names, genes = genes)
```

# BA9 L4-5 EX

## 1

```{r, BA9-L45-EX-1}
if (organism == "human") {
  genes <- c("RORB", "GAL1")
} else if (organism == "mouse") {
  genes <- c("Rorb", "Gal1")
}

gene_plot(subset_list, subset_names = subset_names, genes = genes)
```

## 2

```{r, BA9-L45-EX-2}
if (organism == "human") {
  genes <- c("RORB", "PCP4")
} else if (organism == "mouse") {
  genes <- c("Rorb", "Pcp4")
}

gene_plot(subset_list, subset_names = subset_names, genes = genes)
```

# BA9 L5-6 EX

```{r, BA9-L56-EX}
if (organism == "human") {
  genes <- c("NFIA", "THEMIS")
} else if (organism == "mouse") {
  genes <- c("Nfia", "Themis")
}

gene_plot(subset_list, subset_names = subset_names, genes = genes)
```

# BA10 L4 EX

```{r, BA10-L4-EX}
if (organism == "human") {
  genes <- c("LINGO1", "RASGEF1B", "SLC26A3")
} else if (organism == "mouse") {
  genes <- c("Lingo1", "Rasgef1b", "Slc26a3")
}

gene_plot(subset_list, subset_names = subset_names, genes = genes)
```

# BA9 Chandelier

GAD2 expresssion should be low!

```{r, BA9-chandelier}
if (organism == "human") {
  genes <- c("PVALB", "GAD1", "GAD2", "SCUBE3")
} else if (organism == "mouse") {
  genes <- c("Pvalb", "Gad1", "Gad2", "Scube3")
}

gene_plot(subset_list, subset_names = subset_names, genes = genes)
```

# References

This is the concluding section of the document. Here we write relevant results to disk, output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
