---
title: '05 GSVA - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "05-gsva.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

In this document we perform GSVA analysis.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name `r unlist(strsplit(getwd(), "/"))[6]`.

```{r, boilerplate}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

#    This file is part of tau-vulnerability.
#    Copyright (C) 2019-2020  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# -------------------------------------

packages <- c(
  "conflicted", "RColorBrewer", "Seurat", "dplyr", "ggplot2", "ggrepel", "GSEABase", "GSVA",
  "parallel", "limma", "future", "matrixStats", "GO.db", "DT", "pheatmap"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 5
organism <- "mouse"
subset_names <- "ENT_neuronal"
download_name <- "smrt" # Name of download file corresponding to dataset.
color <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100)
color2 <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100)
color2[50] <- "#DDDDDD"
color_legend <- c("lightgray", "darkgray")
margins <- c(2, 50)
options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 10)

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Backed up caches and data.
tmp_dir <- file.path(getwd(), "..", "..", "tmp", data_name)
results_dir <- file.path(getwd(), "..", "..", "results", data_name)

if (!dir.exists(tmp_dir)) {
  dir.create(tmp_dir, recursive = TRUE)
}

# Unique cache directory for each analysis number.
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}
```

We use data saved from previous analyses.

```{r, prep}
seurat <- readRDS(
  file.path(assets_dir, "cache", paste0("0", analysis_no - 3), paste0(subset_names, "_seurat.rds"))
)
seurat
red_dim_plot(seurat, x = "umap1", y = "umap2", color = "seurat_clusters", type = "cat")
```

# Vuln. L2/3 EX vs. Remaining EX/IN

```{r, L23-EX-vs-remain-EX-IN-edit}
# This chunk typically needs to be edited per dataset.
# ----------------------------------------------------

sub <- vector(mode = "list", length = 2)
sub[[1]] <- c(12, 7, 10, 22, 19, 21, 13, 3, 18, 14, 4, 6, 16, 9, 1, 23, 11, 17, 5)
sub[[2]] <- c(15, 2, 0, 20, 8)
conditions <- c("Remaining EX/IN", "Early Affected L2/3 EX")
save_name <- "L23_EX_vs_remain_EX_IN"
gene_set_list <- c("c5.bp.v7.0.symbols", "AD_GO_names")
if (organism == "mouse") {
  gene_set_list <- paste(gene_set_list, "_mouse", sep = "")
}
```

```{r, L23-EX-vs-remain-EX-IN}
# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]])
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
print(conditions[1])
dim(condition1_mat)
red_dim_plot(condition1_seurat, x = "umap1", y = "umap2", color = "seurat_clusters", type = "cat")

# Now for condition 2.
condition2_seurat <- subset(seurat, idents = sub[[2]])
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
print(conditions[2])
dim(condition2_mat)
red_dim_plot(condition2_seurat, x = "umap1", y = "umap2", color = "seurat_clusters", type = "cat")

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
print("Basic DE analysis.")
print(topTable(fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05))

# GSVA can be memory hungry.
rm(condition1_seurat, condition2_seurat, condition1_mat, condition2_mat)
gc()

for (i in 1:length(gene_set_list)) {
  # The gene set to be used.
  gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(gene_set_list[i], ".gmt")))
  gene_sets

  # Run GSVA and cache the results.
  rds <- file.path(
    assets_dir, "cache", paste0("0", analysis_no), paste0(save_name, "_", gene_set_list[i], ".rds")
  )
  if (file.exists(rds)) {
    gsva <- readRDS(rds)
  } else {
    gsva <- gsva(comb_mat, gene_sets, min.sz = 10, max.sz = 500)
    saveRDS(gsva, rds)
  }

  # Fit another linear model, this time performing differential pathway analysis on GSVA output.
  fit <- lmFit(gsva, design)
  fit <- eBayes(fit)
  DE_gene_sets <- topTable(
    fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
  )
  print(gene_set_list[i])
  print(DE_gene_sets)

  gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
  names(sample_color_map) <- colnames(gsva_cutoff)
  gene_set_clustering <- hclust(
    as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete"
  )
  heatmap(
    gsva_cutoff, ColSideColors = sample_color_map,
    xlab = data_name, margins = margins,
    col = color, ylab = gene_set_list[i],
    labRow = toupper(gsub("^GO_|^REACTOME_", "", rownames(gsva_cutoff))), labCol = " ",
    Rowv = as.dendrogram(gene_set_clustering), Colv = NA
  )
  legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
}
```

We revisit the GO Biological Process GSVA results and retrieve terms also appearing in the AD GO terms list.

```{r, L23-EX-vs-remain-EX-IN-AD-GO-BP}
gsva <- readRDS(
  file.path(
    assets_dir, "cache",
    paste0("0", analysis_no), paste0(save_name, "_", gene_set_list[1], ".rds")
  )
)
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
comb1 <- fit
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = Inf, adjust.method = "holm", p.value = 1e-100
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
rownames(gsva_cutoff) <- gsub("^GO_", "", rownames(gsva_cutoff))

gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(gene_set_list[2], ".gmt")))
AD_GO_names <- lapply(names(gene_sets), function(x) { toupper(x) } )
AD_GO_names <- unlist(AD_GO_names)
AD_GO_names <- gsub("-", "_", AD_GO_names)

gsva_cutoff <- gsva_cutoff[(rownames(gsva_cutoff) %in% AD_GO_names), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete"
)
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map, xlab = data_name, margins = margins,
  col = color, ylab = "AD Relevant GO BP", labRow = rownames(gsva_cutoff), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

To further drill down the list, we calculate differentially expressed genes (DEGs) and filter out hits within the pathways.

```{r, L23-EX-vs-remain-EX-IN-DEG}
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0(save_name, "_", "DEG", ".rds")
)
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  parallel_plan(seurat)
  markers <- FindMarkers(seurat, ident.1 = sub[[2]], ident.2 = sub[[1]], verbose = FALSE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
if (length(rownames(markers)) <= 25) {
  DotPlot(seurat, features = rownames(markers), cols = c("blue", "red")) + RotatedAxis()
} else {
  DotPlot(seurat, features = rownames(markers)[1:25], cols = c("blue", "red")) + RotatedAxis()
}

gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(gene_set_list[1], ".gmt")))
for (i in 1:5) {
  print(gene_sets[[paste0("GO_", rownames(gsva_cutoff)[i])]]@shortDescription)
  genes <- which(
    gene_sets[[paste0("GO_", rownames(gsva_cutoff)[i])]]@geneIds %in% rownames(markers)
  )
  genes <- gene_sets[[paste0("GO_", rownames(gsva_cutoff)[i])]]@geneIds[genes]
  print(genes)
  print(
    DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
      NoLegend() +
      scale_fill_gradient2(
        low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
        high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
        guide = "colourbar", aesthetics = "fill"
      )
  )
}
```

Finally, we use a GO slim to group together hierarchically similar GO terms.

```{r, L23-EX-vs-remain-EX-IN-slim}
# Convert GO names to IDs.
GO_terms <- Term(GOTERM)
GO_terms <- toupper(GO_terms)
GO_terms <- gsub(" ", "_", GO_terms)
GO_terms <- gsub("-", "_", GO_terms)
gsva_cutoff <- gsva_cutoff[which(rownames(gsva_cutoff) %in% GO_terms), ]
GO_terms <- GO_terms[which(GO_terms %in% rownames(gsva_cutoff))]
GO_terms <- GO_terms[match(rownames(gsva_cutoff), GO_terms)]
rownames(gsva_cutoff) <- names(GO_terms)

# Summarize terms by their GO slim.
go_collection <- GOCollection(rownames(gsva_cutoff))
slim_obo <- getOBOCollection("http://data.pantherdb.org/PANTHER14.1/ontology/PANTHERGOslim.obo")
slim <- goSlim(go_collection, slimCollection = slim_obo, ontology = "BP")
slim <- slim[slim$Count > 0, ]
datatable_custom(slim)
```

## Conserved Pathways

While the largest GSVA object is loaded, we also use the IQR to look at pathways conserved and not differentially enriched between cells.

```{r, conserved}
gsva_cutoff <- gsva[order(rowIQRs(gsva)), ]

gsva_cutoff <- gsva_cutoff[1:25, ]
names(sample_color_map) <- colnames(gsva_cutoff)
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = margins,
  col = color, ylab = gene_set_list[i],
  labRow = toupper(gsub("^GO_|^REACTOME_", "", rownames(gsva_cutoff))), labCol = " ",
  Rowv = NA, Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# Vuln. L2/3 EX vs. Remaining EX

```{r, L23-EX-vs-remain-EX-edit}
# This chunk typically needs to be edited per dataset.
# ----------------------------------------------------

sub <- vector(mode = "list", length = 2)
sub[[1]] <- c(14, 4, 6, 16, 9, 1, 23, 11, 17, 5)
sub[[2]] <- c(15, 2, 0, 20, 8)
conditions <- c("Remaining EX", "Early Affected L2/3 EX")
save_name <- "L23_EX_vs_remain_EX"
gene_set_list <- c("c5.bp.v7.0.symbols", "AD_GO_names")
if (organism == "mouse") {
  gene_set_list <- paste(gene_set_list, "_mouse", sep = "")
}
```

```{r, L23-EX-vs-remain-EX}
# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]])
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
print(conditions[1])
dim(condition1_mat)
red_dim_plot(condition1_seurat, x = "umap1", y = "umap2", color = "seurat_clusters", type = "cat")

# Now for condition 2.
condition2_seurat <- subset(seurat, idents = sub[[2]])
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
print(conditions[2])
dim(condition2_mat)
red_dim_plot(condition2_seurat, x = "umap1", y = "umap2", color = "seurat_clusters", type = "cat")

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
print("Basic DE analysis.")
print(topTable(fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05))

# GSVA can be memory hungry.
rm(condition1_seurat, condition2_seurat, condition1_mat, condition2_mat)
gc()

for (i in 1:length(gene_set_list)) {
  # The gene set to be used.
  gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(gene_set_list[i], ".gmt")))
  gene_sets

  # Run GSVA and cache the results.
  rds <- file.path(
    assets_dir, "cache", paste0("0", analysis_no), paste0(save_name, "_", gene_set_list[i], ".rds")
  )
  if (file.exists(rds)) {
    gsva <- readRDS(rds)
  } else {
    gsva <- gsva(comb_mat, gene_sets, min.sz = 10, max.sz = 500)
    saveRDS(gsva, rds)
  }

  # Fit another linear model, this time performing differential pathway analysis on GSVA output.
  fit <- lmFit(gsva, design)
  fit <- eBayes(fit)
  DE_gene_sets <- topTable(
    fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
  )
  print(gene_set_list[i])
  print(DE_gene_sets)

  gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
  names(sample_color_map) <- colnames(gsva_cutoff)
  gene_set_clustering <- hclust(
    as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete"
  )
  heatmap(
    gsva_cutoff, ColSideColors = sample_color_map,
    xlab = data_name, margins = margins,
    col = color, ylab = gene_set_list[i],
    labRow = toupper(gsub("^GO_|^REACTOME_", "", rownames(gsva_cutoff))), labCol = " ",
    Rowv = as.dendrogram(gene_set_clustering), Colv = NA
  )
  legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
}
```

We revisit the GO Biological Process GSVA results and retrieve terms also appearing in the AD GO terms list.

```{r, L23-EX-vs-remain-EX-AD-GO-BP}
gsva <- readRDS(
  file.path(
    assets_dir, "cache",
    paste0("0", analysis_no), paste0(save_name, "_", gene_set_list[1], ".rds")
  )
)
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
comb2 <- fit
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = Inf, adjust.method = "holm", p.value = 1e-50
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
rownames(gsva_cutoff) <- gsub("^GO_", "", rownames(gsva_cutoff))

gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(gene_set_list[2], ".gmt")))
AD_GO_names <- lapply(names(gene_sets), function(x) { toupper(x) } )
AD_GO_names <- unlist(AD_GO_names)
AD_GO_names <- gsub("-", "_", AD_GO_names)

gsva_cutoff <- gsva_cutoff[(rownames(gsva_cutoff) %in% AD_GO_names), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete"
)
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map, xlab = data_name, margins = margins,
  col = color, ylab = "AD Relevant GO BP", labRow = rownames(gsva_cutoff), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

To further drill down the list, we calculate differentially expressed genes (DEGs) and filter out hits within the pathways.

```{r, L23-EX-vs-remain-EX-DEG}
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0(save_name, "_", "DEG", ".rds")
)
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  parallel_plan(seurat)
  markers <- FindMarkers(seurat, ident.1 = sub[[2]], ident.2 = sub[[1]], verbose = FALSE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
if (length(rownames(markers)) <= 25) {
  DotPlot(seurat, features = rownames(markers), cols = c("blue", "red")) + RotatedAxis()
} else {
  DotPlot(seurat, features = rownames(markers)[1:25], cols = c("blue", "red")) + RotatedAxis()
}

gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(gene_set_list[1], ".gmt")))
for (i in 1:5) {
  print(gene_sets[[paste0("GO_", rownames(gsva_cutoff)[i])]]@shortDescription)
  genes <- which(
    gene_sets[[paste0("GO_", rownames(gsva_cutoff)[i])]]@geneIds %in% rownames(markers)
  )
  genes <- gene_sets[[paste0("GO_", rownames(gsva_cutoff)[i])]]@geneIds[genes]
  print(genes)
  print(
    DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
      NoLegend() +
      scale_fill_gradient2(
        low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
        high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
        guide = "colourbar", aesthetics = "fill"
      )
  )
}
```

Finally, we use a GO slim to group together hierarchically similar GO terms.

```{r, L23-EX-vs-remain-EX-slim}
# Convert GO names to IDs.
GO_terms <- Term(GOTERM)
GO_terms <- toupper(GO_terms)
GO_terms <- gsub(" ", "_", GO_terms)
GO_terms <- gsub("-", "_", GO_terms)
gsva_cutoff <- gsva_cutoff[which(rownames(gsva_cutoff) %in% GO_terms), ]
GO_terms <- GO_terms[which(GO_terms %in% rownames(gsva_cutoff))]
GO_terms <- GO_terms[match(rownames(gsva_cutoff), GO_terms)]
rownames(gsva_cutoff) <- names(GO_terms)

# Summarize terms by their GO slim.
go_collection <- GOCollection(rownames(gsva_cutoff))
slim_obo <- getOBOCollection("http://data.pantherdb.org/PANTHER14.1/ontology/PANTHERGOslim.obo")
slim <- goSlim(go_collection, slimCollection = slim_obo, ontology = "BP")
slim <- slim[slim$Count > 0, ]
datatable_custom(slim)
```

# Vuln. L2/3 EX vs. IN

```{r, L23-EX-vs-IN-edit}
# This chunk typically needs to be edited per dataset.
# ----------------------------------------------------

sub <- vector(mode = "list", length = 2)
sub[[1]] <- c(18, 3, 13, 21, 22, 19, 10, 7, 12)
sub[[2]] <- c(15, 2, 0, 20, 8)
conditions <- c("IN", "Early Affected L2/3 EX")
save_name <- "L23_EX_vs_IN"
gene_set_list <- c("c5.bp.v7.0.symbols", "AD_GO_names")
if (organism == "mouse") {
  gene_set_list <- paste(gene_set_list, "_mouse", sep = "")
}
```

```{r, L23-EX-vs-IN}
# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]])
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
print(conditions[1])
dim(condition1_mat)
red_dim_plot(condition1_seurat, x = "umap1", y = "umap2", color = "seurat_clusters", type = "cat")

# Now for condition 2.
condition2_seurat <- subset(seurat, idents = sub[[2]])
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
print(conditions[2])
dim(condition2_mat)
red_dim_plot(condition2_seurat, x = "umap1", y = "umap2", color = "seurat_clusters", type = "cat")

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
print("Basic DE analysis.")
print(topTable(fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05))

# GSVA can be memory hungry.
rm(condition1_seurat, condition2_seurat, condition1_mat, condition2_mat)
gc()

for (i in 1:length(gene_set_list)) {
  # The gene set to be used.
  gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(gene_set_list[i], ".gmt")))
  gene_sets

  # Run GSVA and cache the results.
  rds <- file.path(
    assets_dir, "cache", paste0("0", analysis_no), paste0(save_name, "_", gene_set_list[i], ".rds")
  )
  if (file.exists(rds)) {
    gsva <- readRDS(rds)
  } else {
    gsva <- gsva(comb_mat, gene_sets, min.sz = 10, max.sz = 500)
    saveRDS(gsva, rds)
  }

  # Fit another linear model, this time performing differential pathway analysis on GSVA output.
  fit <- lmFit(gsva, design)
  fit <- eBayes(fit)
  DE_gene_sets <- topTable(
    fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
  )
  print(gene_set_list[i])
  print(DE_gene_sets)

  gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
  names(sample_color_map) <- colnames(gsva_cutoff)
  gene_set_clustering <- hclust(
    as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete"
  )
  heatmap(
    gsva_cutoff, ColSideColors = sample_color_map,
    xlab = data_name, margins = margins,
    col = color, ylab = gene_set_list[i],
    labRow = toupper(gsub("^GO_|^REACTOME_", "", rownames(gsva_cutoff))), labCol = " ",
    Rowv = as.dendrogram(gene_set_clustering), Colv = NA
  )
  legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
}
```

We revisit the GO Biological Process GSVA results and retrieve terms also appearing in the AD GO terms list.

```{r, L23-EX-vs-IN-AD-GO-BP}
gsva <- readRDS(
  file.path(
    assets_dir, "cache",
    paste0("0", analysis_no), paste0(save_name, "_", gene_set_list[1], ".rds")
  )
)
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
comb3 <- fit
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = Inf, adjust.method = "holm", p.value = 1e-250
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
rownames(gsva_cutoff) <- gsub("^GO_", "", rownames(gsva_cutoff))

gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(gene_set_list[2], ".gmt")))
AD_GO_names <- lapply(names(gene_sets), function(x) { toupper(x) } )
AD_GO_names <- unlist(AD_GO_names)
AD_GO_names <- gsub("-", "_", AD_GO_names)

gsva_cutoff <- gsva_cutoff[(rownames(gsva_cutoff) %in% AD_GO_names), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete"
)
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map, xlab = data_name, margins = margins,
  col = color, ylab = "AD Relevant GO BP", labRow = rownames(gsva_cutoff), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

To further drill down the list, we calculate differentially expressed genes (DEGs) and filter out hits within the pathways.

```{r, L23-EX-vs-IN-DEG}
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0(save_name, "_", "DEG", ".rds")
)
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  parallel_plan(seurat)
  markers <- FindMarkers(seurat, ident.1 = sub[[2]], ident.2 = sub[[1]], verbose = FALSE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
if (length(rownames(markers)) <= 25) {
  DotPlot(seurat, features = rownames(markers), cols = c("blue", "red")) + RotatedAxis()
} else {
  DotPlot(seurat, features = rownames(markers)[1:25], cols = c("blue", "red")) + RotatedAxis()
}

gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(gene_set_list[1], ".gmt")))
for (i in 1:5) {
  print(gene_sets[[paste0("GO_", rownames(gsva_cutoff)[i])]]@shortDescription)
  genes <- which(
    gene_sets[[paste0("GO_", rownames(gsva_cutoff)[i])]]@geneIds %in% rownames(markers)
  )
  genes <- gene_sets[[paste0("GO_", rownames(gsva_cutoff)[i])]]@geneIds[genes]
  print(genes)
  print(
    DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
      NoLegend() +
      scale_fill_gradient2(
        low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
        high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
        guide = "colourbar", aesthetics = "fill"
      )
  )
}
```

Finally, we use a GO slim to group together hierarchically similar GO terms.

```{r, L23-EX-vs-IN-slim}
# Convert GO names to IDs.
GO_terms <- Term(GOTERM)
GO_terms <- toupper(GO_terms)
GO_terms <- gsub(" ", "_", GO_terms)
GO_terms <- gsub("-", "_", GO_terms)
gsva_cutoff <- gsva_cutoff[which(rownames(gsva_cutoff) %in% GO_terms), ]
GO_terms <- GO_terms[which(GO_terms %in% rownames(gsva_cutoff))]
GO_terms <- GO_terms[match(rownames(gsva_cutoff), GO_terms)]
rownames(gsva_cutoff) <- names(GO_terms)

# Summarize terms by their GO slim.
go_collection <- GOCollection(rownames(gsva_cutoff))
slim_obo <- getOBOCollection("http://data.pantherdb.org/PANTHER14.1/ontology/PANTHERGOslim.obo")
slim <- goSlim(go_collection, slimCollection = slim_obo, ontology = "BP")
slim <- slim[slim$Count > 0, ]
datatable_custom(slim)
```

# Combined

```{r, L23-EX-combined}
names <- c("Early Affected L2/3 EX", "Remaining EX/IN", "Remaining EX", "IN")

sets <- list(
  topTable(comb1, number = Inf, coef = names[1], adjust.method = "holm", p.value = 1e-150),
  topTable(comb2, number = Inf, coef = names[1], adjust.method = "holm", p.value = 1e-100),
  topTable(comb3, number = Inf, coef = names[1], adjust.method = "holm", p.value = 1e-300)
)
sets <- lapply(sets, function(x) cbind(x , rownames(x)))
sets <- lapply(sets, function(x) x[ , c(dim(x)[2], 1)])
colnames(sets[[1]])[2] <- paste0(names[2], " lods")
colnames(sets[[2]])[2] <- paste0(names[3], " lods")
colnames(sets[[3]])[2] <- paste0(names[4], " lods")
merged_set <- Reduce(function(...) merge(..., all = TRUE), list(
  sets[[1]][ , c(1, 2)], sets[[2]][ , c(1, 2)], sets[[3]][ , c(1, 2)]
))

set1 <- merged_set[order(-abs(merged_set[[paste0(names[2], " lods")]])), ]
set2 <- merged_set[order(-abs(merged_set[[paste0(names[3], " lods")]])), ]
set3 <- merged_set[order(-abs(merged_set[[paste0(names[4], " lods")]])), ]

merged_top <- Reduce(function(...) merge(..., all = TRUE), list(
  set1[ , c(1, 2)], set2[ , c(1, 3)], set3[ , c(1, 4)]
))
top_genes <- merged_top$`rownames(x)`
merged_top <- merged_top[ , 2:dim(merged_top)[2]]
rownames(merged_top) <- top_genes
datatable_custom(merged_top) %>%
  formatRound(columns = c(1:dim(merged_top)[2]), digits = 2)

# Generate heatmap, using grey for 0 in color palette to indicate `NA` values.
merged_top <- as.matrix(merged_top)
rownames(merged_top) <- top_genes
merged_top <- replace(merged_top, is.na(merged_top), 0)
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 0, fontsize = 10, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 4
)

# Generate heatmap without `NA` values.
merged_top[merged_top == 0] <- NA
merged_top <- merged_top[complete.cases(merged_top), ]
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 0, fontsize = 9, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 9
)

markers <- readRDS(
  file.path(assets_dir, "cache", paste0("0", analysis_no), "L23_EX_vs_remain_EX_IN_DEG.rds")
)
markers <- markers[markers$p_val_adj < 0.05, ]

gsva <- readRDS(
  file.path(
    assets_dir, "cache",
    paste0("0", analysis_no), paste0("L23_EX_vs_remain_EX_IN", "_", gene_set_list[1], ".rds")
  )
)
seurat[["GSVA"]] <- CreateAssayObject(counts = gsva)

gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(gene_set_list[1], ".gmt")))
hyphen_names <- gsub("_", "-", rownames(merged_top))
for (i in 1:length(rownames(merged_top))) {
  print(gene_sets[[rownames(merged_top)[i]]]@shortDescription)
  genes <- which(
    gene_sets[[rownames(merged_top)[i]]]@geneIds %in% rownames(markers)
  )
  genes <- gene_sets[[rownames(merged_top)[i]]]@geneIds[genes]
  genes
  print(
    DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
      NoLegend() +
      scale_fill_gradient2(
        low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
        high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
        guide = "colourbar", aesthetics = "fill"
      )
  )
  print(
    FeaturePlot(
      seurat, features = hyphen_names[i], order = TRUE,
      cols = c("blue", "red"), min.cutoff = -0.25, max.cutoff = 0.25
    )
  )
}
```

# References

This is the concluding section of the document. Here we write relevant results to disk, output the `sessionInfo`, and create a bibliography for works cited.

```{r, references}
sessionInfo()
```
