---
title: '05 GSVA - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "05-gsva.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

In this document we perform GSVA analysis.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name `r unlist(strsplit(getwd(), "/"))[6]`.

```{r, boilerplate}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

#    This file is part of tau-vulnerability.
#    Copyright (C) 2019-2020  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# -------------------------------------

packages <- c(
  "conflicted", "Seurat", "dplyr", "ggplot2", "ggrepel",
  "biomaRt", "SCnorm", "SingleCellExperiment", "GSEABase", "limma",
  "GSVA", "RColorBrewer", "parallel", "DT", "pheatmap"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 5
organism <- "mouse"
batch_metadata <- "donor_id"
vars_to_regress <- c("age_label", "sex_label")
parallel_override <- NULL # See function `parallel_plan` in `utils.R`.
subset_names <- "ENT_HIP_neuronal"
download_name <- "smrt" # Name of download file corresponding to dataset.
color <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100)
color[50] <- "#DDDDDD"
options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(fig.width = 15, fig.height = 15)

# TODO: Make less verbose.
idents <- c(14, 36, 0, 15)
idents[[1]] <- 14
idents[[2]] <- 36
idents[[3]] <- 0
idents[[4]] <- 15

ident_names <- c("EC_L2", "EC_L6", "CA1", "CA3") 

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Backed up caches and data.
tmp_dir <- file.path(getwd(), "..", "..", "tmp", data_name)
results_dir <- file.path(getwd(), "..", "..", "results", data_name)

if (!dir.exists(tmp_dir)) {
  dir.create(tmp_dir, recursive = TRUE)
}

# Unique cache directory for each analysis number.
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}
```

We use data saved from previous analyses.

```{r, prep}
# TODO: Make chunk loopable for all `subset_names` and remove hardcoding.
# -----------------------------------------------------------------------

seurat <- readRDS(
  file.path(
    assets_dir, "cache", paste0("0", analysis_no - 3), paste0(subset_names[1], "_seurat.rds")
  )
)
seurat
red_dim_plot(seurat, x = "umap1", y = "umap2", color = "seurat_clusters", type = "cat")

# Remove obsolete data.
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", "GO_BP_AD_filt_mouse.gmt"))
seurat <- seurat[rownames(seurat) %in% unique(unlist(geneIds(gene_sets))), ]

# Create subsets.
seurat <- subset(seurat, idents = c(idents[[1]], idents[[2]], idents[[3]], idents[[4]]))
seurat <- seurat[rowSums(as.matrix(GetAssayData(seurat, slot = "counts")) > 0) > 0, ]
condition1_seurat <- subset(seurat, idents = idents[[1]])
condition2_seurat <- subset(seurat, idents = idents[[2]])
condition3_seurat <- subset(seurat, idents = idents[[3]])
condition4_seurat <- subset(seurat, idents = idents[[4]])
seurat <- merge(
  condition1_seurat, y = list(condition2_seurat, condition3_seurat, condition4_seurat)
)

# Renormalize using SCnorm.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "scnorm.rds")
if (file.exists(rds)) {
  sce <- readRDS(rds)
} else {
  sce <- SingleCellExperiment(
    assays = list(
      counts = cbind(
        as.matrix(GetAssayData(condition1_seurat)), as.matrix(GetAssayData(condition2_seurat)),
        as.matrix(GetAssayData(condition3_seurat)), as.matrix(GetAssayData(condition4_seurat))
      )
    )
  )
  groups <- factor(
    rep(
      c(0, 1, 2, 3),
      c(
        dim(condition1_seurat)[2], dim(condition2_seurat)[2],
        dim(condition3_seurat)[2], dim(condition4_seurat)[2]
      )
    )
  )

  # Get gene lengths and GC content.
  if (organism == "human") {
    dataset = "hsapiens_gene_ensembl"
  } else if (organism == "mouse") {
    dataset = "mmusculus_gene_ensembl"
  }
  mart <- useEnsembl(biomart = "ensembl", dataset = dataset, host = "http://useast.ensembl.org/")
  attributes <- c(
    "external_gene_name", "start_position", "end_position", "percentage_gene_gc_content"
  )
  gene_anno <- getBM(
    attributes = attributes, filters = "external_gene_name", values = rownames(sce), mart = mart
  )
  sce <- sce[rownames(sce) %in% gene_anno$external_gene_name, ]
  gene_anno$length <- gene_anno$end_position - gene_anno$start_position
  len <- tapply(gene_anno$length, gene_anno$external_gene_name, mean)
  gcc <- tapply(gene_anno$percentage_gene_gc_content, gene_anno$external_gene_name, mean)
  rowData(sce) <- data.frame(gc_content = gcc, length = log(len))

  sce <- SCnorm(
    sce, Conditions = groups, withinSample = c(rowData(sce)$gc_content, rowData(sce)$length),
    useZerosToScale = TRUE, NCores = detectCores()
  )
  saveRDS(sce, rds)
}

rm(condition1_seurat, condition2_seurat, condition3_seurat, condition4_seurat)
seurat <- seurat[rownames(seurat) %in% rownames(sce), ]
seurat <- SetAssayData(seurat, slot = "data", new.data = normcounts(sce))
seurat <- SetAssayData(seurat, slot = "scale.data", new.data = log(normcounts(sce) + 1))
rm(sce)
seurat <- RenameIdents(seurat, `14` = "EC_L2")
seurat <- RenameIdents(seurat, `36` = "EC_L6")
seurat <- RenameIdents(seurat, `0` = "CA1")
seurat <- RenameIdents(seurat, `15` = "CA3")
seurat$idents <- seurat@active.ident
seurat
red_dim_plot(seurat, x = "umap1", y = "umap2", color = "idents", type = "cat")

hist(as.matrix(GetAssayData(seurat, slot = "scale.data")))
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "d.rds")
if (file.exists(rds)) {
  d <- readRDS(rds)
} else {
  d <- dist(t(scale(GetAssayData(seurat, slot = "scale.data"))))
  saveRDS(d, rds)
}
mds <- cmdscale(d = d, k = 2)
colnames(mds) <- paste0("mds_", 1:2)
seurat[["mds"]] <- CreateDimReducObject(
  embeddings = mds, key = "mds_", assay = DefaultAssay(seurat)
)
rm(mds)
add_df <- data.frame(Embeddings(seurat, reduction = "mds"))
names(add_df) <- paste0("mds", seq(ncol(add_df)))
seurat$mds1 <- add_df$mds1
seurat$mds2 <- add_df$mds2
red_dim_plot(seurat, x = "mds1", y = "mds2", color = "idents", type = "cat")
red_dim_plot(seurat, x = "mds1", y = "mds2", color = "sex_label", type = "cat")
red_dim_plot(seurat, x = "mds1", y = "mds2", color = "age_label", type = "cat")
red_dim_plot(seurat, x = "mds1", y = "mds2", color = "donor_id", type = "cat")

rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "gsva.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(as.matrix(GetAssayData(seurat, slot = "scale.data")), gene_sets)
  saveRDS(gsva, rds)
}

seurat[["GSVA"]] <- CreateAssayObject(counts = gsva)
rm(gsva)
DefaultAssay(seurat) <- "GSVA"
hist(as.matrix(GetAssayData(seurat, slot = "counts")))

umap <- as.matrix(data.frame(seurat$umap1, seurat$umap2))
colnames(umap) <- paste0("umap_", 1:2)
seurat[["umap"]] <- CreateDimReducObject(
  embeddings = umap, key = "umap_", assay = DefaultAssay(seurat)
)
rm(umap)
```

# Comparisons Against Early Affected EC L2 EX

```{r, vs-EC-L2}
design <- model.matrix(
  ~0 + seurat$idents, data = as.data.frame(GetAssayData(seurat, slot = "counts"))
)
colnames(design) <- rev(ident_names)
fit <- lmFit(
  as.data.frame(GetAssayData(seurat, slot = "counts")), design = design
)
contrast_mat <- makeContrasts(CA3-EC_L2, CA1-EC_L2, EC_L6-EC_L2, levels = design)
fit <- eBayes(contrasts.fit(fit, contrast_mat))
table <- topTable(fit, coef = 1, number = Inf)
results <- decideTests(fit)
EC_L2_up <- table[rownames(table) %in% rownames(results[rowSums(results) == -3, ]), ]
EC_L2_down <- table[rownames(table) %in% rownames(results[rowSums(results) == 3, ]), ]

datatable_custom(EC_L2_up)
datatable_custom(EC_L2_down)

DoHeatmap(seurat, slot = "counts", features = rownames(EC_L2_up)[1:40], size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
DoHeatmap(seurat, slot = "counts", features = rownames(EC_L2_up)[1:40], size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )

FeaturePlot(
  seurat, features = rownames(EC_L2_up)[1:4], order = TRUE, cols = c("lawngreen", "firebrick1"),
  ncol = 2
)
FeaturePlot(
  seurat, features = rownames(EC_L2_down)[1:4], order = TRUE, cols = c("lawngreen", "firebrick1"),
  ncol = 2
)

VlnPlot(seurat, features = rownames(EC_L2_up)[1:4], ncol = 2)
VlnPlot(seurat, features = rownames(EC_L2_down)[1:4], ncol = 2)
```

# References

This is the concluding section of the document. Here we write relevant results to disk, output the `sessionInfo`, and create a bibliography for works cited.

```{r, references}
sessionInfo()
```
