---
title: '01 AD GWAS - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "01-AD-GWAS.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

In this document we probe for vulnerable cell types.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r}
#    This file is part of tau-vulnerability.
#    Copyright (C) 2019-2020  Emir Turkes, UK DRI at UCL, Columbia University Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# This section should be checked per document.
# Blank variables should be set to NULL.
# --------------------------------------------
options(stringsAsFactors = FALSE)
packages <- c(
  "conflicted", "RColorBrewer", "biomaRt", "GO.db", "GOSemSim", "ComplexHeatmap", "clusterProfiler"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

color <- colorRampPalette(rev(brewer.pal(9, "RdBu")))(100)
# --------------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
analysis_no <- 1
data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
assets_dir <- file.path(getwd(), "..", "..", "assets") # Backed up data.

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path(getwd(), "..", "..", "cache", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

knitr::opts_chunk$set(fig.width = 14, fig.height = 10)
# ----------------------------------------------------------
```

# Mouse

```{r}
rds <- file.path(cache_dir, "genes.rds")
if (file.exists(rds)) {
  genes <- readRDS(rds)
} else {
  genes <- read.delim(
    file.path(
      assets_dir, "gene-sets",
      "gwas-association-downloaded_2020-05-12-EFO_0000249-withChildTraits.tsv"
    )
  )
  genes <- genes[order(genes$P.VALUE), ]
  genes$REPORTED.GENE.S. <- gsub(" ", "", genes$REPORTED.GENE.S.)
  genes <- unique(unlist(strsplit(genes$REPORTED.GENE.S., ",")))
  genes <- genes[-grep("ntergenic", genes)]
  genes <- sub("\\*", "", genes)
  genes <- genes[-grep("^NR$", genes)]
  genes <- unique(unlist(strsplit(genes, "x")))
  genes <- human_to_mouse_genes(genes)
  genes <- bitr(genes, "SYMBOL", "ENTREZID", "org.Mm.eg.db")
  saveRDS(genes, rds)
}

rds <- file.path(cache_dir, "go_data.rds")
if (file.exists(rds)) {
  go_data <- readRDS(rds)
} else {
  go_data <- godata("org.Mm.eg.db", "SYMBOL", "BP", FALSE)
  saveRDS(go_data, rds)
}

rds <- file.path(cache_dir, "sem_sim.rds")
if (file.exists(rds)) {
  sem_sim <- readRDS(rds)
} else {
  sem_sim <- mgeneSim(genes$SYMBOL, go_data, drop = NULL, verbose = FALSE)
  saveRDS(sem_sim, rds)
}

split <- 14
heatmap <- Heatmap(
  sem_sim, color, row_split = split, column_split = split,
  show_heatmap_legend = FALSE, show_row_names = FALSE, show_column_names = FALSE
)
draw(heatmap)

clusters <- vector("list", length(column_order(heatmap)))
for (i in seq_along(column_order(heatmap))) {
  clusters[[i]] <- genes$SYMBOL[column_order(heatmap)[[i]]]
  names(clusters)[i] <- paste0("cluster", i)
}

rds <- file.path(cache_dir, "compare_clusters_BP.rds")
if (file.exists(rds)) {
  compare_clusters_BP <- readRDS(rds)
} else {
  compare_clusters_BP <- compareCluster(clusters, OrgDb = "org.Mm.eg.db", ont = "BP")
  saveRDS(compare_clusters_BP, rds)
}

rds <- file.path(cache_dir, "compare_clusters_CC.rds")
if (file.exists(rds)) {
  compare_clusters_CC <- readRDS(rds)
} else {
  compare_clusters_CC <- compareCluster(clusters, OrgDb = "org.Mm.eg.db", ont = "CC")
  saveRDS(compare_clusters_CC, rds)
}

rds <- file.path(cache_dir, "compare_clusters_MF.rds")
if (file.exists(rds)) {
  compare_clusters_MF <- readRDS(rds)
} else {
  compare_clusters_MF <- compareCluster(clusters, OrgDb = "org.Mm.eg.db", ont = "MF")
  saveRDS(compare_clusters_MF, rds)
}

dotplot(compare_clusters_BP)
dotplot(compare_clusters_CC)
dotplot(compare_clusters_MF)
```

# References

This is the concluding section of the document.
Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
