---
title: '04 EC Reln vs. CA1 Wfs1 vs. CA3 Bok vs. IN Sst - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), "%B %d, %Y")`'
bibliography: '../../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "..",
    "results", unlist(strsplit(getwd(), "/"))[6],
    unlist(strsplit(getwd(), "/"))[7], "04-Reln-Wfs1-Bok-Sst.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

In this document we perform gene set enrichment and network analysis.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r}
#    This file is part of tau-vulnerability.
#    Copyright (C) 2019-2020  Emir Turkes, UK DRI at UCL, Columbia University Medical Center
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# -------------------------------------
options(stringsAsFactors = FALSE)
packages <- c(
  "conflicted", "scales", "RColorBrewer", "Seurat", "dplyr", "ggplot2", "ggrepel", "GSEABase",
  "biomaRt", "Biostrings", "SingleCellExperiment", "SCnorm", "viridis", "scater", "GSVA", "limma",
  "DT", "GOSemSim", "factoextra", "dendextend", "ComplexHeatmap", "GO.db", "tm", "reshape2",
  "igraph", "qgraph", "pals", "cluster", "future", "edgeR", "circlize", "visNetwork"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
conflict_prefer("strsplit", "Biostrings", quiet = TRUE)
conflict_prefer("intersect", "GSEABase", quiet = TRUE)
conflict_prefer("select", "AnnotationDbi", quiet = TRUE)
conflict_prefer("degree", "igraph", quiet = TRUE)
conflict_prefer("strsplit", "Biostrings", quiet = TRUE)
conflict_prefer("cpm", "edgeR", quiet = TRUE)
conflict_prefer("groups", "igraph", quiet = TRUE)
source(file.path("..", "..", "utils.R"))

nbclust_fun <- function(x, k) {
  list(cluster = cutree(hclust(get_dist((x), "spearman"), "ward.D2"), k))
}
`%notin%` <- Negate(`%in%`)

protocol <- c("mouse", "smart-seq", "single-cell", "reads") # See `cluster_pipeline` in `utils.R`.

# Metadata to plot after dimensionality reduction and clustering.
# Values in list can include "no_legend and/or "no_label" to exclude those.
# -------------------------------------------------------------------------
metadata_to_plot <- vector("list", 6)
names(metadata_to_plot) <- c(
  "seurat_clusters", "class_label", "subclass_label",
  "region_label", "donor_sex_label", "external_donor_name_id"
)
metadata_to_plot$region_label <- "no_label"
metadata_to_plot$donor_sex_label <- "no_label"
# -------------------------------------------------------------------------
# --------------------------------------------

gene_set_list <- "GO"
subset_names <- "EC_HIP_neuronal"
color <- colorRampPalette(rev(brewer.pal(9, "RdBu")))(100)
color3 <- colorRampPalette(rev(brewer.pal(11, "RdYlBu")))(100)

# Clusters of interest.
# ---------------------
idents <- vector("list", 4)
idents[[1]] <- 23
idents[[2]] <- 3
idents[[3]] <- 29
idents[[4]] <- 18
names(idents) <- c("EC_Reln", "CA1_Wfs1", "CA3_Bok", "IN_Sst")
# ---------------------
# -------------------------------------

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------
analysis_no <- 4
data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
assets_dir <- file.path(getwd(), "..", "..", "..", "assets") # Backed up data.
results_dir <- file.path(getwd(), "..", "..", "..", "results")

# Unique cache directory for each analysis number.
# ------------------------------------------------
cache_dir <- file.path(
  getwd(), "..", "..", "..", "cache", data_name, "05-gsva", paste0("0", analysis_no)
)
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# ------------------------------------------------

if (protocol[1] == "human") {
  organism <- "hsapiens"
} else if (protocol[1] == "mouse") {
  organism <- "mmusculus"
}

knitr::opts_chunk$set(fig.width = 12, fig.height = 12)
# ----------------------------------------------------------
```

```{r}
seurat <- readRDS(file.path(cache_dir, "..", "..", "02", paste0(subset_names[1], "_seurat.rds")))
seurat

for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat") + NoLegend())
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(seurat, "umap1", "umap2", names(metadata_to_plot)[i], "cat"))
  }
}
```

# Prep

```{r, fig.height = 5, fig.width = 7.5}
section_no <- 1
gene_set_name <- gene_set_list[section_no]

rds <- file.path(cache_dir, paste0(gene_set_name, "_gene_sets.rds"))
rds2 <- file.path(cache_dir, paste0(gene_set_name, "_gene_sets_filtered.rds"))
if (file.exists(rds) & file.exists(rds2)) {
  gene_sets <- readRDS(rds)
  gene_sets_filtered <- readRDS(rds2)
} else {
  gene_sets <- getGmt(
    file.path(assets_dir, "gene-sets", paste0(organism, ".", gene_set_name, ".ENSG.gmt"))
  )
  gene_sets <- filterGeneSets(gene_sets, 3, 500)
  gene_sets_filtered <- gene_sets # Put aside for further filtering later.
  for (i in seq(length(gene_sets@.Data))) {
    go_id <- gene_sets[[i]]@setName
    suppressWarnings(gene_sets[[i]]@setName <- gene_sets[[i]]@shortDescription)
    suppressWarnings(gene_sets[[i]]@shortDescription <- go_id)
  }

  filter_sets <- read.delim(file.path(assets_dir, "gene-sets", "genes-09-00593-s001-s5.tsv"))
  filter_sets <- filter_sets$GO.TERM
  gene_sets_filtered <- gene_sets_filtered[names(gene_sets_filtered) %in% filter_sets]
  for (i in seq(length(gene_sets_filtered@.Data))) {
    go_id <- gene_sets_filtered[[i]]@setName
    suppressWarnings(gene_sets_filtered[[i]]@setName <- gene_sets_filtered[[i]]@shortDescription)
    suppressWarnings(gene_sets_filtered[[i]]@shortDescription <- go_id)
  }

  saveRDS(gene_sets, rds)
  saveRDS(gene_sets_filtered, rds2)
}
gene_sets

open_targets_all <- read.csv(
  file.path(assets_dir, "gene-sets", "open-targets-all-mouse.csv"), FALSE
)
open_targets_all <- unique(open_targets_all$V1)

kampmann_comb_ECvuln_AD_CTRL_DE_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva", "01-ECvuln-AD-CTRL-DE-up.csv"
  ),
  FALSE
)
kampmann_comb_ECvuln_AD_CTRL_DE_up <- unique(kampmann_comb_ECvuln_AD_CTRL_DE_up$V1)
kampmann_comb_ECvuln_AD_CTRL_DE_strict_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva", "01-ECvuln-AD-CTRL-DE-strict-up.csv"
  ),
  FALSE
)
kampmann_comb_ECvuln_AD_CTRL_DE_strict_up <- unique(kampmann_comb_ECvuln_AD_CTRL_DE_strict_up$V1)
kampmann_comb_ECvuln_AD_CTRL_DE_stricter_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva", "01-ECvuln-AD-CTRL-DE-stricter-up.csv"
  ),
  FALSE
)
kampmann_comb_ECvuln_AD_CTRL_DE_stricter_up <- unique(
  kampmann_comb_ECvuln_AD_CTRL_DE_stricter_up$V1
)
kampmann_comb_ECvuln_AD_CTRL_DE_strictest_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "01-ECvuln-AD-CTRL-DE-strictest-up.csv"
  ),
  FALSE
)
kampmann_comb_ECvuln_AD_CTRL_DE_strictest_up <- unique(
  kampmann_comb_ECvuln_AD_CTRL_DE_strictest_up$V1
)
kampmann_comb_ECvuln_AD_CTRL_DE_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva", "01-ECvuln-AD-CTRL-DE-down.csv"
  ),
  FALSE
)
kampmann_comb_ECvuln_AD_CTRL_DE_down <- unique(kampmann_comb_ECvuln_AD_CTRL_DE_down$V1)
kampmann_comb_ECvuln_AD_CTRL_DE_strict_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva", "01-ECvuln-AD-CTRL-DE-strict-down.csv"
  ),
  FALSE
)
kampmann_comb_ECvuln_AD_CTRL_DE_strict_down <- unique(
  kampmann_comb_ECvuln_AD_CTRL_DE_strict_down$V1
)
kampmann_comb_ECvuln_AD_CTRL_DE_stricter_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "01-ECvuln-AD-CTRL-DE-stricter-down.csv"
  ),
  FALSE
)
kampmann_comb_ECvuln_AD_CTRL_DE_stricter_down <- unique(
  kampmann_comb_ECvuln_AD_CTRL_DE_stricter_down$V1
)
kampmann_comb_ECvuln_AD_CTRL_DE_strictest_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "01-ECvuln-AD-CTRL-DE-strictest-down.csv"
  ),
  FALSE
)
kampmann_comb_ECvuln_AD_CTRL_DE_strictest_down <- unique(
  kampmann_comb_ECvuln_AD_CTRL_DE_strictest_down$V1
)

kampmann_comb_SFGvuln_AD_CTRL_DE_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva", "03-SFGvuln-AD-CTRL-DE-up.csv"
  ),
  FALSE
)
kampmann_comb_SFGvuln_AD_CTRL_DE_up <- unique(kampmann_comb_SFGvuln_AD_CTRL_DE_up$V1)
kampmann_comb_SFGvuln_AD_CTRL_DE_strict_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "03-SFGvuln-AD-CTRL-DE-strict-up.csv"
  ),
  FALSE
)
kampmann_comb_SFGvuln_AD_CTRL_DE_strict_up <- unique(
  kampmann_comb_SFGvuln_AD_CTRL_DE_strict_up$V1
)
kampmann_comb_SFGvuln_AD_CTRL_DE_stricter_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva",
    "03-SFGvuln-AD-CTRL-DE-stricter-up.csv"
  ),
  FALSE
)
kampmann_comb_SFGvuln_AD_CTRL_DE_stricter_up <- unique(
  kampmann_comb_SFGvuln_AD_CTRL_DE_stricter_up$V1
)
kampmann_comb_SFGvuln_AD_CTRL_DE_strictest_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "03-SFGvuln-AD-CTRL-DE-strictest-up.csv"
  ),
  FALSE
)
kampmann_comb_SFGvuln_AD_CTRL_DE_strictest_up <- unique(
  kampmann_comb_SFGvuln_AD_CTRL_DE_strictest_up$V1
)
kampmann_comb_SFGvuln_AD_CTRL_DE_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva", "03-SFGvuln-AD-CTRL-DE-down.csv"
  ),
  FALSE
)
kampmann_comb_SFGvuln_AD_CTRL_DE_down <- unique(kampmann_comb_SFGvuln_AD_CTRL_DE_down$V1)
kampmann_comb_SFGvuln_AD_CTRL_DE_strict_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "03-SFGvuln-AD-CTRL-DE-strict-down.csv"
  ),
  FALSE
)
kampmann_comb_SFGvuln_AD_CTRL_DE_strict_down <- unique(
  kampmann_comb_SFGvuln_AD_CTRL_DE_strict_down$V1
)
kampmann_comb_SFGvuln_AD_CTRL_DE_stricter_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "03-SFGvuln-AD-CTRL-DE-stricter-down.csv"
  ),
  FALSE
)
kampmann_comb_SFGvuln_AD_CTRL_DE_stricter_down <- unique(
  kampmann_comb_SFGvuln_AD_CTRL_DE_stricter_down$V1
)
kampmann_comb_SFGvuln_AD_CTRL_DE_strictest_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "03-SFGvuln-AD-CTRL-DE-strictest-down.csv"
  ),
  FALSE
)
kampmann_comb_SFGvuln_AD_CTRL_DE_strictest_down <- unique(
  kampmann_comb_SFGvuln_AD_CTRL_DE_strictest_down$V1
)

kampmann_comb_SFGinvuln_AD_CTRL_DE_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva", "02-SFGinvuln-AD-CTRL-DE-up.csv"
  ),
  FALSE
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_up <- unique(kampmann_comb_SFGinvuln_AD_CTRL_DE_up$V1)
kampmann_comb_SFGinvuln_AD_CTRL_DE_strict_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "02-SFGinvuln-AD-CTRL-DE-strict-up.csv"
  ),
  FALSE
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_strict_up <- unique(
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strict_up$V1
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_stricter_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva",
    "02-SFGinvuln-AD-CTRL-DE-stricter-up.csv"
  ),
  FALSE
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_stricter_up <- unique(
  kampmann_comb_SFGinvuln_AD_CTRL_DE_stricter_up$V1
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_strictest_up <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "02-SFGinvuln-AD-CTRL-DE-strictest-up.csv"
  ),
  FALSE
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_strictest_up <- unique(
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strictest_up$V1
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva", "02-SFGinvuln-AD-CTRL-DE-down.csv"
  ),
  FALSE
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_down <- unique(kampmann_comb_SFGinvuln_AD_CTRL_DE_down$V1)
kampmann_comb_SFGinvuln_AD_CTRL_DE_strict_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "02-SFGinvuln-AD-CTRL-DE-strict-down.csv"
  ),
  FALSE
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_strict_down <- unique(
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strict_down$V1
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_stricter_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "02-SFGinvuln-AD-CTRL-DE-stricter-down.csv"
  ),
  FALSE
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_stricter_down <- unique(
  kampmann_comb_SFGinvuln_AD_CTRL_DE_stricter_down$V1
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_strictest_down <- read.csv(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb",
    "05-gsva", "02-SFGinvuln-AD-CTRL-DE-strictest-down.csv"
  ),
  FALSE
)
kampmann_comb_SFGinvuln_AD_CTRL_DE_strictest_down <- unique(
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strictest_down$V1
)

kampmann_all <- read.delim(
  file.path(
    results_dir, "kampmann-2020-snRNAseq", "comb", "05-gsva", "01-ECvuln-AD-CTRL-all.tsv"
  ),
  FALSE
)
kampmann_all <- unique(kampmann_all$V1)

# Create a Seurat object containing only clusters of interest with downsampling.
# Also convert gene symbols to ENSEMBL IDs.
# ------------------------------------------------------------------------------
rds <- file.path(cache_dir, paste0(gene_set_name, "_seurat_cleaned.rds"))
rds2 <- file.path(cache_dir, paste0(gene_set_name, "_gene_anno.rds"))
if (file.exists(rds) & file.exists(rds2)) {
  seurat <- readRDS(rds)
  gene_anno <- readRDS(rds2)
} else {

  # Remove obsolete data.
  # ---------------------
  DefaultAssay(seurat) <- "RNA"
  seurat[["SCT"]] <- NULL
  # ---------------------

  # Create subsets.
  # ---------------
  sample1_seurat <- subset(seurat, subset = external_donor_name_id == 475)
  expr <- FetchData(sample1_seurat, "seurat_clusters")
  sample1_seurat <- sample1_seurat[ , which(expr == idents[[1]])]
  sample2_seurat <- subset(seurat, subset = external_donor_name_id == 476)
  expr <- FetchData(sample2_seurat, "seurat_clusters")
  sample2_seurat <- sample2_seurat[ , which(expr == idents[[1]])]
  sample3_seurat <- subset(seurat, subset = external_donor_name_id == 527)
  expr <- FetchData(sample3_seurat, "seurat_clusters")
  sample3_seurat <- sample3_seurat[ , which(expr == idents[[2]])]
  sample4_seurat <- subset(seurat, subset = external_donor_name_id == 528)
  expr <- FetchData(sample4_seurat, "seurat_clusters")
  sample4_seurat <- sample4_seurat[ , which(expr == idents[[2]])]
  sample5_seurat <- subset(seurat, subset = external_donor_name_id == 527)
  expr <- FetchData(sample5_seurat, "seurat_clusters")
  sample5_seurat <- sample5_seurat[ , which(expr == idents[[3]])]
  sample6_seurat <- subset(seurat, subset = external_donor_name_id == 528)
  expr <- FetchData(sample6_seurat, "seurat_clusters")
  sample6_seurat <- sample6_seurat[ , which(expr == idents[[3]])]
  sample7_seurat <- subset(seurat, subset = external_donor_name_id == 456)
  expr <- FetchData(sample7_seurat, "seurat_clusters")
  sample7_seurat <- sample7_seurat[ , which(expr == idents[[4]])]
  sample8_seurat <- subset(seurat, subset = external_donor_name_id == 471)
  expr <- FetchData(sample8_seurat, "seurat_clusters")
  sample8_seurat <- sample8_seurat[ , which(expr == idents[[4]])]

  downsample <- min(
    dim(sample1_seurat)[2], dim(sample2_seurat)[2], dim(sample3_seurat)[2],
    dim(sample4_seurat)[2], dim(sample5_seurat)[2], dim(sample6_seurat)[2],
    dim(sample7_seurat)[2], dim(sample8_seurat)[2]
  )

  sample1_seurat <- subset(seurat, subset = external_donor_name_id == 475)
  expr <- FetchData(sample1_seurat, "seurat_clusters")
  sample1_seurat <- sample1_seurat[ , which(expr == idents[[1]])]
  sample2_seurat <- subset(seurat, subset = external_donor_name_id == 476)
  expr <- FetchData(sample2_seurat, "seurat_clusters")
  sample2_seurat <- sample2_seurat[ , which(expr == idents[[1]])]
  # downsample <- min(dim(sample1_seurat)[2], dim(sample2_seurat)[2])
  set.seed(1)
  sample1_seurat <- subset(sample1_seurat, cells = sample(Cells(sample1_seurat), downsample))
  set.seed(1)
  sample2_seurat <- subset(sample2_seurat, cells = sample(Cells(sample2_seurat), downsample))
  condition1_seurat <- merge(sample1_seurat, sample2_seurat)

  sample1_seurat <- subset(seurat, subset = external_donor_name_id == 527)
  expr <- FetchData(sample1_seurat, "seurat_clusters")
  sample1_seurat <- sample1_seurat[ , which(expr == idents[[2]])]
  sample2_seurat <- subset(seurat, subset = external_donor_name_id == 528)
  expr <- FetchData(sample2_seurat, "seurat_clusters")
  sample2_seurat <- sample2_seurat[ , which(expr == idents[[2]])]
  # downsample <- min(dim(sample1_seurat)[2], dim(sample2_seurat)[2])
  set.seed(1)
  sample1_seurat <- subset(sample1_seurat, cells = sample(Cells(sample1_seurat), downsample))
  set.seed(1)
  sample2_seurat <- subset(sample2_seurat, cells = sample(Cells(sample2_seurat), downsample))
  condition2_seurat <- merge(sample1_seurat, sample2_seurat)

  sample1_seurat <- subset(seurat, subset = external_donor_name_id == 527)
  expr <- FetchData(sample1_seurat, "seurat_clusters")
  sample1_seurat <- sample1_seurat[ , which(expr == idents[[3]])]
  sample2_seurat <- subset(seurat, subset = external_donor_name_id == 528)
  expr <- FetchData(sample2_seurat, "seurat_clusters")
  sample2_seurat <- sample2_seurat[ , which(expr == idents[[3]])]
  # downsample <- min(dim(sample1_seurat)[2], dim(sample2_seurat)[2], dim(sample3_seurat)[2])
  set.seed(1)
  sample1_seurat <- subset(sample1_seurat, cells = sample(Cells(sample1_seurat), downsample))
  set.seed(1)
  sample2_seurat <- subset(sample2_seurat, cells = sample(Cells(sample2_seurat), downsample))
  condition3_seurat <- merge(sample1_seurat, sample2_seurat)

  sample1_seurat <- subset(seurat, subset = external_donor_name_id == 456)
  expr <- FetchData(sample1_seurat, "seurat_clusters")
  sample1_seurat <- sample1_seurat[ , which(expr == idents[[4]])]
  sample2_seurat <- subset(seurat, subset = external_donor_name_id == 471)
  expr <- FetchData(sample2_seurat, "seurat_clusters")
  sample2_seurat <- sample2_seurat[ , which(expr == idents[[4]])]
  # downsample <- min(dim(sample1_seurat)[2], dim(sample2_seurat)[2], dim(sample3_seurat)[2])
  set.seed(1)
  sample1_seurat <- subset(sample1_seurat, cells = sample(Cells(sample1_seurat), downsample))
  set.seed(1)
  sample2_seurat <- subset(sample2_seurat, cells = sample(Cells(sample2_seurat), downsample))
  condition4_seurat <- merge(sample1_seurat, sample2_seurat)

  # downsample <- min(
  #   dim(condition1_seurat)[2], dim(condition2_seurat)[2], dim(condition3_seurat)[2],
  #   dim(condition4_seurat)[2], dim(condition5_seurat)[2]
  # )
  # set.seed(1)
  # condition1_seurat <- subset(
  #   condition1_seurat, cells = sample(Cells(condition1_seurat), downsample)
  # )
  # set.seed(1)
  # condition2_seurat <- subset(
  #   condition2_seurat, cells = sample(Cells(condition2_seurat), downsample)
  # )
  # set.seed(1)
  # condition3_seurat <- subset(
  #   condition3_seurat, cells = sample(Cells(condition3_seurat), downsample)
  # )

  seurat <- merge(condition1_seurat, c(condition2_seurat, condition3_seurat, condition4_seurat))
  rm(
    sample1_seurat, sample2_seurat, sample3_seurat, sample4_seurat,
    sample5_seurat, sample6_seurat, sample7_seurat, sample8_seurat
  )
  seurat <- seurat[rowSums(as.matrix(GetAssayData(seurat, slot = "counts")) > 0) >= 10, ]
  # ---------------

  # Get gene annotations.
  # ---------------------
  mart <- useEnsembl("ensembl", paste0(organism, "_gene_ensembl"))
  attributes <- c("external_gene_name", "ensembl_gene_id", "chromosome_name")
  gene_anno <- getBM(attributes, "external_gene_name", rownames(seurat), mart)
  gene_anno <- gene_anno[gene_anno$chromosome_name %in% c(1:22), ]
  gene_anno <- gene_anno[gene_anno$ensembl_gene_id %in% unique(unlist(geneIds(gene_sets))), ]

  tmp <- lapply(
    geneIds(gene_sets), function(x, y) na.omit(fastmatch::fmatch(x, y)), gene_anno$ensembl_gene_id
  )
  tmp <- filterGeneSets(tmp, 3, 500)
  gene_sets <- gene_sets[names(gene_sets) %in% names(tmp)]

  gene_anno <- gene_anno[gene_anno$ensembl_gene_id %in% unique(unlist(geneIds(gene_sets))), ]
  # ---------------------

  # For gene symbols with multiple ENSEMBL IDs, duplicate the gene symbol to have an identical row
  # for each ENSEMBL ID.
  # ----------------------------------------------------------------------------------------------
  dup <- gene_anno[duplicated(gene_anno$external_gene_name), ]
  if (nrow(dup) > 0) {
    for (i in 1:dim(dup)[1]) {
      for (j in 1:dim(gene_anno)[1]) {
        if (dup$ensembl_gene_id[i] == gene_anno$ensembl_gene_id[j]) {
          gene_anno$external_gene_name[j] <- paste0(gene_anno$external_gene_name[j], "-alt")
        }
      }
    }
    if (any(duplicated(gene_anno$external_gene_name))) {
      stop("Duplicates in gene_anno.")
    }
    seurat <- seurat[rownames(seurat) %in% gene_anno$external_gene_name, ]
    new_mat <- GetAssayData(seurat, "counts")
    for (i in 1:dim(dup)[1]) {
      for (j in 1:dim(seurat)[1]) {
        if (dup$external_gene_name[i] == rownames(seurat)[j]) {
          new_row <- GetAssayData(seurat[j, ], "counts")
          rownames(new_row) <- paste0(rownames(new_row), "-alt")
          if (rownames(new_row) %in% rownames(new_mat)) {
            rownames(new_row) <- paste0(rownames(new_row), "2")
          }
          new_mat <- rbind(new_mat, new_row)
        }
      }
    }
  } else {
    seurat <- seurat[rownames(seurat) %in% gene_anno$external_gene_name, ]
    new_mat <- GetAssayData(seurat, "counts")
  }
  gene_anno <- gene_anno[gene_anno$external_gene_name %in% rownames(new_mat), ]
  gene_anno <- gene_anno[order(match(gene_anno$external_gene_name, rownames(new_mat))), ]
  rownames(new_mat) <- gene_anno$ensembl_gene_id

  seurat <- CreateSeuratObject(new_mat, meta.data = seurat[[]])
  rm(new_mat)
  seurat$idents <- factor(
    c(
      rep(names(idents)[1], dim(condition1_seurat)[2]),
      rep(names(idents)[2], dim(condition2_seurat)[2]),
      rep(names(idents)[3], dim(condition3_seurat)[2]),
      rep(names(idents)[4], dim(condition4_seurat)[2])
    ),
    levels = names(idents)
  )
  rm(condition1_seurat, condition2_seurat, condition3_seurat, condition4_seurat)
  seurat@active.ident <- seurat$idents
  seurat$groups <- paste0(seurat$idents, "_", seurat$external_donor_name_id)
  # ----------------------------------------------------------------------------------------------
  
  saveRDS(seurat, rds)
  saveRDS(gene_anno, rds2)
}
# ------------------------------------------------------------------------------

# Renormalize using SCnorm using sample-wise condition information and gene-wise gene length and GC
# content information.
# -------------------------------------------------------------------------------------------------
rds <- file.path(cache_dir, paste0(gene_set_name, "_scnorm.rds"))
rds2 <- file.path(cache_dir, paste0(gene_set_name, "_row_data.rds"))
if (file.exists(rds)) {
  sce <- readRDS(rds)
} else {
  sce <- SingleCellExperiment(list(counts = as.matrix(GetAssayData(seurat))))
  row_data <- data.frame(
    getGeneLengthAndGCContent(rownames(sce), "mmusculus_gene_ensembl", "biomart")
  )
  saveRDS(row_data, rds2)
  # row_data <- readRDS(rds2)
  # sce <- SCnorm(
  #   sce, seurat$idents, withinSample = c(row_data$gc, row_data$length),
  #   useZerosToScale = TRUE, ditherCounts = TRUE, NCores = detectCores()
  # )
  sce <- SCnorm(
    sce, rep(1, dim(sce)[2]), withinSample = c(row_data$gc, row_data$length),
    useZerosToScale = TRUE, ditherCounts = TRUE, NCores = detectCores()
  )
  saveRDS(sce, rds)
}
# -------------------------------------------------------------------------------------------------

# Add normalized data to Seurat object.
# -------------------------------------
seurat <- SetAssayData(seurat, new.data = normcounts(sce))
seurat <- SetAssayData(seurat, "scale.data", log2(normcounts(sce) + 1))
# -------------------------------------

# Create MDS plots to identify potential confounders and latent factors.
# ----------------------------------------------------------------------
rds <- file.path(cache_dir, paste0(gene_set_name, "_d.rds"))
d <- dist(scale(t(GetAssayData(seurat, "scale.data"))))
mds <- cmdscale(d, 2)
colnames(mds) <- paste0("mds_", 1:2)
seurat[["mds"]] <- CreateDimReducObject(mds, key = "mds_", assay = DefaultAssay(seurat))
rm(mds)
add_df <- data.frame(Embeddings(seurat, "mds"))
names(add_df) <- paste0("mds", seq(ncol(add_df)))
seurat$mds1 <- add_df$mds1
seurat$mds2 <- add_df$mds2

red_dim_plot(seurat, "umap1", "umap2", "idents", "cat")
red_dim_plot(seurat, "mds1", "mds2", "idents", "cat")
for (i in 1:length(metadata_to_plot)) {
  print(names(metadata_to_plot)[i])
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "mds1", "mds2", names(metadata_to_plot)[i]) + NoLegend())
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "mds1", "mds2", names(metadata_to_plot)[i], "cat") + NoLegend())
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(red_dim_plot(seurat, "mds1", "mds2", names(metadata_to_plot)[i]))
  } else {
    print(red_dim_plot(seurat, "mds1", "mds2", names(metadata_to_plot)[i], "cat"))
  }
}
# ----------------------------------------------------------------------

print("SCnorm normalized counts distribution")
hist(as.matrix(GetAssayData(seurat, "scale.data")))

# Aggregate features to into "pseudo-bulk" counts.
# ------------------------------------------------
rds <- file.path(cache_dir, paste0(gene_set_name, "_pseudobulk.rds"))
if (file.exists(rds)) {
  sce <- readRDS(rds)
} else {
  sce <- as.SingleCellExperiment(seurat)
  counts(sce) <- logcounts(sce)
  sce <- suppressWarnings(
    aggregateAcrossCells(
      sce, colData(sce)[ , c("idents", "external_donor_name_id")],
      use_exprs_values = "counts", BPPARAM = MulticoreParam()
    )
  )
  keep <- filterByExpr(counts(sce), group = sce$idents, min.total.count = 10)
  sce <- sce[keep, ]
  dge <- DGEList(counts(sce))
  dge <- calcNormFactors(dge)
  # dge <- voom(dge, design, plot = TRUE)
  logcounts(sce) <- cpm(dge, log = TRUE, prior.count = 3)
  rm(dge)
  # logcounts(sce) <- log2(counts(sce) + 1)
  # sce <- logNormCounts(sce)
  # logcounts(sce) <- counts(sce)
  saveRDS(sce, rds)
}
print("Pseudo-bulk counts distribution")
hist(logcounts(sce))
# ------------------------------------------------

# Subset other objects.
# ---------------------
gene_anno <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(sce), ]
seurat <- seurat[rownames(seurat) %in% rownames(sce), ]

tmp <- lapply(
  geneIds(gene_sets), function(x, y) na.omit(fastmatch::fmatch(x, y)), gene_anno$ensembl_gene_id
)
tmp <- filterGeneSets(tmp, 3, 500)
gene_sets <- gene_sets[names(gene_sets) %in% names(tmp)]

rds <- file.path(cache_dir, paste0(gene_set_name, "_gene_set_overlap.rds"))
if (file.exists(rds)) {
  overlap <- readRDS(rds)
} else {
  overlap <- computeGeneSetsOverlapMax(gene_sets, unique(unlist(geneIds(gene_sets))))
  saveRDS(overlap, rds)
}
tmp <- rowSums(overlap)
tmp <- tmp[order(tmp, decreasing = TRUE)]
gene_sets_sorted <- gene_sets[match(names(tmp), names(gene_sets))]
rds <- file.path(cache_dir, paste0(gene_set_name, "_gene_set_overlap_sorted.rds"))
if (file.exists(rds)) {
  overlap <- readRDS(rds)
} else {
  overlap <- computeGeneSetsOverlapMax(gene_sets_sorted, unique(unlist(geneIds(gene_sets_sorted))))
  saveRDS(overlap, rds)
}

overlap[upper.tri(overlap)] <- 0
diag(overlap) <- 0
keep <- apply(overlap, 1, max)
keep <- keep[keep < 1]
gene_sets <- gene_sets[names(gene_sets) %in% names(keep)]

sce <- sce[rownames(sce) %in% unique(unlist(geneIds(gene_sets))), ]
gene_anno <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(sce), ]
seurat <- seurat[rownames(seurat) %in% rownames(sce), ]
tmp <- lapply(
  geneIds(gene_sets), function(x, y) na.omit(fastmatch::fmatch(x, y)), gene_anno$ensembl_gene_id
)
tmp <- filterGeneSets(tmp, 3, 500)
gene_sets <- gene_sets[names(gene_sets) %in% names(tmp)]
#----------------------

# # Run GSVA on all cells.
# # ----------------------
# rds <- file.path(cache_dir, paste0(gene_set_name, "_gsva_allcells.rds"))
# if (file.exists(rds)) {
#   gsva <- readRDS(rds)
# } else {
#   gsva <- gsva(as.matrix(GetAssayData(seurat, "scale.data")), gene_sets)
#   saveRDS(gsva, rds)
# }
# seurat[["GSVA"]] <- CreateAssayObject(gsva)
# DefaultAssay(seurat) <- "GSVA"
# print("GSVA distribution (all cells)")
# hist(as.matrix(GetAssayData(seurat, "counts")))
# # ----------------------

# Run GSVA on pseudo-bulk counts.
# -------------------------------
rds <- file.path(cache_dir, paste0(gene_set_name, "_gsva_pseudobulk.rds"))
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(logcounts(sce), gene_sets, method = "ssgsea", ssgsea.norm = FALSE)
  colnames(gsva) <- sce$groups
  saveRDS(gsva, rds)
}
# -------------------------------

# Add previous UMAP to Seurat object.
# -----------------------------------
umap <- as.matrix(data.frame(seurat$umap1, seurat$umap2))
colnames(umap) <- paste0("umap_", 1:2)
seurat[["umap"]] <- CreateDimReducObject(umap, key = "umap_", assay = DefaultAssay(seurat))
rm(umap)
# -----------------------------------
```

## DE `r gene_set_name` Ontology Tables

```{r}
# Fit model on pseudo-bulk GSVA results.
# --------------------------------------
design <- model.matrix(~ 0 + sce$idents)
colnames(design) <- names(idents)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(
  EC_Reln-IN_Sst, CA1_Wfs1-IN_Sst, CA3_Bok-IN_Sst,
  EC_Reln-CA1_Wfs1, CA1_Wfs1-CA3_Bok, levels = design
)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat))
tests <- decideTests(cont_fit, "nestedF", p.value = 0.05)
tests_up <- tests[tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1, ]
discard <- which(rownames(tests_up) %in% rownames(tests_up[rowSums(tests_up == -1) > 0, ]))
if (length(discard) > 0) {
  tests_up <- tests_up[-discard, ]
}
tests_down <- tests[tests[ , 1] == -1 & tests[ , 2] == -1 & tests[ , 3] == -1, ]
discard <- which(rownames(tests_down) %in% rownames(tests_down[rowSums(tests_down == 1) > 0, ]))
if (length(discard) > 0) {
  tests_down <- tests_down[-discard, ]
}
results <- topTable(cont_fit, number = 100)
# --------------------------------------

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}
gsva_list_pseudobulk <- list(all, up, down)
names(gsva_list_pseudobulk) <- c(
  "All DE", paste0("Upregulated in ", names(idents)[1]),
  paste0("Downregulated in ", names(idents)[2])
)

# Fit a strict model on pseudo-bulk GSVA results.
# -----------------------------------------------
design <- model.matrix(~ 0 + sce$idents)
colnames(design) <- names(idents)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(
  EC_Reln-IN_Sst, EC_Reln-CA3_Bok, EC_Reln-CA1_Wfs1,
  CA1_Wfs1-CA3_Bok, CA3_Bok-IN_Sst, levels = design
)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat))
tests <- decideTests(cont_fit, "nestedF", p.value = 0.05)
tests_up <- tests[tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1, ]
discard <- which(rownames(tests_up) %in% rownames(tests_up[rowSums(tests_up == -1) > 0, ]))
if (length(discard) > 0) {
  tests_up <- tests_up[-discard, ]
}
tests_down <- tests[tests[ , 1] == -1 & tests[ , 2] == -1 & tests[ , 3] == -1, ]
discard <- which(rownames(tests_down) %in% rownames(tests_down[rowSums(tests_down == 1) > 0, ]))
if (length(discard) > 0) {
  tests_down <- tests_down[-discard, ]
}
results <- topTable(cont_fit, number = 200)
# -----------------------------------------------

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}
gsva_list_pseudobulk_strict <- list(all, up, down)
names(gsva_list_pseudobulk_strict) <- c(
  "All DE", paste0("Upregulated in ", names(idents)[1]),
  paste0("Downregulated in ", names(idents)[2])
)

# Fit a stricter model on pseudo-bulk GSVA results.
# -------------------------------------------------
design <- model.matrix(~ 0 + sce$idents)
colnames(design) <- names(idents)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(
  EC_Reln-IN_Sst, EC_Reln-CA3_Bok, CA1_Wfs1-IN_Sst, CA1_Wfs1-CA3_Bok,
  EC_Reln-CA1_Wfs1, CA3_Bok-IN_Sst, levels = design
)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat))
tests <- decideTests(cont_fit, "nestedF", p.value = 0.05)
tests_up <- tests[tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1 & tests[ , 4] == 1, ]
discard <- which(rownames(tests_up) %in% rownames(tests_up[rowSums(tests_up == -1) > 0, ]))
if (length(discard) > 0) {
  tests_up <- tests_up[-discard, ]
}
tests_down <- tests[tests[ , 1] == -1 & tests[ , 2] == -1 & tests[ , 3] == -1 & tests[ , 4] == -1, ]
discard <- which(rownames(tests_down) %in% rownames(tests_down[rowSums(tests_down == 1) > 0, ]))
if (length(discard) > 0) {
  tests_down <- tests_down[-discard, ]
}
results <- topTable(cont_fit, number = 300)
# -------------------------------------------------

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}
gsva_list_pseudobulk_stricter <- list(all, up, down)
names(gsva_list_pseudobulk_stricter) <- c(
  "All DE", paste0("Upregulated in ", names(idents)[1]),
  paste0("Downregulated in ", names(idents)[2])
)

# Fit the strictest model on pseudo-bulk GSVA results.
# ----------------------------------------------------
design <- model.matrix(~ 0 + sce$idents)
colnames(design) <- names(idents)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(EC_Reln-CA1_Wfs1, CA1_Wfs1-CA3_Bok, CA3_Bok-IN_Sst, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat))
tests <- decideTests(cont_fit, "nestedF", p.value = 0.05)
tests_up <- tests[tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1, ]
discard <- which(rownames(tests_up) %in% rownames(tests_up[rowSums(tests_up == -1) > 0, ]))
if (length(discard) > 0) {
  tests_up <- tests_up[-discard, ]
}
tests_down <- tests[tests[ , 1] == -1 & tests[ , 2] == -1 & tests[ , 3] == -1, ]
discard <- which(rownames(tests_down) %in% rownames(tests_down[rowSums(tests_down == 1) > 0, ]))
if (length(discard) > 0) {
  tests_down <- tests_down[-discard, ]
}
results <- topTable(cont_fit, number = 400)
# ----------------------------------------------------

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}
gsva_list_pseudobulk_strictest <- list(all, up, down)
names(gsva_list_pseudobulk_strictest) <- c(
  "All DE", paste0("Upregulated in ", names(idents)[1]),
  paste0("Downregulated in ", names(idents)[2])
)
```

**`r names(idents)[1]` Upregulated `r gene_set_name` Pseudo-bulk**

```{r}
ncol <- ncol(gsva_list_pseudobulk[[2]])
datatable_download_exp(gsva_list_pseudobulk[[2]][(ncol-3):ncol])
```

**`r names(idents)[1]` Upregulated `r gene_set_name` Pseudo-bulk Strict Model**

```{r}
ncol <- ncol(gsva_list_pseudobulk_strict[[2]])
datatable_download_exp(gsva_list_pseudobulk_strict[[2]][(ncol-3):ncol])
```

**`r names(idents)[1]` Upregulated `r gene_set_name` Pseudo-bulk Stricter Model**

```{r}
ncol <- ncol(gsva_list_pseudobulk_stricter[[2]])
datatable_download_exp(gsva_list_pseudobulk_stricter[[2]][(ncol-3):ncol])
```

**`r names(idents)[1]` Upregulated `r gene_set_name` Pseudo-bulk Strictest Model**

```{r}
ncol <- ncol(gsva_list_pseudobulk_strictest[[2]])
datatable_download_exp(gsva_list_pseudobulk_strictest[[2]][(ncol-3):ncol])
```

**`r names(idents)[1]` Downregulated `r gene_set_name` Pseudo-bulk**

```{r}
ncol <- ncol(gsva_list_pseudobulk[[3]])
datatable_download_exp(gsva_list_pseudobulk[[3]][(ncol-3):ncol])
```

**`r names(idents)[1]` Downregulated `r gene_set_name` Pseudo-bulk Strict**

```{r}
ncol <- ncol(gsva_list_pseudobulk_strict[[3]])
datatable_download_exp(gsva_list_pseudobulk_strict[[3]][(ncol-3):ncol])
```

**`r names(idents)[1]` Downregulated `r gene_set_name` Pseudo-bulk Stricter**

```{r}
ncol <- ncol(gsva_list_pseudobulk_stricter[[3]])
datatable_download_exp(gsva_list_pseudobulk_stricter[[3]][(ncol-3):ncol])
```

**`r names(idents)[1]` Downregulated `r gene_set_name` Pseudo-bulk Strict**

```{r}
ncol <- ncol(gsva_list_pseudobulk_strictest[[3]])
datatable_download_exp(gsva_list_pseudobulk_strictest[[3]][(ncol-3):ncol])
```

## DE Genes Tables

```{r}
# Fit model on pseudo-bulk counts.
# --------------------------------
design <- model.matrix(~ 0 + sce$idents)
colnames(design) <- names(idents)
fit <- lmFit(logcounts(sce), design)
contrast_mat <- makeContrasts(
  EC_Reln-IN_Sst, CA1_Wfs1-IN_Sst, CA3_Bok-IN_Sst,
  EC_Reln-CA1_Wfs1, CA1_Wfs1-CA3_Bok, levels = design
)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), robust = TRUE, trend = TRUE)
tests <- decideTests(cont_fit, "nestedF", p.value = 0.001)
tests_up <- tests[tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1, ]
discard <- which(rownames(tests_up) %in% rownames(tests_up[rowSums(tests_up == -1) > 0, ]))
if (length(discard) > 0) {
  tests_up <- tests_up[-discard, ]
}
tests_down <- tests[tests[ , 1] == -1 & tests[ , 2] == -1 & tests[ , 3] == -1, ]
discard <- which(rownames(tests_down) %in% rownames(tests_down[rowSums(tests_down == 1) > 0, ]))
if (length(discard) > 0) {
  tests_down <- tests_down[-discard, ]
}
results <- topTable(cont_fit, number = Inf)
# --------------------------------------

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}
markers_list_pseudobulk <- list(all, up, down)

# Fit a strict model on pseudo-bulk counts.
# -----------------------------------------
design <- model.matrix(~ 0 + sce$idents)
colnames(design) <- names(idents)
fit <- lmFit(logcounts(sce), design)
contrast_mat <- makeContrasts(
  EC_Reln-IN_Sst, EC_Reln-CA3_Bok, EC_Reln-CA1_Wfs1,
  CA1_Wfs1-CA3_Bok, CA3_Bok-IN_Sst, levels = design
)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), robust = TRUE, trend = TRUE)
tests <- decideTests(cont_fit, "nestedF", p.value = 0.005)
tests_up <- tests[tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1, ]
discard <- which(rownames(tests_up) %in% rownames(tests_up[rowSums(tests_up == -1) > 0, ]))
if (length(discard) > 0) {
  tests_up <- tests_up[-discard, ]
}
tests_down <- tests[tests[ , 1] == -1 & tests[ , 2] == -1 & tests[ , 3] == -1, ]
discard <- which(rownames(tests_down) %in% rownames(tests_down[rowSums(tests_down == 1) > 0, ]))
if (length(discard) > 0) {
  tests_down <- tests_down[-discard, ]
}
results <- topTable(cont_fit, number = Inf)
# -----------------------------------------

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}
markers_list_pseudobulk_strict <- list(all, up, down)

# Fit a stricter model on pseudo-bulk counts.
# -------------------------------------------
design <- model.matrix(~ 0 + sce$idents)
colnames(design) <- names(idents)
fit <- lmFit(logcounts(sce), design)
contrast_mat <- makeContrasts(
  EC_Reln-IN_Sst, EC_Reln-CA3_Bok, CA1_Wfs1-IN_Sst, CA1_Wfs1-CA3_Bok,
  EC_Reln-CA1_Wfs1, CA3_Bok-IN_Sst, levels = design
)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), robust = TRUE, trend = TRUE)
tests <- decideTests(cont_fit, "nestedF", p.value = 0.01)
tests_up <- tests[tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1 & tests[ , 4] == 1, ]
discard <- which(rownames(tests_up) %in% rownames(tests_up[rowSums(tests_up == -1) > 0, ]))
if (length(discard) > 0) {
  tests_up <- tests_up[-discard, ]
}
tests_down <- tests[tests[ , 1] == -1 & tests[ , 2] == -1 & tests[ , 3] == -1 & tests[ , 4] == -1, ]
discard <- which(rownames(tests_down) %in% rownames(tests_down[rowSums(tests_down == 1) > 0, ]))
if (length(discard) > 0) {
  tests_down <- tests_down[-discard, ]
}
results <- topTable(cont_fit, number = Inf)
# -------------------------------------------

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}
markers_list_pseudobulk_stricter <- list(all, up, down)

# Fit the strictest model on pseudo-bulk counts.
# ----------------------------------------------
design <- model.matrix(~ 0 + sce$idents)
colnames(design) <- names(idents)
fit <- lmFit(logcounts(sce), design)
contrast_mat <- makeContrasts(EC_Reln-CA1_Wfs1, CA1_Wfs1-CA3_Bok, CA3_Bok-IN_Sst, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), robust = TRUE, trend = TRUE)
tests <- decideTests(cont_fit, "nestedF", p.value = 0.05)
tests_up <- tests[tests[ , 1] == 1 & tests[ , 2] == 1 & tests[ , 3] == 1, ]
discard <- which(rownames(tests_up) %in% rownames(tests_up[rowSums(tests_up == -1) > 0, ]))
if (length(discard) > 0) {
  tests_up <- tests_up[-discard, ]
}
tests_down <- tests[tests[ , 1] == -1 & tests[ , 2] == -1 & tests[ , 3] == -1, ]
discard <- which(rownames(tests_down) %in% rownames(tests_down[rowSums(tests_down == 1) > 0, ]))
if (length(discard) > 0) {
  tests_down <- tests_down[-discard, ]
}
results <- topTable(cont_fit, number = Inf)
# ----------------------------------------------

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$F), ]
}
markers_list_pseudobulk_strictest <- list(all, up, down)

# Convert ENSEMBL IDs back to gene symbols.
# -----------------------------------------
genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk[[2]]), ]
genes <- genes[order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk[[2]]))), ]
up <- markers_list_pseudobulk[[2]]
rownames(up) <- genes$external_gene_name

genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk[[3]]), ]
genes <- genes[order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk[[3]]))), ]
down <- markers_list_pseudobulk[[3]]
rownames(down) <- genes$external_gene_name

all <- rbind(up, down)
all <- all[order(-all$F), ]
markers_list_symbols_pseudobulk <- list(all, up, down)

genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk_strict[[2]]), ]
genes <- genes[order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk_strict[[2]]))), ]
up <- markers_list_pseudobulk_strict[[2]]
rownames(up) <- genes$external_gene_name

genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk_strict[[3]]), ]
genes <- genes[order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk_strict[[3]]))), ]
down <- markers_list_pseudobulk_strict[[3]]
rownames(down) <- genes$external_gene_name

all <- rbind(up, down)
all <- all[order(-all$F), ]
markers_list_symbols_pseudobulk_strict <- list(all, up, down)

genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk_stricter[[2]]), ]
genes <- genes[
  order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk_stricter[[2]]))),
]
up <- markers_list_pseudobulk_stricter[[2]]
rownames(up) <- genes$external_gene_name

genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk_stricter[[3]]), ]
genes <- genes[
  order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk_stricter[[3]]))),
]
down <- markers_list_pseudobulk_stricter[[3]]
rownames(down) <- genes$external_gene_name

all <- rbind(up, down)
all <- all[order(-all$F), ]
markers_list_symbols_pseudobulk_stricter <- list(all, up, down)

genes <- gene_anno[
  gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk_strictest[[2]]),
]
genes <- genes[
  order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk_strictest[[2]]))),
]
up <- markers_list_pseudobulk_strictest[[2]]
rownames(up) <- genes$external_gene_name

genes <- gene_anno[
  gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk_strictest[[3]]),
]
genes <- genes[
  order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk_strictest[[3]]))),
]
down <- markers_list_pseudobulk_strictest[[3]]
rownames(down) <- genes$external_gene_name

all <- rbind(up, down)
all <- all[order(-all$F), ]
markers_list_symbols_pseudobulk_strictest <- list(all, up, down)
# -----------------------------------------

# Create new Seurat object with gene symbols.
# -------------------------------------------
new_mat <- GetAssayData(seurat)
rownames(new_mat) <- gene_anno$external_gene_name
seurat_symbols <- CreateSeuratObject(new_mat, meta.data = seurat[[]])
Idents(seurat_symbols) <- "idents"
seurat_symbols <- SetAssayData(
  seurat_symbols, new.data = log2(as.matrix(GetAssayData(seurat_symbols, "counts")) + 1)
)
seurat_symbols <- SetAssayData(
  seurat_symbols, "scale.data", t(scale(t(GetAssayData(seurat_symbols)), FALSE))
)
# DefaultAssay(seurat) <- "GSVA"
# -------------------------------------------

# color2 <- colorRampPalette(brewer.pal(9, "YlOrBr"))(round(max(GetAssayData(seurat_symbols))))
# color2[1] <- "#DDDDDD"
```

**`r names(idents)[1]` Upregulated Pseudo-bulk**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk[[2]])
datatable_download_exp(markers_list_symbols_pseudobulk[[2]][(ncol-3):ncol])
```

**`r names(idents)[1]` Upregulated Pseudo-bulk Strict Model**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk_strict[[2]])
datatable_download_exp(markers_list_symbols_pseudobulk_strict[[2]][(ncol-3):ncol])
```

**`r names(idents)[1]` Upregulated Pseudo-bulk Stricter Model**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk_stricter[[2]])
datatable_download_exp(markers_list_symbols_pseudobulk_stricter[[2]][(ncol-3):ncol])
```

**`r names(idents)[1]` Upregulated Pseudo-bulk Strictest Model**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk_strictest[[2]])
datatable_download_exp(markers_list_symbols_pseudobulk_strictest[[2]][(ncol-3):ncol])
```

**`r names(idents)[1]` Downregulated Pseudo-bulk**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk[[3]])
datatable_download_exp(markers_list_symbols_pseudobulk[[3]][(ncol-3):ncol])
```

**`r names(idents)[1]` Downregulated Pseudo-bulk Strict Model**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk_strict[[3]])
datatable_download_exp(markers_list_symbols_pseudobulk_strict[[3]][(ncol-3):ncol])
```

**`r names(idents)[1]` Downregulated Pseudo-bulk Stricter Model**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk_stricter[[3]])
datatable_download_exp(markers_list_symbols_pseudobulk_stricter[[3]][(ncol-3):ncol])
```

**`r names(idents)[1]` Downregulated Pseudo-bulk Strictest Model**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk_strictest[[3]])
datatable_download_exp(markers_list_symbols_pseudobulk_strictest[[3]][(ncol-3):ncol])
```

## DE GO Ontology Heatmaps

### `r gene_set_name` Ontologies

```{r, fig.width = 20, fig.height = 40}
anno_color <- hue_pal()(length(unique(sce$external_donor_name_id)))
gsva_heatmap <- vector("list", 3)

color2 <- colorRamp2(
  c(min(gsva), (max(gsva) + min(gsva)) / 2, max(gsva)),
  c(brewer.pal(3, "RdBu")[3], brewer.pal(3, "RdBu")[2], brewer.pal(3, "RdBu")[1])
)

gsva_list_pseudobulk_all <- list(
  c(
    rownames(gsva_list_pseudobulk[[1]]), rownames(gsva_list_pseudobulk_strict[[1]]),
    rownames(gsva_list_pseudobulk_stricter[[1]]), rownames(gsva_list_pseudobulk_strictest[[1]])
  ),
  c(
    rownames(gsva_list_pseudobulk[[2]]), rownames(gsva_list_pseudobulk_strict[[2]]),
    rownames(gsva_list_pseudobulk_stricter[[2]]), rownames(gsva_list_pseudobulk_strictest[[2]])
  ),
  c(
    rownames(gsva_list_pseudobulk[[3]]), rownames(gsva_list_pseudobulk_strict[[3]]),
    rownames(gsva_list_pseudobulk_stricter[[3]]), rownames(gsva_list_pseudobulk_strictest[[3]])
  )
)
markers_list_pseudobulk_all <- list(
  c(
    rownames(markers_list_pseudobulk[[1]]),
    rownames(markers_list_pseudobulk_strict[[1]]),
    rownames(markers_list_pseudobulk_stricter[[1]]),
    rownames(markers_list_pseudobulk_strictest[[1]])
  ),
  c(
    rownames(markers_list_pseudobulk[[2]]),
    rownames(markers_list_pseudobulk_strict[[2]]),
    rownames(markers_list_pseudobulk_stricter[[2]]),
    rownames(markers_list_pseudobulk_strictest[[2]])
  ),
  c(
    rownames(markers_list_pseudobulk[[3]]),
    rownames(markers_list_pseudobulk_strict[[3]]),
    rownames(markers_list_pseudobulk_stricter[[3]]),
    rownames(markers_list_pseudobulk_strictest[[3]])
  )
)
markers_list_symbols_pseudobulk_all <- list(
  c(
    rownames(markers_list_symbols_pseudobulk[[1]]),
    rownames(markers_list_symbols_pseudobulk_strict[[1]]),
    rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
    rownames(markers_list_symbols_pseudobulk_strictest[[1]])
  ),
  c(
    rownames(markers_list_symbols_pseudobulk[[2]]),
    rownames(markers_list_symbols_pseudobulk_strict[[2]]),
    rownames(markers_list_symbols_pseudobulk_stricter[[2]]),
    rownames(markers_list_symbols_pseudobulk_strictest[[2]])
  ),
  c(
    rownames(markers_list_symbols_pseudobulk[[3]]),
    rownames(markers_list_symbols_pseudobulk_strict[[3]]),
    rownames(markers_list_symbols_pseudobulk_stricter[[3]]),
    rownames(markers_list_symbols_pseudobulk_strictest[[3]])
  )
)

# if (organism == "hsapiens") {
#   go_data <- godata("org.Hs.eg.db", "ENSEMBL", unlist(strsplit(gene_set_name, "\\."))[2], FALSE)
# } else if (organism == "mmusculus") {
#   go_data <- godata("org.Mm.eg.db", "ENSEMBL", unlist(strsplit(gene_set_name, "\\."))[2], FALSE)
# }

# for (i in seq_along(gsva_list_pseudobulk)) {
for (i in 1) {
  if (length(gsva_list_pseudobulk_all[[i]]) > 0) {
      i <- 1
      mat <- gsva
      mat <- mat[rownames(mat) %in% gsva_list_pseudobulk_all[[1]], ]
      mat_up <- mat[rownames(mat) %in% gsva_list_pseudobulk_all[[2]], ]
      mat_down <- mat[rownames(mat) %in% gsva_list_pseudobulk_all[[3]], ]
      gene_sets_sub_up <- gene_sets[names(gene_sets) %in% rownames(mat_up)]
      up_vector <- vector()
      for (j in seq(length(gene_sets_sub_up@.Data))) {
        if (any(gene_sets_sub_up[[j]]@geneIds %in% markers_list_pseudobulk_all[[2]])) {
          up_vector <- append(up_vector, names(gene_sets_sub_up)[j])
        }
      }
      gene_sets_sub_down <- gene_sets[names(gene_sets) %in% rownames(mat_down)]
      down_vector <- c()
      for (j in seq(length(gene_sets_sub_down@.Data))) {
        if (any(gene_sets_sub_down[[j]]@geneIds %in% markers_list_pseudobulk_all[[3]])) {
          down_vector <- append(down_vector, names(gene_sets_sub_down)[j])
        }
      }
      all_vector <- append(up_vector, down_vector)
      mat <- mat[rownames(mat) %in% all_vector, ]
      gene_sets_sub <- gene_sets[names(gene_sets) %in% rownames(mat)]
      # gene_sets_sub <- gene_sets_sub[names(gene_sets_sub) %in% names(gene_sets_filtered)]
      # mat <- mat[rownames(mat) %in% names(gene_sets_sub), ]
      gene_sets_sub_ids <- gene_sets_sub
      for (j in seq(length(gene_sets_sub_ids@.Data))) {
        suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
      }
      # keep <- which(names(gene_sets_sub_ids) %in% go_data@geneAnno$GO)
      # gene_sets_sub_ids <- gene_sets_sub_ids[keep, ]
      # gene_sets_sub <- gene_sets_sub[keep, ]
      # mat <- mat[keep, ]
    if (is.null(nrow(mat)) || dim(mat)[1] == 0) {
    } else {
      # gene_set_overlap <- mgoSim(
      #   names(gene_sets_sub_ids), names(gene_sets_sub_ids), go_data, "Jiang", NULL
      # )
      # discard <- colnames(gene_set_overlap)[
      #   colSums(is.na(gene_set_overlap)) == ncol(gene_set_overlap)
      # ]
      # discard <- which(names(gene_sets_sub_ids) %in% discard)
      # mat <- mat[-discard, ]
      # gene_sets_sub <- gene_sets_sub[-discard, ]
      # gene_sets_sub_ids <- gene_sets_sub_ids[-discard, ]

      # discard <- grep("positive", rownames(mat))
      # mat <- mat[-discard, ]
      # gene_sets_sub <- gene_sets_sub[-discard, ]
      # gene_sets_sub_ids <- gene_sets_sub_ids[-discard, ]
      # discard <- grep("negative", rownames(mat))
      # mat <- mat[-discard, ]
      # gene_sets_sub <- gene_sets_sub[-discard, ]
      # gene_sets_sub_ids <- gene_sets_sub_ids[-discard, ]

      discard <- computeGeneSetsOverlapMax(
        gene_sets_sub_ids, unique(unlist(geneIds(gene_sets_sub_ids)))
      )
      discard[!upper.tri(discard)] <- 0
      discard <- colnames(discard)[colSums(discard) == 0]
      discard <- which(names(gene_sets_sub_ids) %in% discard)
      mat <- mat[-discard, ]
      gene_sets_sub <- gene_sets_sub[-discard, ]
      gene_sets_sub_ids <- gene_sets_sub_ids[-discard, ]

      # gene_set_overlap <- mgoSim(
      #   names(gene_sets_sub_ids), names(gene_sets_sub_ids), go_data, combine = NULL
      # )
      gene_set_overlap <- computeGeneSetsOverlapMax(
        gene_sets_sub_ids, unique(unlist(geneIds(gene_sets_sub_ids)))
      )
      clustering <- hclust(get_dist(gene_set_overlap, "spearman"), "ward.D2")

      # rds <- file.path(cache_dir, paste0(gene_set_name, "_nbclust.rds"))
      # if (file.exists(rds)) {
      #   nbclust <- readRDS(rds)
      # } else {
      #   set.seed(1)
      #   nbclust <- clusGap(gene_set_overlap, nbclust_fun, 25, d.power = 2, verbose = FALSE)
      #   saveRDS(nbclust, rds)
      # }
      # optimal_k <- c(25)
      # for (l in 1:25) {
      #   if (optimal_k == 25) {
      #     optimal_k <- maxSE(nbclust$Tab[ , "gap"], nbclust$Tab[ , "SE.sim"], "Tibs2001SEmax", l)
      #   }
      # }
      # optimal_k <- 15
      # module_colors <- rev(cols25(optimal_k))
      # clustering <- color_branches(clustering, optimal_k, col = module_colors)
      # plot(clustering)
      # heights_per_k <- heights_per_k.dendrogram(as.dendrogram(clustering))
      # nbclust <- fviz_nbclust(
      #   gene_set_overlap, nbclust_fun, k.max = 25, print.summary = FALSE, verbose = FALSE
      # )
      # fviz_gap_stat(nbclust, maxSE = list(method = "Tibs2001SEmax", SE.factor = 2))

      anno <- HeatmapAnnotation(
        donor = sce$external_donor_name_id,
        cell_type = anno_block(
          gp = gpar(fill = "lightgray", col = "lightgray"),
          labels = c(
            names(idents)[[1]], names(idents)[[2]], names(idents)[[3]], names(idents)[[4]]
          ),
          labels_gp = gpar(col = "black", fontsize = 20)
        ),
        col = list(
          donor = c(
            "475" = anno_color[1], "476" = anno_color[2],
            "527" = anno_color[3], "528" = anno_color[4],
            "456" = anno_color[5], "471" = anno_color[6]
          )
        )
      )
      if (nrow(mat) >= 8) {
        modules_up <- rownames(mat)[rownames(mat) %in% gsva_list_pseudobulk_all[[2]]]
        modules_down <- rownames(mat)[rownames(mat) %in% gsva_list_pseudobulk_all[[3]]]

        module_gene_sets_sub <- gene_sets[names(gene_sets) %in% modules_up]
        frequent_genes_tmp <- as.data.frame(table(unlist(geneIds(module_gene_sets_sub))))
        frequent_genes_tmp <- frequent_genes_tmp[
          frequent_genes_tmp$Var1 %in% gene_anno$ensembl_gene_id,
        ]
        rownames(frequent_genes_tmp) <- NULL
        # if (any(frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[1]])) {
        #   discard <- which(
        #     frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[1]] &
        #       frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[3]]
        #   )
        #   if (length(discard) > 0) {
        #     frequent_genes_tmp <- frequent_genes_tmp[-discard, ]
        #   }
        # }
        # frequent_genes_tmp <- frequent_genes_tmp[frequent_genes_tmp$Freq >= 2, ]
        # frequent_genes_tmp <- frequent_genes_tmp[
        #   frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[2]],
        # ]
        frequent_genes_up <- frequent_genes_tmp

        module_gene_sets_sub <- gene_sets[names(gene_sets) %in% modules_down]
        frequent_genes_tmp <- as.data.frame(table(unlist(geneIds(module_gene_sets_sub))))
        frequent_genes_tmp <- frequent_genes_tmp[
          frequent_genes_tmp$Var1 %in% gene_anno$ensembl_gene_id,
        ]
        rownames(frequent_genes_tmp) <- NULL
        # if (any(frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[1]])) {
        #   discard <- which(
        #     frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[1]] &
        #       frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[2]]
        #   )
        #   if (length(discard) > 0) {
        #     frequent_genes_tmp <- frequent_genes_tmp[-discard, ]
        #   }
        # }
        # frequent_genes_tmp <- frequent_genes_tmp[frequent_genes_tmp$Freq >= 2, ]
        # frequent_genes_tmp <- frequent_genes_tmp[
        #   frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[3]],
        # ]
        frequent_genes_down <- frequent_genes_tmp
        frequent_genes <- rbind(frequent_genes_up, frequent_genes_down)

        node_data <- melt(geneIds(gene_sets_sub))
        node_data <- node_data[node_data$value %in% frequent_genes$Var1, ]
        names(node_data)[1] <- "ensembl_gene_id"
        names(node_data)[2] <- "gene_set"
        node_data$direction <- ifelse(
          node_data$gene_set %in% gsva_list_pseudobulk_all[[2]], "up", "down"
        )

        unique_direction <- as.data.frame.matrix(table(node_data[ , c(1, 3)]), )
        unique_direction$down <- unique_direction$down * -1
        # discard <- which(rowSums(unique_direction) == 0)
        # if (length(discard) > 0) {
        #   unique_direction <- unique_direction[-discard, ]
        # }
        module_membership_norm <- unique_direction
        # unique_direction <- ifelse(rowSums(unique_direction) > 0, "up", "down")
        # node_data <- node_data[node_data$ensembl_gene_id %in% names(unique_direction), ]
        # frequent_genes_tmp <- as.data.frame(table(node_data$ensembl_gene_id))
        # frequent_genes <- frequent_genes_tmp
        module_membership_norm <- module_membership_norm[
          abs(rowSums(module_membership_norm)) >= 2,
        ]

        hub_gene_all_GO <- table(unlist(geneIds(gene_sets)))
        hub_gene_all_GO <- hub_gene_all_GO[
          names(hub_gene_all_GO) %in% rownames(module_membership_norm)
        ]

        # mart <- useEnsembl("ensembl", paste0(organism, "_gene_ensembl"))
        # attributes <- c("ensembl_gene_id", "go_id")
        # hub_gene_all_GO <- getBM(
        #   attributes, "ensembl_gene_id", rownames(module_membership_norm), mart
        # )
        # hub_gene_all_GO <- as.data.frame.matrix(table(hub_gene_all_GO), )
        # hub_gene_all_GO <- rowSums(hub_gene_all_GO[ , -1])

        # module_membership_upper_quart <- summary(
        #   as.numeric(
        #     (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
        #     abs(rowSums(module_membership_norm))
        #   )
        # )[5][[1]]
        # module_membership_upper_quart <- summary(abs(rowSums(module_membership_norm)))[5][[1]]
        module_membership_upper_quart <- median(
          (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
            abs(rowSums(module_membership_norm))
        )
        # module_membership_upper_quart <- summary(
        #   as.numeric(
        #     (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
        #     abs(rowSums(module_membership_norm))
        #   )
        # )[2][[1]]

        modules_with_hubs <- vector("list", 25)
        modules_with_hubs[[1]] <- 0
        hub_gene_list <- list(gene_anno)
        for (l in 2:25) {
          set.seed(1)
          heatmap <- Heatmap(
            mat, cluster_rows = clustering, cluster_columns = FALSE, show_column_names = FALSE,
            show_row_names = FALSE, show_heatmap_legend = FALSE, row_split = l,
          )
          modules <- row_order(heatmap)
          hub_gene_list <- vector("list", length(modules))
          for (m in seq_along(modules)) {
            module_sets <- rownames(mat)[modules[[m]]]
            modules_up <- module_sets[module_sets %in% gsva_list_pseudobulk_all[[2]]]
            modules_down <- module_sets[module_sets %in% gsva_list_pseudobulk_all[[3]]]
            module_hub_genes <- vector()

            if (length(modules_up) == 0) {
            } else {
              module_gene_sets_sub <- gene_sets[names(gene_sets) %in% modules_up]
              frequent_genes <- as.data.frame(table(unlist(geneIds(module_gene_sets_sub))))
              hub_gene_all_GO_sub <- hub_gene_all_GO[
                names(hub_gene_all_GO) %in% frequent_genes$Var1
              ]
              frequent_genes <- frequent_genes[frequent_genes$Var1 %in% names(hub_gene_all_GO), ]
              frequent_genes$Freq <-
                (frequent_genes$Freq / hub_gene_all_GO_sub) * frequent_genes$Freq
              frequent_genes <- frequent_genes[
                frequent_genes$Freq >= module_membership_upper_quart,
              ]
              module_hub_genes_tmp <- gene_anno[
                gene_anno$ensembl_gene_id %in% frequent_genes$Var1,
              ]
              module_hub_genes_tmp <- module_hub_genes_tmp[
                module_hub_genes_tmp$ensembl_gene_id %in% markers_list_pseudobulk_all[[2]],
              ]
              module_hub_genes <- module_hub_genes_tmp$ensembl_gene_id
            }

            if (length(modules_down) == 0) {
            } else {
              module_gene_sets_sub <- gene_sets[names(gene_sets) %in% modules_down]
              frequent_genes <- as.data.frame(table(unlist(geneIds(module_gene_sets_sub))))
              hub_gene_all_GO_sub <- hub_gene_all_GO[
                names(hub_gene_all_GO) %in% frequent_genes$Var1
              ]
              frequent_genes <- frequent_genes[frequent_genes$Var1 %in% names(hub_gene_all_GO), ]
              frequent_genes$Freq <-
                (frequent_genes$Freq / hub_gene_all_GO_sub) * frequent_genes$Freq
              frequent_genes <- frequent_genes[
                frequent_genes$Freq >= module_membership_upper_quart,
              ]
              module_hub_genes_tmp <- gene_anno[
                gene_anno$ensembl_gene_id %in% frequent_genes$Var1,
              ]
              module_hub_genes_tmp <- module_hub_genes_tmp[
                module_hub_genes_tmp$ensembl_gene_id %in% markers_list_pseudobulk_all[[3]],
              ]
              module_hub_genes <- append(module_hub_genes, module_hub_genes_tmp$ensembl_gene_id)
            }

            hub_gene_list[[m]] <- module_hub_genes
          }
          modules_with_hubs[[l]] <- length(which(lengths(hub_gene_list) != 0)) / l
        }
        optimal_k <- 26 - which.max(rev(modules_with_hubs))
        module_colors <- cols25(optimal_k)
        clustering <- color_branches(clustering, optimal_k, col = module_colors)
        set.seed(1)
        heatmap <- Heatmap(
          mat, cluster_rows = clustering, cluster_columns = FALSE, show_column_names = FALSE,
          show_row_names = FALSE, show_heatmap_legend = FALSE, row_split = optimal_k
        )

        modules <- row_order(heatmap)
        module_names <- modules
        go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
        # for (l in seq_along(module_names)) {
        #   module_names[[l]] <- paste0(
        #     rownames(mat)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]]
        #   )
        # }
        for (l in seq_along(module_names)) {
          module_names[[l]] <- paste0(rownames(mat)[module_names[[l]]])
        }
        top_words_tmp <- data.frame(
          gene_sets = unlist(module_names),
          module = rep(seq_along(module_names), lapply(module_names, length))
        )
        top_words_all <- suppressWarnings(
          as.character(
            tapply(top_words_tmp$gene_sets, top_words_tmp$module, word_cloud)
          )
        )
        # width <- min(lengths(strsplit(top_words, " ")))
        top_words <- suppressWarnings(
          as.character(
            tapply(top_words_tmp$gene_sets, top_words_tmp$module, word_cloud, width = 5)
          )
        )
        row_anno <-  HeatmapAnnotation(
          top_words = anno_block(
            gp = gpar(col = module_colors), labels = seq(optimal_k),
            labels_rot = 0, labels_gp = gpar(fontsize = 20)
          ),
          which = "row"
        )

        row_colors <- rownames(mat)
        row_colors <- data.frame(
          row_colors = row_colors,
          DE_strict = row_colors %in% rownames(gsva_list_pseudobulk_strict[[1]]),
          DE_stricter = row_colors %in% rownames(gsva_list_pseudobulk_stricter[[1]]),
          DE_strictest = row_colors %in% rownames(gsva_list_pseudobulk_strictest[[1]])
        ) %>%
          mutate(
            type = case_when(
              DE_strictest == TRUE ~ "firebrick",
              DE_stricter == TRUE ~ "darkorange",
              DE_strict == TRUE ~ "darkgoldenrod",
              TRUE ~ "black"
            )
          )
        row_colors <- row_colors$type

        gsva_heatmap[[i]] <- Heatmap(
          t(
            apply(
              mat, 1,
              function (x) (x - min(x)) / (max(x) - min(x))
            )
          ),
          color,
          cluster_rows = clustering,
          cluster_columns = FALSE,
          row_names_max_width = max_text_width(rownames(mat), gpar(fontsize = 8)),
          show_column_names = FALSE,
          column_split = sce$idents,
          top_annotation = anno,
          rect_gp = gpar(col = "lightgray"),
          column_title = " ",
          row_split = optimal_k,
          row_title = top_words,
          row_title_rot = 0,
          row_title_gp = gpar(fontsize = 10),
          row_names_gp = gpar(fontsize = 8, col = unlist(row_colors)),
          left_annotation = row_anno,
          row_gap = unit(2, "mm"),
          column_gap = unit(2, "mm"),
          heatmap_legend_param = list(title = "scaled ssGSEA score")
        )
        draw(gsva_heatmap[[i]])

        gsva_heatmap[[i]] <- Heatmap(
          mat,
          color2,
          cluster_rows = clustering,
          cluster_columns = FALSE,
          row_names_max_width = max_text_width(rownames(mat), gpar(fontsize = 8)),
          show_column_names = FALSE,
          column_split = sce$idents,
          top_annotation = anno,
          rect_gp = gpar(col = "lightgray"),
          column_title = " ",
          row_split = optimal_k,
          row_title = top_words,
          row_title_rot = 0,
          row_title_gp = gpar(fontsize = 10),
          row_names_gp = gpar(fontsize = 8, col = unlist(row_colors)),
          left_annotation = row_anno,
          row_gap = unit(2, "mm"),
          column_gap = unit(2, "mm"),
          heatmap_legend_param = list(title = "ssGSEA score")
        )
        draw(gsva_heatmap[[i]])

        for (l in 2:3) {
          exp_mat <- mat[rownames(mat) %in% gsva_list_pseudobulk_all[[l]], ]
          sort_mat <- mat[unlist(row_order(heatmap)), ]
          sort_mat <- sort_mat[rownames(sort_mat) %in% rownames(exp_mat), ]
          exp_mat <- exp_mat[match(rownames(sort_mat), rownames(exp_mat)), ]
          draw(
            Heatmap(
              exp_mat,
              color2,
              cluster_columns = FALSE,
              row_names_max_width = max_text_width(rownames(exp_mat),gpar(fontsize = 12)),
              show_column_names = FALSE,
              column_split = sce$idents,
              top_annotation = anno,
              row_names_gp = gpar(fontsize = 12),
              rect_gp = gpar(col = "lightgray"),
              column_title = " ",
              heatmap_legend_param = list(title = "ssGSEA score"),
              column_gap = unit(2, "mm")
            )
          )
        }

        if (i == 1) {
          hub_gene_data <- vector()
          for (l in seq_along(modules)) {
            module_sets <- rownames(mat)[modules[[l]]]
            modules_up <- module_sets[module_sets %in% gsva_list_pseudobulk_all[[2]]]
            modules_down <- module_sets[module_sets %in% gsva_list_pseudobulk_all[[3]]]
            frequent_genes <- vector()
            frequent_genes_up <- vector()
            frequent_genes_down <- vector()

            if (length(modules_up) == 0) {
            } else {
              module_gene_sets_sub <- gene_sets[names(gene_sets) %in% modules_up]
              frequent_genes_tmp <- as.data.frame(table(unlist(geneIds(module_gene_sets_sub))))
              frequent_genes_tmp <- frequent_genes_tmp[
                frequent_genes_tmp$Var1 %in% gene_anno$ensembl_gene_id,
              ]
              rownames(frequent_genes_tmp) <- NULL
              if (any(frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[1]])) {
                discard <- which(
                  frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[1]] &
                  frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[3]]
                )
                if (length(discard) > 0) {
                  frequent_genes_tmp <- frequent_genes_tmp[-discard, ]
                }
              }
              # frequent_genes_tmp <- frequent_genes_tmp[frequent_genes_tmp$Freq >= 2, ]
              # frequent_genes_tmp <- frequent_genes_tmp[
              #   frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[2]],
              # ]
              frequent_genes_up <- frequent_genes_tmp
            }

            if (length(modules_down) == 0) {
            } else {
              module_gene_sets_sub <- gene_sets[names(gene_sets) %in% modules_down]
              frequent_genes_tmp <- as.data.frame(table(unlist(geneIds(module_gene_sets_sub))))
              frequent_genes_tmp <- frequent_genes_tmp[
                frequent_genes_tmp$Var1 %in% gene_anno$ensembl_gene_id,
              ]
              rownames(frequent_genes_tmp) <- NULL
              if (any(frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[1]])) {
                discard <- which(
                  frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[1]] &
                  frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[2]]
                )
                if (length(discard) > 0) {
                  frequent_genes_tmp <- frequent_genes_tmp[-discard, ]
                }
              }
              # frequent_genes_tmp <- frequent_genes_tmp[frequent_genes_tmp$Freq >= 2, ]
              # frequent_genes_tmp <- frequent_genes_tmp[
              #   frequent_genes_tmp$Var1 %in% markers_list_pseudobulk_all[[3]],
              # ]
              frequent_genes_down <- frequent_genes_tmp
            }

            frequent_genes <- rbind(frequent_genes_up, frequent_genes_down)
            if (dim(frequent_genes)[1] != 0) {
              module_gene_sets_sub <- gene_sets[names(gene_sets) %in% module_sets]
              node_data <- melt(geneIds(module_gene_sets_sub))
              node_data <- node_data[node_data$value %in% frequent_genes$Var1, ]
              names(node_data)[1] <- "ensembl_gene_id"
              names(node_data)[2] <- "gene_set"
              gene_anno_sub <- gene_anno[
                gene_anno$ensembl_gene_id %in% node_data$ensembl_gene_id, -3
              ]
              node_data <- node_data[node_data$ensembl_gene_id %in% gene_anno_sub$ensembl_gene_id, ]
              node_data <- left_join(node_data, gene_anno_sub, by = "ensembl_gene_id")
              node_data$direction <- ifelse(
                node_data$gene_set %in% gsva_list_pseudobulk_all[[2]], "up", "down"
              )
              node_data$DE <- ifelse(
                node_data$ensembl_gene_id %in% markers_list_pseudobulk_all[[1]], "yes", "no"
              )
              node_data$module <- l
              hub_gene_data <- rbind(hub_gene_data, node_data)
            }
          }

          unique_direction <- as.data.frame.matrix(table(hub_gene_data[ , c(3, 4)]), )
          unique_direction$down <- unique_direction$down * -1
          discard <- which(rowSums(unique_direction) == 0)
          if (length(discard) > 0) {
            unique_direction <- unique_direction[-discard, ]
          }
          module_membership_norm <- unique_direction
          unique_direction <- ifelse(rowSums(unique_direction) > 0, "up", "down")
          hub_gene_data <- hub_gene_data[
            hub_gene_data$external_gene_name %in% names(unique_direction),
          ]

          rownames(hub_gene_data) <- NULL
          discard <- vector()
          for (i_gene in seq_along(hub_gene_data$external_gene_name)) {
            sub_gene <- hub_gene_data$external_gene_name[i_gene]
            sub_unique_direction <- unique_direction[which(names(unique_direction) == sub_gene)]
            if (hub_gene_data$direction[i_gene] != sub_unique_direction[[1]]) {
              discard <- append(discard, i_gene)
            }
          }
          hub_gene_data <- hub_gene_data[-discard, ]

          frequent_genes_tmp <- as.data.frame(table(hub_gene_data$external_gene_name))
          frequent_genes <- frequent_genes_tmp
          # frequent_genes <- frequent_genes_tmp[frequent_genes_tmp$Freq >= 2, ]
          hub_gene_data <- hub_gene_data[
            hub_gene_data$external_gene_name %in% frequent_genes$Var1,
          ]

          hub_gene_data_up <- hub_gene_data[which(hub_gene_data$direction == "up"), ]
          hub_gene_data_down <- hub_gene_data[which(hub_gene_data$direction == "down"), ]

          hub_mat <- as.data.frame.matrix(table(hub_gene_data_up[ , c(3, 2)]), )
          hub_mat <- hub_mat == 1
          co_mat <- hub_mat %*% t(hub_mat)
          diag(co_mat) <- 0
          graph_up <- graph_from_adjacency_matrix(co_mat, "upper", TRUE)

          edges_up <- get.edgelist(graph_up, FALSE)
          set.seed(1)
          layout_up <- qgraph.layout.fruchtermanreingold(
            edges_up, vcount = vcount(graph_up), E(graph_up)$weight,
            area = vcount(graph_up) ^ 2, repulse.rad = vcount(graph_up) ^ 2
          )

          # mart <- useEnsembl("ensembl", paste0(organism, "_gene_ensembl"))
          # attributes <- c("ensembl_gene_id", "go_id")
          # hub_gene_all_GO <- getBM(
          #   attributes, "ensembl_gene_id", unique(hub_gene_data$ensembl_gene_id), mart
          # )
          # gene_anno_hub_sub <- filterGeneSets(gene_sets, 3, 500)
          # for (i in seq(length(gene_anno_hub_sub@.Data))) {
          #   go_id <- gene_anno_hub_sub[[i]]@setName
          #   suppressWarnings(
          #     gene_anno_hub_sub[[i]]@setName <- gene_anno_hub_sub[[i]]@shortDescription
          #   )
          #   suppressWarnings(gene_anno_hub_sub[[i]]@shortDescription <- go_id)
          # }
          # hub_gene_all_GO <- hub_gene_all_GO[hub_gene_all_GO$go_id %in% names(gene_anno_hub_sub), ]
          # hub_gene_all_GO <- as.data.frame.matrix(table(hub_gene_all_GO), )
          # hub_gene_all_GO <- hub_gene_all_GO[ , -1]

          hub_gene_all_GO <- table(unlist(geneIds(gene_sets)))
          hub_gene_all_GO <- hub_gene_all_GO[
            names(hub_gene_all_GO) %in% unique(hub_gene_data_up$ensembl_gene_id)
          ]
          sub_gene_anno <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(hub_gene_all_GO), ]
          rownames(sub_gene_anno) <- NULL
          sub_gene_anno <- sub_gene_anno[
            match(rownames(hub_gene_all_GO), sub_gene_anno$ensembl_gene_id),
          ]
          rownames(hub_gene_all_GO) <- sub_gene_anno$external_gene_name
          # hub_gene_all_GO <- rowSums(hub_gene_all_GO)
          hub_gene_all_GO_up <- hub_gene_all_GO[order(names(hub_gene_all_GO))]

          module_membership_up <- as.data.frame.matrix(table(hub_gene_data_up[ , c(3, 6)]), )
          module_membership_binary_up <- module_membership_up
          module_membership_binary_up[module_membership_binary_up > 0] <- 1
          module_membership_list_up <- lapply(
            split(module_membership_up, row.names(module_membership_up)), unlist
          )
          module_network_colors_up <- module_colors[unique(hub_gene_data_up$module)]
          V(graph_up)$pie.color = list(module_network_colors_up)

          module_membership_norm_up <- module_membership_norm[
            rownames(module_membership_norm) %in% rownames(module_membership_up),
          ]

          V(graph_up)$vertex.pie = list(module_network_colors_up)
          V(graph_up)$vertex.shape <- "pie"
          V(graph_up)$vertex.color <- module_membership_list_up
          V(graph_up)$vertex.label.family = "sans"

          target_all_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% open_targets_all) {
              target_all_up <- append(target_all_up, "yes")
            } else {
              target_all_up <- append(target_all_up, "no")
            }
          }

          DE_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% rownames(markers_list_symbols_pseudobulk[[2]])) {
              DE_up <- append(DE_up, "yes")
            } else {
              DE_up <- append(DE_up, "no")
            }
          }

          DE_opp_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% markers_list_symbols_pseudobulk_all[[3]]) {
              DE_opp_up <- append(DE_opp_up, "yes")
            } else {
              DE_opp_up <- append(DE_opp_up, "no")
            }
          }

          DE_strict_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% rownames(markers_list_symbols_pseudobulk_strict[[2]])) {
              DE_strict_up <- append(DE_strict_up, "yes")
            } else {
              DE_strict_up <- append(DE_strict_up, "no")
            }
          }

          DE_stricter_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% rownames(markers_list_symbols_pseudobulk_stricter[[2]])) {
              DE_stricter_up <- append(DE_stricter_up, "yes")
            } else {
              DE_stricter_up <- append(DE_stricter_up, "no")
            }
          }

          DE_strictest_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% rownames(markers_list_symbols_pseudobulk_strictest[[2]])) {
              DE_strictest_up <- append(DE_strictest_up, "yes")
            } else {
              DE_strictest_up <- append(DE_strictest_up, "no")
            }
          }

          ECvuln_DEup_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% kampmann_comb_ECvuln_AD_CTRL_DE_up) {
              ECvuln_DEup_up <- append(ECvuln_DEup_up, "yes")
            } else {
              ECvuln_DEup_up <- append(ECvuln_DEup_up, "no")
            }
          }

          ECvuln_DEdown_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% kampmann_comb_ECvuln_AD_CTRL_DE_down) {
              ECvuln_DEdown_up <- append(ECvuln_DEdown_up, "yes")
            } else {
              ECvuln_DEdown_up <- append(ECvuln_DEdown_up, "no")
            }
          }

          SFGvuln_DEup_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% kampmann_comb_SFGvuln_AD_CTRL_DE_up) {
              SFGvuln_DEup_up <- append(SFGvuln_DEup_up, "yes")
            } else {
              SFGvuln_DEup_up <- append(SFGvuln_DEup_up, "no")
            }
          }

          SFGvuln_DEdown_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% kampmann_comb_SFGvuln_AD_CTRL_DE_down) {
              SFGvuln_DEdown_up <- append(SFGvuln_DEdown_up, "yes")
            } else {
              SFGvuln_DEdown_up <- append(SFGvuln_DEdown_up, "no")
            }
          }

          SFGinvuln_DEup_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% kampmann_comb_SFGinvuln_AD_CTRL_DE_up) {
              SFGinvuln_DEup_up <- append(SFGinvuln_DEup_up, "yes")
            } else {
              SFGinvuln_DEup_up <- append(SFGinvuln_DEup_up, "no")
            }
          }

          SFGinvuln_DEdown_up <- vector()
          for (l in seq(length(V(graph_up)))) {
            if (V(graph_up)[l]$name %in% kampmann_comb_SFGinvuln_AD_CTRL_DE_down) {
              SFGinvuln_DEdown_up <- append(SFGinvuln_DEdown_up, "yes")
            } else {
              SFGinvuln_DEdown_up <- append(SFGinvuln_DEdown_up, "no")
            }
          }

          direction_up <- vector()
          for (l in seq(length(E(graph_up)))) {
            hub_gene_sub_up <- hub_gene_data_up[
              hub_gene_data_up$external_gene_name %in% head_of(graph_up, l)[[1]]$name,
            ]
            direction_up <- append(direction_up, unique(hub_gene_sub_up$direction))
          }
          E(graph_up)$edge.color <- ifelse(direction_up == "up", "firebrick1", "dodgerblue")

          hub_mat <- as.data.frame.matrix(table(hub_gene_data_down[ , c(3, 2)]), )
          hub_mat <- hub_mat == 1
          co_mat <- hub_mat %*% t(hub_mat)
          diag(co_mat) <- 0
          graph_down <- graph_from_adjacency_matrix(co_mat, "upper", TRUE)

          edges_down <- get.edgelist(graph_down, FALSE)
          set.seed(1)
          layout_down <- qgraph.layout.fruchtermanreingold(
            edges_down, vcount = vcount(graph_down), E(graph_down)$weight,
            area = vcount(graph_down) ^ 2, repulse.rad = vcount(graph_down) ^ 2
          )

          # mart <- useEnsembl("ensembl", paste0(organism, "_gene_ensembl"))
          # attributes <- c("ensembl_gene_id", "go_id")
          # hub_gene_all_GO <- getBM(
          #   attributes, "ensembl_gene_id", unique(hub_gene_data$ensembl_gene_id), mart
          # )
          # gene_anno_hub_sub <- filterGeneSets(gene_sets, 3, 500)
          # for (i in seq(length(gene_anno_hub_sub@.Data))) {
          #   go_id <- gene_anno_hub_sub[[i]]@setName
          #   suppressWarnings(
          #     gene_anno_hub_sub[[i]]@setName <- gene_anno_hub_sub[[i]]@shortDescription
          #   )
          #   suppressWarnings(gene_anno_hub_sub[[i]]@shortDescription <- go_id)
          # }
          # hub_gene_all_GO <- hub_gene_all_GO[hub_gene_all_GO$go_id %in% names(gene_anno_hub_sub), ]
          # hub_gene_all_GO <- as.data.frame.matrix(table(hub_gene_all_GO), )
          # hub_gene_all_GO <- hub_gene_all_GO[ , -1]

          hub_gene_all_GO <- table(unlist(geneIds(gene_sets)))
          hub_gene_all_GO <- hub_gene_all_GO[
            names(hub_gene_all_GO) %in% unique(hub_gene_data_down$ensembl_gene_id)
          ]
          sub_gene_anno <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(hub_gene_all_GO), ]
          rownames(sub_gene_anno) <- NULL
          sub_gene_anno <- sub_gene_anno[
            match(rownames(hub_gene_all_GO), sub_gene_anno$ensembl_gene_id),
          ]
          rownames(hub_gene_all_GO) <- sub_gene_anno$external_gene_name
          # hub_gene_all_GO <- rowSums(hub_gene_all_GO)
          hub_gene_all_GO_down <- hub_gene_all_GO[order(names(hub_gene_all_GO))]

          module_membership_down <- as.data.frame.matrix(table(hub_gene_data_down[ , c(3, 6)]), )
          module_membership_binary_down <- module_membership_down
          module_membership_binary_down[module_membership_binary_down > 0] <- 1
          module_membership_list_down <- lapply(
            split(module_membership_down, row.names(module_membership_down)), unlist
          )
          module_network_colors_down <- module_colors[unique(hub_gene_data_down$module)]
          V(graph_down)$pie.color = list(module_network_colors_down)

          module_membership_norm_down <- module_membership_norm[
            rownames(module_membership_norm) %in% rownames(module_membership_down),
          ]

          V(graph_down)$pie.color = list(module_network_colors_down)
          V(graph_down)$vertex.shape <- "pie"
          V(graph_down)$vertex.color <- module_membership_list_down
          V(graph_down)$vertex.label.family = "sans"

          target_all_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% open_targets_all) {
              target_all_down <- append(target_all_down, "yes")
            } else {
              target_all_down <- append(target_all_down, "no")
            }
          }

          DE_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% rownames(markers_list_symbols_pseudobulk[[3]])) {
              DE_down <- append(DE_down, "yes")
            } else {
              DE_down <- append(DE_down, "no")
            }
          }

          DE_opp_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% markers_list_symbols_pseudobulk_all[[2]]) {
              DE_opp_down <- append(DE_opp_down, "yes")
            } else {
              DE_opp_down <- append(DE_opp_down, "no")
            }
          }

          DE_strict_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% rownames(markers_list_symbols_pseudobulk_strict[[3]])) {
              DE_strict_down <- append(DE_strict_down, "yes")
            } else {
              DE_strict_down <- append(DE_strict_down, "no")
            }
          }

          DE_stricter_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (
              V(graph_down)[l]$name %in% rownames(markers_list_symbols_pseudobulk_stricter[[3]])
            ) {
              DE_stricter_down <- append(DE_stricter_down, "yes")
            } else {
              DE_stricter_down <- append(DE_stricter_down, "no")
            }
          }

          DE_strictest_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (
              V(graph_down)[l]$name %in% rownames(markers_list_symbols_pseudobulk_strictest[[3]])
            ) {
              DE_strictest_down <- append(DE_strictest_down, "yes")
            } else {
              DE_strictest_down <- append(DE_strictest_down, "no")
            }
          }

          ECvuln_DEup_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% kampmann_comb_ECvuln_AD_CTRL_DE_up) {
              ECvuln_DEup_down <- append(ECvuln_DEup_down, "yes")
            } else {
              ECvuln_DEup_down <- append(ECvuln_DEup_down, "no")
            }
          }

          ECvuln_DEdown_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% kampmann_comb_ECvuln_AD_CTRL_DE_down) {
              ECvuln_DEdown_down <- append(ECvuln_DEdown_down, "yes")
            } else {
              ECvuln_DEdown_down <- append(ECvuln_DEdown_down, "no")
            }
          }

          SFGvuln_DEup_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% kampmann_comb_SFGvuln_AD_CTRL_DE_up) {
              SFGvuln_DEup_down <- append(SFGvuln_DEup_down, "yes")
            } else {
              SFGvuln_DEup_down <- append(SFGvuln_DEup_down, "no")
            }
          }

          SFGvuln_DEdown_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% kampmann_comb_SFGvuln_AD_CTRL_DE_down) {
              SFGvuln_DEdown_down <- append(SFGvuln_DEdown_down, "yes")
            } else {
              SFGvuln_DEdown_down <- append(SFGvuln_DEdown_down, "no")
            }
          }

          SFGinvuln_DEup_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% kampmann_comb_SFGinvuln_AD_CTRL_DE_up) {
              SFGinvuln_DEup_down <- append(SFGinvuln_DEup_down, "yes")
            } else {
              SFGinvuln_DEup_down <- append(SFGinvuln_DEup_down, "no")
            }
          }

          SFGinvuln_DEdown_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% kampmann_comb_SFGinvuln_AD_CTRL_DE_down) {
              SFGinvuln_DEdown_down <- append(SFGinvuln_DEdown_down, "yes")
            } else {
              SFGinvuln_DEdown_down <- append(SFGinvuln_DEdown_down, "no")
            }
          }

          direction_down <- vector()
          for (l in seq(length(E(graph_down)))) {
            hub_gene_sub_down <- hub_gene_data_down[
              hub_gene_data_down$external_gene_name %in% head_of(graph_down, l)[[1]]$name,
            ]
            direction_down <- append(direction_down, unique(hub_gene_sub_down$direction))
          }
          E(graph_down)$edge.color <- ifelse(direction_down == "up", "firebrick1", "dodgerblue")

          tmp <- vector()
          for (l in seq(length(E(graph_down)))) {
            if (head_of(graph_down, l)[[1]]$name == "Mapt") {
              tmp <- append(tmp, tail_of(graph_down, l)[[1]]$name)
            }
          }
          mapt_head_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% "Mapt") {
              mapt_head_down <- append(mapt_head_down, "yes")
            } else {
              mapt_head_down <- append(mapt_head_down, "no")
            }
          }
          mapt_tail_down <- vector()
          for (l in seq(length(V(graph_down)))) {
            if (V(graph_down)[l]$name %in% tmp) {
              mapt_tail_down <- append(mapt_tail_down, "yes")
            } else {
              mapt_tail_down <- append(mapt_tail_down, "no")
            }
          }

          hub_mat <- as.data.frame.matrix(table(hub_gene_data[ , c(3, 2)]), )
          hub_mat <- hub_mat == 1
          co_mat <- hub_mat %*% t(hub_mat)
          diag(co_mat) <- 0
          graph <- graph_from_adjacency_matrix(co_mat, "upper", TRUE)

          edges <- get.edgelist(graph, FALSE)
          set.seed(1)
          layout <- qgraph.layout.fruchtermanreingold(
            edges, vcount = vcount(graph), E(graph)$weight,
            area = vcount(graph) ^ 2, repulse.rad = vcount(graph) ^ 2
          )

          # mart <- useEnsembl("ensembl", paste0(organism, "_gene_ensembl"))
          # attributes <- c("ensembl_gene_id", "go_id")
          # hub_gene_all_GO <- getBM(
          #   attributes, "ensembl_gene_id", unique(hub_gene_data$ensembl_gene_id), mart
          # )
          # gene_anno_hub_sub <- filterGeneSets(gene_sets, 3, 500)
          # for (i in seq(length(gene_anno_hub_sub@.Data))) {
          #   go_id <- gene_anno_hub_sub[[i]]@setName
          #   suppressWarnings(
          #     gene_anno_hub_sub[[i]]@setName <- gene_anno_hub_sub[[i]]@shortDescription
          #   )
          #   suppressWarnings(gene_anno_hub_sub[[i]]@shortDescription <- go_id)
          # }
          # hub_gene_all_GO <- hub_gene_all_GO[hub_gene_all_GO$go_id %in% names(gene_anno_hub_sub), ]
          # hub_gene_all_GO <- as.data.frame.matrix(table(hub_gene_all_GO), )
          # hub_gene_all_GO <- hub_gene_all_GO[ , -1]

          hub_gene_all_GO <- table(unlist(geneIds(gene_sets)))
          hub_gene_all_GO <- hub_gene_all_GO[
            names(hub_gene_all_GO) %in% unique(hub_gene_data$ensembl_gene_id)
          ]
          sub_gene_anno <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(hub_gene_all_GO), ]
          rownames(sub_gene_anno) <- NULL
          sub_gene_anno <- sub_gene_anno[
            match(rownames(hub_gene_all_GO), sub_gene_anno$ensembl_gene_id),
          ]
          rownames(hub_gene_all_GO) <- sub_gene_anno$external_gene_name
          # hub_gene_all_GO <- rowSums(hub_gene_all_GO)
          hub_gene_all_GO <- hub_gene_all_GO[order(names(hub_gene_all_GO))]

          module_membership <- as.data.frame.matrix(table(hub_gene_data[ , c(3, 6)]), )
          module_membership_binary <- module_membership
          module_membership_binary[module_membership_binary > 0] <- 1
          module_membership_list <- lapply(
            split(module_membership, row.names(module_membership)), unlist
          )
          module_network_colors <- module_colors
          module_membership_norm <- module_membership_norm[
            rownames(module_membership_norm) %in% rownames(module_membership),
          ]

          V(graph)$pie.color = list(module_network_colors)
          V(graph)$vertex.shape <- "pie"
          V(graph)$vertex.color <- module_membership_list
          V(graph)$vertex.label.family = "sans"

          module_max_genes <- apply(module_membership_norm, 2, function (x) which(x == max(x)))
          module_max_genes <- rownames(module_membership_norm)[unlist(module_max_genes)]
          module_max <- vector()
          for (l in seq(length(V(graph)))) {
            if (V(graph)[l]$name %in% module_max_genes) {
              module_max <- append(module_max, "yes")
            } else {
              module_max <- append(module_max, "no")
            }
          }

          target_all <- vector()
          for (l in seq(length(V(graph)))) {
            if (V(graph)[l]$name %in% open_targets_all) {
              target_all <- append(target_all, "yes")
            } else {
              target_all <- append(target_all, "no")
            }
          }

          unique_markers <- markers_list_symbols_pseudobulk_all[[2]][
            markers_list_symbols_pseudobulk_all[[2]] %notin%
              hub_gene_data_down$external_gene_name
          ]
          unique_markers <- append(
            unique_markers,
            markers_list_symbols_pseudobulk_all[[3]][
              markers_list_symbols_pseudobulk_all[[3]] %notin%
                hub_gene_data_up$external_gene_name
            ]
          )
          DE <- vector()
          for (l in seq(length(V(graph)))) {
            if (V(graph)[l]$name %in% unique_markers) {
              DE <- append(DE, "yes")
            } else {
              DE <- append(DE, "no")
            }
          }

          # kampmann <- vector()
          # for (l in seq(length(V(graph)))) {
          #   if (V(graph)[l]$name %in% kampmann_comb_ECvuln_AD_CTRL_DE) {
          #     kampmann <- append(kampmann, "yes")
          #   } else {
          #     kampmann <- append(kampmann, "no")
          #   }
          # }

          kampmann_all_genes <- vector()
          for (l in seq(length(V(graph)))) {
            if (V(graph)[l]$name %in% kampmann_all) {
              kampmann_all_genes <- append(kampmann_all_genes, "yes")
            } else {
              kampmann_all_genes <- append(kampmann_all_genes, "no")
            }
          }

          direction <- vector()
          for (l in seq(length(E(graph)))) {
            hub_gene_sub <- hub_gene_data[
              hub_gene_data$external_gene_name %in% head_of(graph, l)[[1]]$name,
            ]
            direction <- append(direction, unique(hub_gene_sub$direction))
          }
          E(graph)$edge.color <- ifelse(direction == "up", "firebrick1", "dodgerblue")

          tmp <- vector()
          for (l in seq(length(E(graph)))) {
            if (head_of(graph, l)[[1]]$name == "Mapt") {
              tmp <- append(tmp, tail_of(graph, l)[[1]]$name)
            }
          }
          mapt_head <- vector()
          for (l in seq(length(V(graph)))) {
            if (V(graph)[l]$name %in% "Mapt") {
              mapt_head <- append(mapt_head, "yes")
            } else {
              mapt_head <- append(mapt_head, "no")
            }
          }
          mapt_tail <- vector()
          for (l in seq(length(V(graph)))) {
            if (V(graph)[l]$name %in% tmp) {
              mapt_tail <- append(mapt_tail, "yes")
            } else {
              mapt_tail <- append(mapt_tail, "no")
            }
          }

        }

      } else {
        set.seed(1)
        gsva_heatmap[[i]] <- Heatmap(
          mat,
          color2,
          cluster_rows = clustering,
          cluster_columns = FALSE,
          row_names_max_width = max_text_width(rownames(mat)),
          show_column_names = FALSE,
          column_split = sce$idents,
          show_heatmap_legend = FALSE,
          top_annotation = anno,
          rect_gp = gpar(col = "lightgray"),
          column_title = paste0(names(gsva_list_pseudobulk)[i], " GO Terms"),
          row_gap = unit(2, "mm"),
          column_gap = unit(2, "mm")
        )
        draw(gsva_heatmap[[i]])
      }

      # mat <- as.matrix(GetAssayData(seurat, "counts"))
      # mat <- mat[rownames(mat) %in% rownames(gsva_list_pseudobulk[[i]]), ]
      # mat <- mat[rownames(mat) %in% names(gene_sets_sub), ]
      # ha <- HeatmapAnnotation(
      #   donor = seurat$external_donor_name_id,
      #   cell_type = anno_block(
      #     gp = gpar(fill = "lightgray", col = "lightgray"),
      #     labels = c(
      #       names(idents)[[1]], names(idents)[[2]], names(idents)[[3]],
      #       names(idents)[[4]], names(idents)[[5]]
      #     ),
      #     labels_gp = gpar(col = "black")
      #   ),
      #   col = list(
      #     donor = c(
      #       "453" = anno_color[1], "454" = anno_color[2], "448" = anno_color[3],
      #       "490" = anno_color[4], "492" = anno_color[5], "493" = anno_color[6],
      #       "410" = anno_color[7], "441" = anno_color[8], "407" = anno_color[9]
      #     )
      #   )
      # )
      # if (nrow(mat) >= 4) {
      #   set.seed(1)
      #   draw(
      #     Heatmap(
      #       mat,
      #       color,
      #       cluster_rows = clustering,
      #       cluster_columns = FALSE,
      #       row_names_max_width = max_text_width(rownames(mat)),
      #       show_column_names = FALSE,
      #       column_split = seurat$idents,
      #       show_heatmap_legend = FALSE,
      #       top_annotation = ha,
      #       column_title = paste0(names(gsva_list_allcells)[i], " AD Relevent GO Terms"),
      #       row_split = optimal_k,
      #       row_title = top_words,
      #       row_title_rot = 0,
      #       left_annotation = row_anno,
      #       row_gap = unit(2, "mm"),
      #       column_gap = unit(2, "mm")
      #     )
      #   )
      # } else {
      #   set.seed(1)
      #   draw(
      #     Heatmap(
      #       mat,
      #       color,
      #       cluster_rows = clustering,
      #       cluster_columns = FALSE,
      #       row_names_max_width = max_text_width(rownames(mat)),
      #       show_column_names = FALSE,
      #       column_split = seurat$idents,
      #       show_heatmap_legend = FALSE,
      #       top_annotation = ha,
      #       column_title = paste0(names(gsva_list_allcells)[i], " AD Relevent GO Terms"),
      #       row_gap = unit(2, "mm"),
      #       column_gap = unit(2, "mm")
      #     )
      #   )
      # }
      # print(
      #   FeaturePlot(
      #     seurat, rownames(mat)[unlist(row_order(gsva_heatmap[[i]]))[1:4]], order = TRUE,
      #     cols = c("lawngreen", "firebrick1"), ncol = 2
      #   )
      # )
      # print(VlnPlot(seurat, rownames(mat)[unlist(row_order(gsva_heatmap[[i]]))[1:6]], ncol = 2))
    }
  }
}
```

```{r, fig.width = 20}
module_names <- modules
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(mat)[module_names[[l]]])
}
row_colors <- module_names
for (l in seq_along(row_colors)) {
  tmp <- data.frame(
    row_colors = row_colors[[l]],
    DE_strict = row_colors[[l]] %in% rownames(gsva_list_pseudobulk_strict[[1]]),
    DE_stricter = row_colors[[l]] %in% rownames(gsva_list_pseudobulk_stricter[[1]]),
    DE_strictest = row_colors[[l]] %in% rownames(gsva_list_pseudobulk_strictest[[1]])
  ) %>%
    mutate(
      type = case_when(
        DE_strictest == TRUE ~ "firebrick",
        DE_stricter == TRUE ~ "darkorange",
        DE_strict == TRUE ~ "darkgoldenrod",
        TRUE ~ "black"
      )
    )
  tmp <- tmp$type
  row_colors[[l]] <- tmp
}

for (l in seq_along(module_names)) {
  exp_mat <- mat[rownames(mat) %in% module_names[[l]], ]
  sort_mat <- mat[unlist(row_order(heatmap)), ]
  sort_mat <- sort_mat[rownames(sort_mat) %in% rownames(exp_mat), ]
  exp_mat <- exp_mat[match(rownames(sort_mat), rownames(exp_mat)), ]
  draw(
    Heatmap(
      t(
        apply(
          exp_mat, 1,
          function (x) (x - min(x)) / (max(x) - min(x))
        )
      ),
      color,
      cluster_columns = FALSE,
      cluster_rows = FALSE,
      row_names_max_width = max_text_width(rownames(exp_mat), gpar(fontsize = 20)),
      show_column_names = FALSE,
      column_split = sce$idents,
      top_annotation = anno,
      row_names_gp = gpar(fontsize = 20, col = row_colors[[l]]),
      rect_gp = gpar(col = "lightgray"),
      column_title = paste0("Module ", l, ":\n", top_words[l]),
      column_title_gp = gpar(fontsize = 20),
      heatmap_legend_param = list(title = "scaled ssGSEA score"),
      column_gap = unit(2, "mm")
    )
  )
}
```

### Hub Gene Network

```{r}
datatable_download(as.data.frame(top_words_all))
datatable_download(go_anno)
datatable_download(hub_gene_data)
datatable_download(
  data.frame(genes = names(hub_gene_all_GO), freq = hub_gene_all_GO, row.names = NULL)
)

all_DE_EC_up <- c(
  rownames(markers_list_symbols_pseudobulk[[1]]),
  rownames(markers_list_symbols_pseudobulk_strict[[1]]),
  rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
  rownames(markers_list_symbols_pseudobulk_strictest[[1]]),
  kampmann_comb_ECvuln_AD_CTRL_DE_up,
  kampmann_comb_ECvuln_AD_CTRL_DE_strict_up,
  kampmann_comb_ECvuln_AD_CTRL_DE_stricter_up,
  kampmann_comb_ECvuln_AD_CTRL_DE_strictest_up
)
all_DE_EC_down <- c(
  rownames(markers_list_symbols_pseudobulk[[1]]),
  rownames(markers_list_symbols_pseudobulk_strict[[1]]),
  rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
  rownames(markers_list_symbols_pseudobulk_strictest[[1]]),
  kampmann_comb_ECvuln_AD_CTRL_DE_down,
  kampmann_comb_ECvuln_AD_CTRL_DE_strict_down,
  kampmann_comb_ECvuln_AD_CTRL_DE_stricter_down,
  kampmann_comb_ECvuln_AD_CTRL_DE_strictest_down
)
all_DE_SFGvuln_up <- c(
  rownames(markers_list_symbols_pseudobulk[[1]]),
  rownames(markers_list_symbols_pseudobulk_strict[[1]]),
  rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
  rownames(markers_list_symbols_pseudobulk_strictest[[1]]),
  kampmann_comb_SFGinvuln_AD_CTRL_DE_up,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strict_up,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_stricter_up,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strictest_up
)
all_DE_SFGvuln_down <- c(
  rownames(markers_list_symbols_pseudobulk[[1]]),
  rownames(markers_list_symbols_pseudobulk_strict[[1]]),
  rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
  rownames(markers_list_symbols_pseudobulk_strictest[[1]]),
  kampmann_comb_SFGinvuln_AD_CTRL_DE_down,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strict_down,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_stricter_down,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strictest_down
)
all_DE_SFGinvuln_up <- c(
  rownames(markers_list_symbols_pseudobulk[[1]]),
  rownames(markers_list_symbols_pseudobulk_strict[[1]]),
  rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
  rownames(markers_list_symbols_pseudobulk_strictest[[1]]),
  kampmann_comb_SFGinvuln_AD_CTRL_DE_up,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strict_up,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_stricter_up,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strictest_up
)
all_DE_SFGinvuln_down <- c(
  rownames(markers_list_symbols_pseudobulk[[1]]),
  rownames(markers_list_symbols_pseudobulk_strict[[1]]),
  rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
  rownames(markers_list_symbols_pseudobulk_strictest[[1]]),
  kampmann_comb_SFGinvuln_AD_CTRL_DE_down,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strict_down,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_stricter_down,
  kampmann_comb_SFGinvuln_AD_CTRL_DE_strictest_down
)

DE_alt <- c(
  rownames(markers_list_symbols_pseudobulk[[1]]),
  rownames(markers_list_symbols_pseudobulk_strict[[1]]),
  rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
  rownames(markers_list_symbols_pseudobulk_strictest[[1]])
)

module_membership_upper_quart <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
      abs(rowSums(module_membership_norm))
  )
)[5][[1]]
module_membership_median <- median(
  (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
  abs(rowSums(module_membership_norm))
)
module_membership_lower_quart <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
      abs(rowSums(module_membership_norm))
  )
)[2][[1]]

# all_DE_match <- as.data.frame(
#   table(V(graph_up)$name[match(all_DE_EC_up, V(graph_up)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_up)$name[V(graph_up)$name %notin% all_DE_EC_up], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[5][[1]]
# module_membership_median_up <- median(
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#   abs(rowSums(module_membership_norm_up))
# )
# module_membership_lower_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[2][[1]]
# membership_up <-
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#     abs(rowSums(module_membership_norm_up))
# 
# vertex_frame_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         ECvuln_DEup_up == "yes" ~ "white",
#       ECvuln_DEup_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         ECvuln_DEup_up == "yes" ~ "white",
#       ECvuln_DEup_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_up)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-ECup-up.png"),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_up,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_up),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_up)$weight ^ 2 / 4,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_up,
#     edge.color = ifelse(direction_up == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       ECvuln_DEup_up == "yes" & DE_opp_up == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_up$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_up$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# all_DE_match <- as.data.frame(
#   table(V(graph_up)$name[match(all_DE_EC_down, V(graph_up)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_up)$name[V(graph_up)$name %notin% all_DE_EC_down], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[5][[1]]
# module_membership_median_up <- median(
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#   abs(rowSums(module_membership_norm_up))
# )
# module_membership_lower_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[2][[1]]
# membership_up <-
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#     abs(rowSums(module_membership_norm_up))
# 
# vertex_frame_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         ECvuln_DEdown_up == "yes" ~ "white",
#       ECvuln_DEdown_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         ECvuln_DEdown_up == "yes" ~ "white",
#       ECvuln_DEdown_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# vertex_label <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_up)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-ECdown-up.png"),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_up,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_up),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_up)$weight ^ 2 / 4,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_up,
#     edge.color = ifelse(direction_up == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       ECvuln_DEdown_up == "yes" & DE_opp_up == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_up$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_up$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# all_DE_match <- as.data.frame(
#   table(V(graph_up)$name[match(all_DE_SFGvuln_up, V(graph_up)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_up)$name[V(graph_up)$name %notin% all_DE_SFGvuln_up], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[5][[1]]
# module_membership_median_up <- median(
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#   abs(rowSums(module_membership_norm_up))
# )
# module_membership_lower_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[2][[1]]
# membership_up <-
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#     abs(rowSums(module_membership_norm_up))
# 
# vertex_frame_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         SFGvuln_DEup_up == "yes" ~ "white",
#       SFGvuln_DEup_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         SFGvuln_DEup_up == "yes" ~ "white",
#       SFGvuln_DEup_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_up)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-SFGvulnup-up.png"),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_up,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_up),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_up)$weight ^ 2 / 4,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_up,
#     edge.color = ifelse(direction_up == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       SFGinvuln_DEup_up == "yes" & DE_opp_up == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_up$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_up$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# all_DE_match <- as.data.frame(
#   table(V(graph_up)$name[match(all_DE_SFGvuln_down, V(graph_up)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_up)$name[V(graph_up)$name %notin% all_DE_SFGvuln_down], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[5][[1]]
# module_membership_median_up <- median(
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#   abs(rowSums(module_membership_norm_up))
# )
# module_membership_lower_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[2][[1]]
# membership_up <-
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#     abs(rowSums(module_membership_norm_up))
# 
# vertex_frame_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         SFGvuln_DEdown_up == "yes" ~ "white",
#       SFGvuln_DEdown_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) & SFGvuln_DEdown_up == "yes" ~ "white",
#       SFGvuln_DEdown_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_up)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-SFGvulndown-up.png"),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_up,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_up),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_up)$weight ^ 2 / 4,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_up,
#     edge.color = ifelse(direction_up == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       SFGinvuln_DEdown_up == "yes" & DE_opp_up == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_up$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_up$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# all_DE_match <- as.data.frame(
#   table(V(graph_up)$name[match(all_DE_SFGinvuln_up, V(graph_up)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_up)$name[V(graph_up)$name %notin% all_DE_SFGinvuln_up], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[5][[1]]
# module_membership_median_up <- median(
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#   abs(rowSums(module_membership_norm_up))
# )
# module_membership_lower_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[2][[1]]
# membership_up <-
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#     abs(rowSums(module_membership_norm_up))
# 
# vertex_frame_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         SFGinvuln_DEup_up == "yes" ~ "white",
#       SFGinvuln_DEup_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         SFGinvuln_DEup_up == "yes" ~ "white",
#       SFGinvuln_DEup_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_up)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-SFGinvulnup-up.png"),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_up,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_up),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_up)$weight ^ 2 / 4,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_up,
#     edge.color = ifelse(direction_up == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       SFGinvuln_DEup_up == "yes" & DE_opp_up == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_up$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_up$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# all_DE_match <- as.data.frame(
#   table(V(graph_up)$name[match(all_DE_SFGinvuln_down, V(graph_up)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_up)$name[V(graph_up)$name %notin% all_DE_SFGinvuln_down], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[5][[1]]
# module_membership_median_up <- median(
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#   abs(rowSums(module_membership_norm_up))
# )
# module_membership_lower_quart_up <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#   )
# )[2][[1]]
# membership_up <-
#   (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#     abs(rowSums(module_membership_norm_up))
# 
# vertex_frame_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         SFGinvuln_DEdown_up == "yes" ~ "white",
#       SFGinvuln_DEdown_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
#       ) &
#         SFGinvuln_DEdown_up == "yes" ~ "white",
#       SFGinvuln_DEdown_up == "yes" & DE_opp_up == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_up = DE_up) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_up)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(
#     results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-SFGinvulndown-up.png"
#   ),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_up,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_up),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_up)$weight ^ 2 / 4,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_up,
#     edge.color = ifelse(direction_up == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       SFGinvuln_DEdown_up == "yes" & DE_opp_up == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_up$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_up$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-nolabel-up.png"),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_up,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_up),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_up)$weight ^ 2 / 4,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_up,
#     edge.color = ifelse(direction_up == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = 1,
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = NA
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_up$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_up$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()

all_DE_match <- as.data.frame(
  table(V(graph_up)$name[match(DE_alt, V(graph_up)$name)]), stringsAsFactors = FALSE
)
all_DE_match_tmp <- data.frame(Var1 = V(graph_up)$name[V(graph_up)$name %notin% DE_alt], Freq = 0)
all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]

module_membership_upper_quart_up <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
      abs(rowSums(module_membership_norm_up))
  )
)[5][[1]]
module_membership_median_up <- median(
  (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
  abs(rowSums(module_membership_norm_up))
)
module_membership_lower_quart_up <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
      abs(rowSums(module_membership_norm_up))
  )
)[2][[1]]
membership_up <-
  (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
    abs(rowSums(module_membership_norm_up))

vertex_frame_color <- data.frame(DE_up = DE_up) %>%
  mutate(
    type = case_when(
      (
        DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
      ) & DE_opp_up == "no" ~ "white",
      TRUE ~ "darkgray"
    )
  )
vertex_frame_color <- vertex_frame_color$type

vertex_label_color <- data.frame(DE_up = DE_up) %>%
  mutate(
    type = case_when(
      (
        DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
      ) & DE_opp_up == "no" ~ "white",
      TRUE ~ "darkgray"
    )
  )
vertex_label_color <- vertex_label_color$type

vertex_label_font <- data.frame(DE_up = DE_up) %>%
  mutate(
    type = case_when(
      TRUE ~ 2
    )
  )
vertex_label_font <- vertex_label_font$type

vertex_label <- data.frame(DE_up = DE_up) %>%
  mutate(
    type = case_when(
      TRUE ~ V(graph_up)$name
    )
  )
vertex_label <- vertex_label$type
vertex_label[vertex_label == "NA"] <- NA

png(
  file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-alt-up.png"), 150,
  150, "in",
  res = 72
)
par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
plot(
    graph_up,
    rescale = FALSE,
    xlim = c(-1, 1),
    ylim = c(-1, 1),
    layout = norm_coords(layout_up),
    vertex.size = sqrt(
      (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
      abs(rowSums(module_membership_norm_up))
    ) * 5,
    vertex.label.color = vertex_label_color,
    edge.width = E(graph_up)$weight ^ 2 / 4,
    vertex.shape = "pie",
    vertex.pie = module_membership_list_up,
    edge.color = ifelse(direction_up == "up", "firebrick1", "dodgerblue"),
    vertex.label.family = "sans",
    vertex.label.font = vertex_label_font,
    vertex.label.cex = ifelse(
      (
        DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
      ) &
        DE_opp_up == "no",
      (all_DE_match$Freq + 3) * 2, 3
    ),
    vertex.frame.color = vertex_frame_color,
    vertex.label = vertex_label
)
legend(
    "right",
    legend = paste0(
      "Module ", unique(hub_gene_data_up$module),
      ": ", gsub("\n", " ", top_words[unique(hub_gene_data_up$module)])
    ),
    pch = 20,
    col = module_network_colors_up,
    pt.cex = 12,
    cex = 8,
    bty = "n",
    text.col = "white",
    inset = c(-0.5, 0)
)
dev.off()

all_DE_match <- as.data.frame(
  table(V(graph_up)$name[match(DE_alt, V(graph_up)$name)]), stringsAsFactors = FALSE
)
all_DE_match_tmp <- data.frame(Var1 = V(graph_up)$name[V(graph_up)$name %notin% DE_alt], Freq = 0)
all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]

module_membership_upper_quart_up <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
      abs(rowSums(module_membership_norm_up))
  )
)[5][[1]]
module_membership_median_up <- median(
  (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
  abs(rowSums(module_membership_norm_up))
)
module_membership_lower_quart_up <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
      abs(rowSums(module_membership_norm_up))
  )
)[2][[1]]
membership_up <-
  (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
    abs(rowSums(module_membership_norm_up))

vertex_frame_color <- data.frame(DE_up = DE_up) %>%
  mutate(
    type = case_when(
      (
        DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
      ) &
        target_all_up == "yes" ~ "white",
      target_all_up == "yes" ~ "darkgray",
      TRUE ~ "black"
    )
  )
vertex_frame_color <- vertex_frame_color$type

vertex_label_color <- data.frame(DE_up = DE_up) %>%
  mutate(
    type = case_when(
      (
        DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"
      ) &
        target_all_up == "yes" ~ "white",
      target_all_up == "yes" ~ "darkgray"
    )
  )
vertex_label_color <- vertex_label_color$type

vertex_label_font <- data.frame(DE_up = DE_up) %>%
  mutate(
    type = case_when(
      TRUE ~ 2
    )
  )
vertex_label_font <- vertex_label_font$type

vertex_label <- data.frame(DE_up = DE_up) %>%
  mutate(
    type = case_when(
      DE_opp_up == "yes" ~ "NA",
      target_all_up == "yes" ~ V(graph_up)$name
    )
  )
vertex_label <- vertex_label$type
vertex_label[vertex_label == "NA"] <- NA

# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-up-alt2.png"), 150,
#   150, "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_up,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_up),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
#       abs(rowSums(module_membership_norm_up))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_up)$weight ^ 2 / 4,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_up,
#     edge.color = ifelse(direction_up == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = 5,
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_up$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_up$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()

png(
  file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-up-alt3.png"), 150,
  150, "in",
  res = 72
)
par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
plot(
    graph_up,
    rescale = FALSE,
    xlim = c(-1, 1),
    ylim = c(-1, 1),
    layout = norm_coords(layout_up),
    vertex.size = sqrt(
      (abs(rowSums(module_membership_norm_up)) / hub_gene_all_GO_up) *
      abs(rowSums(module_membership_norm_up))
    ) * 5,
    vertex.label.color = vertex_label_color,
    edge.width = E(graph_up)$weight ^ 2 / 4,
    vertex.shape = "pie",
    vertex.pie = module_membership_list_up,
    edge.color = ifelse(direction_up == "up", "firebrick1", "dodgerblue"),
    vertex.label.family = "sans",
    vertex.label.font = vertex_label_font,
    vertex.label.cex = ifelse(
      (DE_up == "yes" | DE_strict_up == "yes" | DE_stricter_up == "yes" | DE_strictest_up == "yes"),
      (all_DE_match$Freq + 3) * 2, 3
    ),
    vertex.frame.color = vertex_frame_color,
    vertex.label = vertex_label
)
legend(
    "right",
    legend = paste0(
      "Module ", unique(hub_gene_data_up$module),
      ": ", gsub("\n", " ", top_words[unique(hub_gene_data_up$module)])
    ),
    pch = 20,
    col = module_network_colors_up,
    pt.cex = 12,
    cex = 8,
    bty = "n",
    text.col = "white",
    inset = c(-0.5, 0)
)
dev.off()

# all_DE_match <- as.data.frame(
#   table(V(graph_down)$name[match(all_DE_EC_up, V(graph_down)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_down)$name[V(graph_down)$name %notin% all_DE_EC_up], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[5][[1]]
# module_membership_median_down <- median(
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#   abs(rowSums(module_membership_norm_down))
# )
# module_membership_lower_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[2][[1]]
# membership_down <-
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#     abs(rowSums(module_membership_norm_down))
# 
# vertex_frame_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         ECvuln_DEup_down == "yes" ~ "white",
#       ECvuln_DEup_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         ECvuln_DEup_down == "yes" ~ "white",
#       ECvuln_DEup_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_down)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-ECup-down.png"), 150,
#   150, "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_down,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_down),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#     ) * 7,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_down)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_down,
#     edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       ECvuln_DEup_down == "yes" & DE_opp_down == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_down$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# all_DE_match <- as.data.frame(
#   table(V(graph_down)$name[match(all_DE_EC_down, V(graph_down)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_down)$name[V(graph_down)$name %notin% all_DE_EC_down], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[5][[1]]
# module_membership_median_down <- median(
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#   abs(rowSums(module_membership_norm_down))
# )
# module_membership_lower_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[2][[1]]
# membership_down <-
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#     abs(rowSums(module_membership_norm_down))
# 
# vertex_frame_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         ECvuln_DEdown_down == "yes" ~ "white",
#       ECvuln_DEdown_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         ECvuln_DEdown_down == "yes" ~ "white",
#       ECvuln_DEdown_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_down)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-ECdown-down.png"),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_down,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_down),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#     ) * 7,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_down)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_down,
#     edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       ECvuln_DEdown_down == "yes" & DE_opp_down == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_down$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# all_DE_match <- as.data.frame(
#   table(V(graph_down)$name[match(all_DE_SFGvuln_up, V(graph_down)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_down)$name[V(graph_down)$name %notin% all_DE_SFGvuln_up], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[5][[1]]
# module_membership_median_down <- median(
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#   abs(rowSums(module_membership_norm_down))
# )
# module_membership_lower_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[2][[1]]
# membership_down <-
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#     abs(rowSums(module_membership_norm_down))
# 
# vertex_frame_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         SFGvuln_DEup_down == "yes" ~ "white",
#       SFGvuln_DEup_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         SFGvuln_DEup_down == "yes" ~ "white",
#       SFGvuln_DEup_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_down)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-SFGvulnup-down.png"),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_down,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_down),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#     ) * 7,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_down)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_down,
#     edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       SFGinvuln_DEup_down == "yes" & DE_opp_down == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_down$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# all_DE_match <- as.data.frame(
#   table(V(graph_down)$name[match(all_DE_SFGvuln_down, V(graph_down)$name)]),
#   stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_down)$name[V(graph_down)$name %notin% all_DE_SFGvuln_down], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[5][[1]]
# module_membership_median_down <- median(
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#   abs(rowSums(module_membership_norm_down))
# )
# module_membership_lower_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[2][[1]]
# membership_down <-
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#     abs(rowSums(module_membership_norm_down))
# 
# vertex_frame_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         SFGvuln_DEdown_down == "yes" ~ "white",
#       SFGvuln_DEdown_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         SFGvuln_DEdown_down == "yes" ~ "white",
#       SFGvuln_DEdown_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_down)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(
#     results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-SFGvulndown-down.png"
#   ),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_down,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_down),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#     ) * 7,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_down)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_down,
#     edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       SFGinvuln_DEdown_down == "yes" & DE_opp_down == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_down$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# all_DE_match <- as.data.frame(
#   table(V(graph_down)$name[match(all_DE_SFGinvuln_up, V(graph_down)$name)]),
#   stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_down)$name[V(graph_down)$name %notin% all_DE_SFGinvuln_up], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[5][[1]]
# module_membership_median_down <- median(
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#   abs(rowSums(module_membership_norm_down))
# )
# module_membership_lower_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[2][[1]]
# membership_down <-
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#     abs(rowSums(module_membership_norm_down))
# 
# vertex_frame_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         SFGinvuln_DEup_down == "yes" ~ "white",
#       SFGinvuln_DEup_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         SFGinvuln_DEup_down == "yes" ~ "white",
#       SFGinvuln_DEup_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_down)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(
#     results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-SFGinvulnup-down.png"
#   ),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_down,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_down),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#     ) * 7,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_down)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_down,
#     edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       SFGinvuln_DEup_down == "yes" & DE_opp_down == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_down$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# all_DE_match <- as.data.frame(
#   table(V(graph_down)$name[match(all_DE_SFGinvuln_down, V(graph_down)$name)]),
#   stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_down)$name[V(graph_down)$name %notin% all_DE_SFGinvuln_down], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[5][[1]]
# module_membership_median_down <- median(
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#   abs(rowSums(module_membership_norm_down))
# )
# module_membership_lower_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[2][[1]]
# membership_down <-
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#     abs(rowSums(module_membership_norm_down))
# 
# vertex_frame_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         SFGinvuln_DEdown_down == "yes" ~ "white",
#       SFGinvuln_DEdown_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       (
#         DE_down == "yes" | DE_strict_down == "yes" |
#           DE_stricter_down == "yes" | DE_strictest_down == "yes"
#       ) &
#         SFGinvuln_DEdown_down == "yes" ~ "white",
#       SFGinvuln_DEdown_down == "yes" & DE_opp_down == "no" ~ "darkgray",
#       TRUE ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ V(graph_down)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(
#     results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-SFGinvulndown-down.png"
#   ),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_down,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_down),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#     ) * 7,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_down)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_down,
#     edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       SFGinvuln_DEdown_down == "yes" & DE_opp_down == "no", (all_DE_match$Freq + 3) * 2, 3
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_down$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-down-nolabel.png"),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_down,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_down),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#     ) * 7,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_down)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_down,
#     edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = 1,
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = NA
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_down$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()

all_DE_match <- as.data.frame(
  table(V(graph_down)$name[match(DE_alt, V(graph_down)$name)]), stringsAsFactors = FALSE
)
all_DE_match_tmp <- data.frame(
  Var1 = V(graph_down)$name[V(graph_down)$name %notin% DE_alt], Freq = 0
)
all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]

module_membership_upper_quart_down <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
      abs(rowSums(module_membership_norm_down))
  )
)[5][[1]]
module_membership_median_down <- median(
  (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
  abs(rowSums(module_membership_norm_down))
)
module_membership_lower_quart_down <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
      abs(rowSums(module_membership_norm_down))
  )
)[2][[1]]
membership_down <-
  (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
    abs(rowSums(module_membership_norm_down))

vertex_frame_color <- data.frame(DE_down = DE_down) %>%
  mutate(
    type = case_when(
      (
        DE_down == "yes" | DE_strict_down == "yes" |
          DE_stricter_down == "yes" | DE_strictest_down == "yes"
      ) &
        DE_opp_down == "no" ~ "white",
      TRUE ~ "darkgray"
    )
  )
vertex_frame_color <- vertex_frame_color$type

vertex_label_color <- data.frame(DE_down = DE_down) %>%
  mutate(
    type = case_when(
      (
        DE_down == "yes" | DE_strict_down == "yes" |
          DE_stricter_down == "yes" | DE_strictest_down == "yes"
      ) &
        DE_opp_down == "no" ~ "white",
      TRUE ~ "darkgray"
    )
  )
vertex_label_color <- vertex_label_color$type

vertex_label_font <- data.frame(DE_down = DE_down) %>%
  mutate(
    type = case_when(
      TRUE ~ 2
    )
  )
vertex_label_font <- vertex_label_font$type

vertex_label <- data.frame(DE_down = DE_down) %>%
  mutate(
    type = case_when(
      TRUE ~ V(graph_down)$name
    )
  )
vertex_label <- vertex_label$type
vertex_label[vertex_label == "NA"] <- NA

png(
  file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-down-alt.png"), 150,
  150, "in",
  res = 72
)
par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
plot(
    graph_down,
    rescale = FALSE,
    xlim = c(-1, 1),
    ylim = c(-1, 1),
    layout = norm_coords(layout_down),
    vertex.size = sqrt(
      (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
      abs(rowSums(module_membership_norm_down))
    ) * 7,
    vertex.label.color = vertex_label_color,
    edge.width = E(graph_down)$weight ^ 2 / 2,
    vertex.shape = "pie",
    vertex.pie = module_membership_list_down,
    edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
    vertex.label.family = "sans",
    vertex.label.font = vertex_label_font,
    vertex.label.cex = ifelse(
      (
        DE_down == "yes" | DE_strict_down == "yes" |
          DE_stricter_down == "yes" | DE_strictest_down == "yes"
      ) &
        DE_opp_down == "no",
      (all_DE_match$Freq + 3) * 2, 3
    ),
    vertex.frame.color = vertex_frame_color,
    vertex.label = vertex_label
)
legend(
    "right",
    legend = paste0(
      "Module ", unique(hub_gene_data_down$module),
      ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
    ),
    pch = 20,
    col = module_network_colors_down,
    pt.cex = 12,
    cex = 8,
    bty = "n",
    text.col = "white",
    inset = c(-0.5, 0)
)
dev.off()

all_DE_match <- as.data.frame(
  table(V(graph_down)$name[match(DE_alt, V(graph_down)$name)]), stringsAsFactors = FALSE
)
all_DE_match_tmp <- data.frame(
  Var1 = V(graph_down)$name[V(graph_down)$name %notin% DE_alt], Freq = 0
)
all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]

module_membership_upper_quart_down <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
      abs(rowSums(module_membership_norm_down))
  )
)[5][[1]]
module_membership_median_down <- median(
  (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
  abs(rowSums(module_membership_norm_down))
)
module_membership_lower_quart_down <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
      abs(rowSums(module_membership_norm_down))
  )
)[2][[1]]
membership_down <-
  (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
    abs(rowSums(module_membership_norm_down))

vertex_frame_color <- data.frame(DE_down = DE_down) %>%
  mutate(
    type = case_when(
      (
        DE_down == "yes" | DE_strict_down == "yes" |
          DE_stricter_down == "yes" | DE_strictest_down == "yes"
      ) &
        target_all_down == "yes" ~ "white",
      target_all_down == "yes" ~ "darkgray",
      TRUE ~ "black"
    )
  )
vertex_frame_color <- vertex_frame_color$type

vertex_label_color <- data.frame(DE_down = DE_down) %>%
  mutate(
    type = case_when(
      (
        DE_down == "yes" | DE_strict_down == "yes" |
          DE_stricter_down == "yes" | DE_strictest_down == "yes"
      ) &
        target_all_down == "yes" ~ "white",
      target_all_down == "yes" ~ "darkgray"
    )
  )
vertex_label_color <- vertex_label_color$type

vertex_label_font <- data.frame(DE_down = DE_down) %>%
  mutate(
    type = case_when(
      TRUE ~ 2
    )
  )
vertex_label_font <- vertex_label_font$type

vertex_label <- data.frame(DE_down = DE_down) %>%
  mutate(
    type = case_when(
      DE_opp_down == "yes" ~ "NA",
      target_all_down == "yes" ~ V(graph_down)$name
    )
  )
vertex_label <- vertex_label$type
vertex_label[vertex_label == "NA"] <- NA

# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-down-alt2.png"), 150,
#   150, "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_down,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_down),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#     ) * 7,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_down)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_down,
#     edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = 5,
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_down$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()

png(
  file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-down-alt3.png"), 150,
  150, "in",
  res = 72
)
par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
plot(
    graph_down,
    rescale = FALSE,
    xlim = c(-1, 1),
    ylim = c(-1, 1),
    layout = norm_coords(layout_down),
    vertex.size = sqrt(
      (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
      abs(rowSums(module_membership_norm_down))
    ) * 7,
    vertex.label.color = vertex_label_color,
    edge.width = E(graph_down)$weight ^ 2 / 2,
    vertex.shape = "pie",
    vertex.pie = module_membership_list_down,
    edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
    vertex.label.family = "sans",
    vertex.label.font = vertex_label_font,
    vertex.label.cex = ifelse(
      (
        DE_down == "yes" | DE_strict_down == "yes" |
          DE_stricter_down == "yes" | DE_strictest_down == "yes"
      ),
      (all_DE_match$Freq + 3) * 2, 3
    ),
    vertex.frame.color = vertex_frame_color,
    vertex.label = vertex_label
)
legend(
    "right",
    legend = paste0(
      "Module ", unique(hub_gene_data_down$module),
      ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
    ),
    pch = 20,
    col = module_network_colors_down,
    pt.cex = 12,
    cex = 8,
    bty = "n",
    text.col = "white",
    inset = c(-0.5, 0)
)
dev.off()

# all_DE_match <- as.data.frame(
#   table(V(graph_down)$name[match(DE_alt, V(graph_down)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(
#   Var1 = V(graph_down)$name[V(graph_down)$name %notin% DE_alt], Freq = 0
# )
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[5][[1]]
# module_membership_median_down <- median(
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#   abs(rowSums(module_membership_norm_down))
# )
# module_membership_lower_quart_down <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#   )
# )[2][[1]]
# membership_down <-
#   (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#     abs(rowSums(module_membership_norm_down))
# 
# vertex_frame_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       mapt_head_down == "yes" ~ "white",
#       mapt_tail_down == "yes" &
#         (
#           DE_down == "yes" | DE_strict_down == "yes" |
#             DE_stricter_down == "yes" | DE_strictest_down == "yes"
#         ) ~ "darkgray",
#       mapt_tail_down == "yes" ~ "dimgray",
#       TRUE ~ "black"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       mapt_head_down == "yes" ~ "white",
#       mapt_tail_down == "yes" &
#         (DE_down == "yes" | DE_strict_down == "yes" |
#            DE_stricter_down == "yes" | DE_strictest_down == "yes"
#         ) ~ "darkgray",
#       mapt_tail_down == "yes" ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE_down = DE_down) %>%
#   mutate(
#     type = case_when(
#       mapt_head_down == "yes" | mapt_tail_down == "yes" ~ V(graph_down)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-down-mapt.png"), 150,
#   150, "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph_down,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout_down),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm_down)) / hub_gene_all_GO_down) *
#       abs(rowSums(module_membership_norm_down))
#     ) * 7,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph_down)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list_down,
#     edge.color = ifelse(direction_down == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = ifelse(
#       mapt_tail_down == "yes", (all_DE_match$Freq + 3) * 2, max(all_DE_match$Freq + 4) * 2
#     ),
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data_down$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data_down$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()

all_DE_match <- as.data.frame(
  table(V(graph)$name[match(DE_alt, V(graph)$name)]), stringsAsFactors = FALSE
)
all_DE_match_tmp <- data.frame(Var1 = V(graph)$name[V(graph)$name %notin% DE_alt], Freq = 0)
all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]

module_membership_upper_quart <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
      abs(rowSums(module_membership_norm))
  )
)[5][[1]]
module_membership_median <- median(
  (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
  abs(rowSums(module_membership_norm))
)
module_membership_lower_quart <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
      abs(rowSums(module_membership_norm))
  )
)[2][[1]]
membership <-
  (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) * abs(rowSums(module_membership_norm))

vertex_frame_color <- data.frame(DE = DE) %>%
  mutate(
    type = case_when(
      DE == "yes" ~ "white",
      TRUE ~ "darkgray"
    )
  )
vertex_frame_color <- vertex_frame_color$type

vertex_label_color <- data.frame(DE = DE) %>%
  mutate(
    type = case_when(
      DE == "yes" ~ "white",
      TRUE ~ "darkgray"
    )
  )
vertex_label_color <- vertex_label_color$type

vertex_label_font <- data.frame(DE = DE) %>%
  mutate(
    type = case_when(
      TRUE ~ 2
    )
  )
vertex_label_font <- vertex_label_font$type

vertex_label <- data.frame(DE = DE) %>%
  mutate(
    type = case_when(
      TRUE ~ V(graph)$name
    )
  )
vertex_label <- vertex_label$type
vertex_label[vertex_label == "NA"] <- NA

png(
  file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-alt.png"), 150, 150,
  "in", res = 72
)
par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
plot(
    graph,
    rescale = FALSE,
    xlim = c(-1, 1),
    ylim = c(-1, 1),
    layout = norm_coords(layout),
    vertex.size = sqrt(
      (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
      abs(rowSums(module_membership_norm))
    ) * 5,
    vertex.label.color = vertex_label_color,
    edge.width = E(graph)$weight ^ 2 / 2,
    vertex.shape = "pie",
    vertex.pie = module_membership_list,
    edge.color = ifelse(direction == "up", "firebrick1", "dodgerblue"),
    vertex.label.family = "sans",
    vertex.label.font = vertex_label_font,
    vertex.label.cex = (all_DE_match$Freq + 1) * 3, 1,
    vertex.frame.color = vertex_frame_color,
    vertex.label = vertex_label
)
legend(
    "right",
    legend = paste0(
      "Module ", unique(hub_gene_data$module),
      ": ", gsub("\n", " ", top_words[unique(hub_gene_data$module)])
    ),
    pch = 20,
    col = module_network_colors,
    pt.cex = 12,
    cex = 8,
    bty = "n",
    text.col = "white",
    inset = c(-0.5, 0)
)
dev.off()

# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-nolabel.png"),
#   150,
#   150,
#   "in",
#   res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
#       abs(rowSums(module_membership_norm))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list,
#     edge.color = ifelse(direction == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = 1,
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = NA
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()

all_DE_match <- as.data.frame(
  table(V(graph)$name[match(DE_alt, V(graph)$name)]), stringsAsFactors = FALSE
)
all_DE_match_tmp <- data.frame(Var1 = V(graph)$name[V(graph)$name %notin% DE_alt], Freq = 0)
all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]

module_membership_upper_quart <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
      abs(rowSums(module_membership_norm))
  )
)[5][[1]]
module_membership_median <- median(
  (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
  abs(rowSums(module_membership_norm))
)
module_membership_lower_quart <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
      abs(rowSums(module_membership_norm))
  )
)[2][[1]]
membership <-
  (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) * abs(rowSums(module_membership_norm))

vertex_frame_color <- data.frame(DE = DE) %>%
  mutate(
    type = case_when(
      DE == "yes" & target_all == "yes" ~ "white",
      target_all == "yes" ~ "darkgray",
      TRUE ~ "black"
    )
  )
vertex_frame_color <- vertex_frame_color$type

vertex_label_color <- data.frame(DE = DE) %>%
  mutate(
    type = case_when(
      DE == "yes" & target_all == "yes" ~ "white",
      target_all == "yes" ~ "darkgray"
    )
  )
vertex_label_color <- vertex_label_color$type

vertex_label_font <- data.frame(DE = DE) %>%
  mutate(
    type = case_when(
      TRUE ~ 2
    )
  )
vertex_label_font <- vertex_label_font$type

vertex_label <- data.frame(DE = DE) %>%
  mutate(
    type = case_when(
      DE == "yes" & target_all == "yes" ~ V(graph)$name,
      target_all == "yes" ~ V(graph)$name
    )
  )
vertex_label <- vertex_label$type
vertex_label[vertex_label == "NA"] <- NA

# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-alt2.png"), 150, 150,
#   "in", res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
#       abs(rowSums(module_membership_norm))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list,
#     edge.color = ifelse(direction == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = 5,
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()

png(
  file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-alt3.png"), 150, 150,
  "in", res = 72
)
par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
plot(
    graph,
    rescale = FALSE,
    xlim = c(-1, 1),
    ylim = c(-1, 1),
    layout = norm_coords(layout),
    vertex.size = sqrt(
      (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
      abs(rowSums(module_membership_norm))
    ) * 5,
    vertex.label.color = vertex_label_color,
    edge.width = E(graph)$weight ^ 2 / 2,
    vertex.shape = "pie",
    vertex.pie = module_membership_list,
    edge.color = ifelse(direction == "up", "firebrick1", "dodgerblue"),
    vertex.label.family = "sans",
    vertex.label.font = vertex_label_font,
    vertex.label.cex = (all_DE_match$Freq + 1) * 3, 1,
    vertex.frame.color = vertex_frame_color,
    vertex.label = vertex_label
)
legend(
    "right",
    legend = paste0(
      "Module ", unique(hub_gene_data$module),
      ": ", gsub("\n", " ", top_words[unique(hub_gene_data$module)])
    ),
    pch = 20,
    col = module_network_colors,
    pt.cex = 12,
    cex = 8,
    bty = "n",
    text.col = "white",
    inset = c(-0.5, 0)
)
dev.off()

# all_DE_match <- as.data.frame(
#   table(V(graph)$name[match(DE_alt, V(graph)$name)]), stringsAsFactors = FALSE
# )
# all_DE_match_tmp <- data.frame(Var1 = V(graph)$name[V(graph)$name %notin% DE_alt], Freq = 0)
# all_DE_match <- rbind(all_DE_match, all_DE_match_tmp)
# all_DE_match <- all_DE_match[order(all_DE_match$Var1), ]
# 
# module_membership_upper_quart <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
#       abs(rowSums(module_membership_norm))
#   )
# )[5][[1]]
# module_membership_median <- median(
#   (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
#   abs(rowSums(module_membership_norm))
# )
# module_membership_lower_quart <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
#       abs(rowSums(module_membership_norm))
#   )
# )[2][[1]]
# membership <-
#   (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) * abs(rowSums(module_membership_norm))
# 
# vertex_frame_color <- data.frame(DE = DE) %>%
#   mutate(
#     type = case_when(
#       mapt_head == "yes" ~ "white",
#       DE == "yes" & mapt_tail == "yes" ~ "darkgray",
#       mapt_tail == "yes" ~ "dimgray",
#       TRUE ~ "black"
#     )
#   )
# vertex_frame_color <- vertex_frame_color$type
# 
# vertex_label_color <- data.frame(DE = DE) %>%
#   mutate(
#     type = case_when(
#       mapt_head == "yes" ~ "white",
#       DE == "yes" & mapt_tail == "yes" ~ "darkgray",
#       mapt_tail == "yes" ~ "dimgray"
#     )
#   )
# vertex_label_color <- vertex_label_color$type
# 
# vertex_label_font <- data.frame(DE = DE) %>%
#   mutate(
#     type = case_when(
#       TRUE ~ 2
#     )
#   )
# vertex_label_font <- vertex_label_font$type
# 
# vertex_label <- data.frame(DE = DE) %>%
#   mutate(
#     type = case_when(
#       mapt_head == "yes" | mapt_tail == "yes" ~ V(graph)$name
#     )
#   )
# vertex_label <- vertex_label$type
# vertex_label[vertex_label == "NA"] <- NA
# 
# png(
#   file.path(results_dir, data_name, "05-gsva", "04-Reln-Wfs1-Bok-Sst-network-mapt.png"), 150, 150,
#   "in", res = 72
# )
# par(bg = "black", mar = c(5.1, 4.1, 4.1, 250.1), xpd = TRUE)
# plot(
#     graph,
#     rescale = FALSE,
#     xlim = c(-1, 1),
#     ylim = c(-1, 1),
#     layout = norm_coords(layout),
#     vertex.size = sqrt(
#       (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
#       abs(rowSums(module_membership_norm))
#     ) * 5,
#     vertex.label.color = vertex_label_color,
#     edge.width = E(graph)$weight ^ 2 / 2,
#     vertex.shape = "pie",
#     vertex.pie = module_membership_list,
#     edge.color = ifelse(direction == "up", "firebrick1", "dodgerblue"),
#     vertex.label.family = "sans",
#     vertex.label.font = vertex_label_font,
#     vertex.label.cex = 18,
#     vertex.frame.color = vertex_frame_color,
#     vertex.label = vertex_label
# )
# legend(
#     "right",
#     legend = paste0(
#       "Module ", unique(hub_gene_data$module),
#       ": ", gsub("\n", " ", top_words[unique(hub_gene_data$module)])
#     ),
#     pch = 20,
#     col = module_network_colors_up,
#     pt.cex = 12,
#     cex = 8,
#     bty = "n",
#     text.col = "white",
#     inset = c(-0.5, 0)
# )
# dev.off()
```

## Expression of GO Genes

```{r}
anno <- HeatmapAnnotation(
  donor = seurat_symbols$external_donor_name_id,
  cell_type = anno_block(
    gp = gpar(fill = "lightgray", col = "lightgray"),
    labels = c(names(idents)[[1]], names(idents)[[2]], names(idents)[[3]], names(idents)[[4]]),
    labels_gp = gpar(col = "black", fontsize = 20)
  ),
  col = list(
    donor = c(
      "475" = anno_color[1], "476" = anno_color[2],
      "527" = anno_color[3], "528" = anno_color[4],
      "456" = anno_color[5], "471" = anno_color[6]
    )
  )
)
```

### `r names(idents)[1]` Upregulation

```{r, fig.height = 5, fig.width = 20}
# TODO: Fix problem of contents appearing out of order.
# -----------------------------------------------------
set <- 2 # Up or downregulation?
if (length(gsva_list_pseudobulk_all[[set]]) > 0) {
  exp_mat <- mat
  exp_mat <- exp_mat[rownames(exp_mat) %in% gsva_list_pseudobulk_all[[set]], ]
  sort_mat <- mat[unlist(row_order(heatmap)), ]
  sort_mat <- sort_mat[rownames(sort_mat) %in% rownames(exp_mat), ]
  exp_mat <- exp_mat[match(rownames(sort_mat), rownames(exp_mat)), ]
  for (i in 1:nrow(exp_mat)) {
    genes <- which(
      gene_sets[[rownames(exp_mat)[i]]]@geneIds %in% markers_list_pseudobulk_all[[set]]
    )
    genes <- gene_sets[[rownames(exp_mat)[i]]]@geneIds[genes]
    genes <- gene_anno[gene_anno$ensembl_gene_id %in% genes, ]
    genes <- genes$external_gene_name
    title <- paste0(
      gene_sets[[rownames(exp_mat)[i]]]@shortDescription, ": ",
      gene_sets[[rownames(exp_mat)[i]]]@setName
    )
    print(title)

    if (length(genes) > 0) {
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " DE genes: ",
          paste(genes, collapse = ", ")
        )
      )
      strict_genes <- genes[genes %in% rownames(markers_list_symbols_pseudobulk_strict[[1]])]
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " DE genes in strict model: ",
          paste(strict_genes, collapse = ", ")
        )
      )
      stricter_genes <- genes[genes %in% rownames(markers_list_symbols_pseudobulk_stricter[[1]])]
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " DE genes in stricter model: ",
          paste(stricter_genes, collapse = ", ")
        )
      )
      strictest_genes <- genes[genes %in% rownames(markers_list_symbols_pseudobulk_strictest[[1]])]
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " DE genes in strictest model: ",
          paste(strictest_genes, collapse = ", ")
        )
      )
      open_target_genes <- genes[genes %in% open_targets_all]
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " DE genes in Open Targets: ",
          paste(open_target_genes, collapse = ", ")
        )
      )
      row_colors <- rownames(as.matrix(GetAssayData(seurat_symbols)[genes, , drop = FALSE]))
      row_colors <- data.frame(
        row_colors = row_colors,
        DE_strict = row_colors %in% rownames(markers_list_symbols_pseudobulk_strict[[1]]),
        DE_stricter = row_colors %in% rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
        DE_strictest = row_colors %in% rownames(markers_list_symbols_pseudobulk_strictest[[1]])
      ) %>%
        mutate(
          type = case_when(
            DE_strictest == TRUE ~ "firebrick",
            DE_stricter == TRUE ~ "darkorange",
            DE_strict == TRUE ~ "darkgoldenrod",
            TRUE ~ "black"
          )
        )
      row_colors <- row_colors$type
      draw(
        Heatmap(
          t(
            apply(
              as.matrix(GetAssayData(seurat_symbols)[genes, , drop = FALSE]), 1,
              function (x) (x - min(x)) / (max(x) - min(x))
            )
          ),
          color3,
          cluster_columns = FALSE,
          show_column_names = FALSE,
          column_split = seurat_symbols$idents,
          show_heatmap_legend = FALSE,
          top_annotation = anno,
          column_title = title,
          column_title_gp = gpar(fontsize = 20),
          row_names_gp = gpar(fontsize = 18, col = unlist(row_colors)),
          column_gap = unit(2, "mm")
        )
      )
    }

    genes <- gene_sets[[rownames(exp_mat)[i]]]@geneIds
    genes <- gene_anno[gene_anno$ensembl_gene_id %in% genes, ]
    genes <- genes$external_gene_name
    if (length(genes) > 0) {
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription,
          " all genes: ",
          paste(genes, collapse = ", ")
        )
      )
      draw(
        Heatmap(
          t(
            apply(
              as.matrix(GetAssayData(seurat_symbols)[genes, , drop = FALSE]), 1,
              function (x) (x - min(x)) / (max(x) - min(x))
            )
          ),
          color3,
          cluster_columns = FALSE,
          show_column_names = FALSE,
          column_split = seurat_symbols$idents,
          show_heatmap_legend = FALSE,
          top_annotation = anno,
          column_title = title,
          column_title_gp = gpar(fontsize = 20),
          column_gap = unit(2, "mm")
        )
      )
    }
  }
}
# -----------------------------------------------------
```

### `r names(idents)[1]` Downregulation

```{r, fig.height = 5, fig.width = 20}
# TODO: Fix problem of contents appearing out of order.
# -----------------------------------------------------
set <- 3 # Up or downregulation?
if (length(gsva_list_pseudobulk_all[[set]]) > 0) {
  exp_mat <- mat
  exp_mat <- exp_mat[rownames(exp_mat) %in% gsva_list_pseudobulk_all[[set]], ]
  exp_mat <- exp_mat[rownames(exp_mat) %in% rownames(mat), ]
  sort_mat <- mat[unlist(row_order(heatmap)), ]
  sort_mat <- sort_mat[rownames(sort_mat) %in% rownames(exp_mat), ]
  exp_mat <- exp_mat[match(rownames(sort_mat), rownames(exp_mat)), ]
  for (i in 1:nrow(exp_mat)) {
    genes <- which(
      gene_sets[[rownames(exp_mat)[i]]]@geneIds %in% markers_list_pseudobulk_all[[set]]
    )
    genes <- gene_sets[[rownames(exp_mat)[i]]]@geneIds[genes]
    genes <- gene_anno[gene_anno$ensembl_gene_id %in% genes, ]
    genes <- genes$external_gene_name
    title <- paste0(
      gene_sets[[rownames(exp_mat)[i]]]@shortDescription, ": ",
      gene_sets[[rownames(exp_mat)[i]]]@setName
    )
    print(title)

    if (length(genes) > 0) {
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " DE genes: ",
          paste(genes, collapse = ", ")
        )
      )
      strict_genes <- genes[genes %in% rownames(markers_list_symbols_pseudobulk_strict[[1]])]
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " DE genes in strict model: ",
          paste(strict_genes, collapse = ", ")
        )
      )
      stricter_genes <- genes[genes %in% rownames(markers_list_symbols_pseudobulk_stricter[[1]])]
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " DE genes in stricter model: ",
          paste(stricter_genes, collapse = ", ")
        )
      )
      strictest_genes <- genes[genes %in% rownames(markers_list_symbols_pseudobulk_strictest[[1]])]
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " DE genes in strictest model: ",
          paste(strictest_genes, collapse = ", ")
        )
      )
      open_target_genes <- genes[genes %in% open_targets_all]
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " DE genes in Open Targets: ",
          paste(open_target_genes, collapse = ", ")
        )
      )
      row_colors <- rownames(as.matrix(GetAssayData(seurat_symbols)[genes, , drop = FALSE]))
      row_colors <- data.frame(
        row_colors = row_colors,
        DE_strict = row_colors %in% rownames(markers_list_symbols_pseudobulk_strict[[1]]),
        DE_stricter = row_colors %in% rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
        DE_strictest = row_colors %in% rownames(markers_list_symbols_pseudobulk_strictest[[1]])
      ) %>%
        mutate(
          type = case_when(
            DE_strictest == TRUE ~ "firebrick",
            DE_stricter == TRUE ~ "darkorange",
            DE_strict == TRUE ~ "darkgoldenrod",
            TRUE ~ "black"
          )
        )
      row_colors <- row_colors$type
      draw(
        Heatmap(
          t(
            apply(
              as.matrix(GetAssayData(seurat_symbols)[genes, , drop = FALSE]), 1,
              function (x) (x - min(x)) / (max(x) - min(x))
            )
          ),
          color3,
          cluster_columns = FALSE,
          show_column_names = FALSE,
          column_split = seurat_symbols$idents,
          show_heatmap_legend = FALSE,
          top_annotation = anno,
          column_title = title,
          column_title_gp = gpar(fontsize = 20),
          row_names_gp = gpar(fontsize = 18, col = unlist(row_colors)),
          column_gap = unit(2, "mm")
        )
      )
    }

    genes <- gene_sets[[rownames(exp_mat)[i]]]@geneIds
    genes <- gene_anno[gene_anno$ensembl_gene_id %in% genes, ]
    genes <- genes$external_gene_name
    if (length(genes) > 0) {
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " all genes: ",
          paste(genes, collapse = ", ")
        )
      )
      open_target_genes <- genes[genes %in% open_targets_all]
      print(
        paste0(
          gene_sets[[rownames(exp_mat)[i]]]@shortDescription, " all genes in Open Targets: ",
          paste(open_target_genes, collapse = ", ")
        )
      )
      draw(
        Heatmap(
          t(
            apply(
              as.matrix(GetAssayData(seurat_symbols)[genes, , drop = FALSE]), 1,
              function (x) (x - min(x)) / (max(x) - min(x))
            )
          ),
          color3,
          cluster_columns = FALSE,
          show_column_names = FALSE,
          column_split = seurat_symbols$idents,
          show_heatmap_legend = FALSE,
          top_annotation = anno,
          column_title = title,
          column_title_gp = gpar(fontsize = 20),
          column_gap = unit(2, "mm")
        )
      )
    }
  }
}
# -----------------------------------------------------
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
