---
title: '02 Subsets - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "02-subsets.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).

In this particular document we create useful subsets of the original data.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

#    This file is part of tau-vulnerability.
#    Copyright (C) 2019  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# ------------------------------------------

packages <- c(
  "conflicted", "Seurat", "dplyr", "scater", "ggplot2", "ggrepel", "viridis", "DT", "data.table"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 2
protocol_type <- "smart-seq"
download_name <- "transcrip.tome" # Name of download file corresponding to dataset.

knitr::opts_chunk$set(fig.width = 15, fig.height = 13)
options(stringsAsFactors = FALSE)
# plan("multiprocess") # Parallel processing for Seurat. # TODO: Get `future` working for big data.

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
data_name_stem <- unlist(strsplit(data_name, "-"))[1] # `data_name` up to the first hyphen.
data_dir <- file.path(getwd(), "..", "..", "data", data_name) # Publicly available data.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Misc binaries and temp files.
results_dir <- file.path(getwd(), "..", "..", "results", data_name)

# Unique cache and results directory for each analysis number.
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}
if (!dir.exists(file.path(results_dir, "data", paste0("0", analysis_no)))) {
  dir.create(file.path(results_dir, "data", paste0("0", analysis_no)), recursive = TRUE)
}
```

We use data saved from previous analyses.

```{r}
seurat_orig <- readRDS(file.path(results_dir, "data", paste0("0", analysis_no - 1), "seurat.rds"))
seurat_orig
```

# MTG

```{r}
sub_name <- "MTG"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat_orig, region_label == sub_name)

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 30 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 30 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
MTG_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

MTG_markers <- markers
```

## Neuronal

```{r}
sub_name <- "MTG_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat, class_label == c("GABAergic", "Glutamatergic"))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 25 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 25 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
MTG_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

MTG_neuronal_markers <- markers
```

### EX

```{r}
sub_name <- "MTG_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(MTG_neuronal_seurat, class_label == "Glutamatergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 20 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 20 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
MTG_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

MTG_neuronal_EX_markers <- markers
```

### IN

```{r}
sub_name <- "MTG_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(MTG_neuronal_seurat, class_label == "GABAergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 20 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 20 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
MTG_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

MTG_neuronal_IN_markers <- markers
```

# V1C

```{r}
sub_name <- "V1C"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat_orig, region_label == sub_name)

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 25 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 25 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
V1C_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

V1C_markers <- markers
```

## Neuronal

```{r}
sub_name <- "V1C_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat, class_label == c("GABAergic", "Glutamatergic"))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 20 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 20 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
V1C_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

V1C_neuronal_markers <- markers
```

### EX

```{r}
sub_name <- "V1C_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(V1C_neuronal_seurat, class_label == "Glutamatergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
V1C_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

V1C_neuronal_EX_markers <- markers
```

### IN

```{r}
sub_name <- "V1C_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(V1C_neuronal_seurat, class_label == "GABAergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
V1C_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

V1C_neuronal_IN_markers <- markers
```

# CgG

```{r}
sub_name <- "CgG"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat_orig, region_label == sub_name)

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 20 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 20 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
CgG_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

CgG_markers <- markers
```

## Neuronal

```{r}
sub_name <- "CgG_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat, class_label == c("GABAergic", "Glutamatergic"))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 20 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 20 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
CgG_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

CgG_neuronal_markers <- markers
```

### EX

```{r}
sub_name <- "CgG_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(CgG_neuronal_seurat, class_label == "Glutamatergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
CgG_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

CgG_neuronal_EX_markers <- markers
```

### IN

```{r}
sub_name <- "CgG_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(CgG_neuronal_seurat, class_label == "GABAergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
CgG_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

CgG_neuronal_IN_markers <- markers
```

# M1lm

```{r}
sub_name <- "M1lm"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat_orig, region_label == sub_name)

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
M1lm_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

M1lm_markers <- markers
```

## Neuronal

```{r}
sub_name <- "M1lm_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat, class_label == c("GABAergic", "Glutamatergic"))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
M1lm_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

M1lm_neuronal_markers <- markers
```

### EX

```{r}
sub_name <- "M1lm_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(M1lm_neuronal_seurat, class_label == "Glutamatergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
M1lm_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

M1lm_neuronal_EX_markers <- markers
```

### IN

```{r}
sub_name <- "M1lm_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(M1lm_neuronal_seurat, class_label == "GABAergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 10 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 10 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
M1lm_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

M1lm_neuronal_IN_markers <- markers
```

# S1ul

```{r}
sub_name <- "S1ul"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat_orig, region_label == sub_name)

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
S1ul_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

S1ul_markers <- markers
```

## Neuronal

```{r}
sub_name <- "S1ul_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat, class_label == c("GABAergic", "Glutamatergic"))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
S1ul_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

S1ul_neuronal_markers <- markers
```

### EX

```{r}
sub_name <- "S1ul_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(S1ul_neuronal_seurat, class_label == "Glutamatergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
S1ul_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

S1ul_neuronal_EX_markers <- markers
```

### IN

```{r}
sub_name <- "S1ul_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(S1ul_neuronal_seurat, class_label == "GABAergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
S1ul_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

S1ul_neuronal_IN_markers <- markers
```

# S1lm

```{r}
sub_name <- "S1lm"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat_orig, region_label == sub_name)

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
S1lm_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

S1lm_markers <- markers
```

## Neuronal

```{r}
sub_name <- "S1lm_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat, class_label == c("GABAergic", "Glutamatergic"))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 10 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 10 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
S1lm_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

S1lm_neuronal_markers <- markers
```

### EX

```{r}
sub_name <- "S1lm_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(S1lm_neuronal_seurat, class_label == "Glutamatergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
S1lm_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

S1lm_neuronal_EX_markers <- markers
```

### IN

```{r}
sub_name <- "S1lm_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(S1lm_neuronal_seurat, class_label == "GABAergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 10 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 10 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
S1lm_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

S1lm_neuronal_IN_markers <- markers
```

# M1ul

```{r}
sub_name <- "M1ul"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat_orig, region_label == sub_name)

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
M1ul_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

M1ul_markers <- markers
```

## Neuronal

```{r}
sub_name <- "M1ul_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat, class_label == c("GABAergic", "Glutamatergic"))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 10 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 10 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
M1ul_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

M1ul_neuronal_markers <- markers
```

### EX

```{r}
sub_name <- "M1ul_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(M1ul_neuronal_seurat, class_label == "Glutamatergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 10 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 10 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
M1ul_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

M1ul_neuronal_EX_markers <- markers
```

### IN

```{r}
sub_name <- "M1ul_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(M1ul_neuronal_seurat, class_label == "GABAergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 10 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 10 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat")
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
M1ul_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

M1ul_neuronal_IN_markers <- markers
```

# A1C

```{r}
sub_name <- "A1C"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat_orig, region_label == sub_name)

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 20 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 20 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
A1C_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

A1C_markers <- markers
```

## Neuronal

```{r}
sub_name <- "A1C_neuronal"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(seurat, class_label == c("GABAergic", "Glutamatergic"))

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 20 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 20 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
A1C_neuronal_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

A1C_neuronal_markers <- markers
```

### EX

```{r}
sub_name <- "A1C_neuronal_EX"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(A1C_neuronal_seurat, class_label == "Glutamatergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 15 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 15 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
A1C_neuronal_EX_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

A1C_neuronal_EX_markers <- markers
```

### IN

```{r}
sub_name <- "A1C_neuronal_IN"

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_seurat.rds"))
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- subset(A1C_neuronal_seurat, class_label == "GABAergic")

  sce <- as.SingleCellExperiment(seurat)
  ribo_genes <- read.table(
    file.path(getwd(), "..", "..", "hugo-ribo-genes.txt"), sep = "\t", header = TRUE
  )
  is_ribo <- which(rowData(sce)$external_gene_name %in% ribo_genes$Approved.symbol)
  is_mito <- which(rowData(sce)$chromosome_name == "MT")
  sce <- addPerCellQC(
    sce, subsets = list(ribo = is_ribo, mito = is_mito), BPPARAM = MulticoreParam()
  )
  sce <- addPerFeatureQC(sce, BPPARAM = MulticoreParam())

  seurat <- as.Seurat(sce, data = NULL)
  seurat <- NormalizeData(seurat, verbose = FALSE)
  seurat <- FindVariableFeatures(seurat, verbose = FALSE)
  seurat <- ScaleData(
    seurat, vars.to.regress = c("subsets_mito_percent", "subsets_ribo_percent"), verbose = FALSE
  )

  seurat <- RunPCA(seurat, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "pca")[ , 1:2])
  names(add_df) <- paste0("pca", seq(ncol(add_df)))
  seurat$pca1 <- add_df$pca1
  seurat$pca2 <- add_df$pca2
  elbow <- 10 # TODO: Automate this.

  seurat <- RunUMAP(seurat, dims = 1:elbow, min.dist = 0.75, verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "umap"))
  names(add_df) <- paste0("umap", seq(ncol(add_df)))
  seurat$umap1 <- add_df$umap1
  seurat$umap2 <- add_df$umap2

  resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
  seurat <- FindNeighbors(seurat, dims = 1:elbow, verbose = FALSE)
  seurat <- FindClusters(seurat, resolution = resolution, verbose = FALSE)

  rm(clusters, ribo_genes, sce)
  saveRDS(seurat, rds)
}

print(paste0(sub_name, " Seurat dims: ", dim(seurat))) # TODO: Improve formatting.

ElbowPlot(seurat, ndims = 50)
elbow <- 10 # TODO: Automate this.
print(paste0("Retain first ", elbow, " PCs."))

red_dim_plot(seurat, "pca1", "pca2", "batch")
VizDimLoadings(seurat, dims = 1:2)
DimHeatmap(seurat, dims = 1:elbow, ncol = 3) # TODO: Customize color.

red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "batch")
red_dim_plot(seurat, "umap1", "umap2", "sum", "cont")

resolution <- (dim(seurat)[2] / 3000) * 0.8 # Default is optimal for 3K cells so we scale it.
print(paste0("Seurat resolution is: ", resolution))
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

# TODO: Reduce hardcoding in rest of chunk.
# -----------------------------------------

DefaultAssay(seurat) <- "RNA" # It is currently recommended to use the `RNA` assay for DEGs.
A1C_neuronal_IN_seurat <- seurat

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), paste0(sub_name, "_markers.rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

A1C_neuronal_IN_markers <- markers
```

# References

```{r}
saveRDS(MTG_seurat, file.path(results_dir, "data", paste0("0", analysis_no), "MTG_seurat.rds"))
saveRDS(MTG_markers, file.path(results_dir, "data", paste0("0", analysis_no), "MTG_markers.rds"))
saveRDS(
  MTG_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "MTG_neuronal_seurat.rds")
)
saveRDS(
  MTG_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "MTG_neuronal_markers.rds")
)
saveRDS(
  MTG_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "MTG_neuronal_EX_seurat.rds")
)
saveRDS(
  MTG_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "MTG_neuronal_EX_markers.rds")
)
saveRDS(
  MTG_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "MTG_neuronal_IN_seurat.rds")
)
saveRDS(
  MTG_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "MTG_neuronal_IN_markers.rds")
)

saveRDS(V1C_seurat, file.path(results_dir, "data", paste0("0", analysis_no), "V1C_seurat.rds"))
saveRDS(V1C_markers, file.path(results_dir, "data", paste0("0", analysis_no), "V1C_markers.rds"))
saveRDS(
  V1C_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "V1C_neuronal_seurat.rds")
)
saveRDS(
  V1C_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "V1C_neuronal_markers.rds")
)
saveRDS(
  V1C_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "V1C_neuronal_EX_seurat.rds")
)
saveRDS(
  V1C_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "V1C_neuronal_EX_markers.rds")
)
saveRDS(
  V1C_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "V1C_neuronal_IN_seurat.rds")
)
saveRDS(
  V1C_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "V1C_neuronal_IN_markers.rds")
)

saveRDS(CgG_seurat, file.path(results_dir, "data", paste0("0", analysis_no), "CgG_seurat.rds"))
saveRDS(CgG_markers, file.path(results_dir, "data", paste0("0", analysis_no), "CgG_markers.rds"))
saveRDS(
  CgG_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "CgG_neuronal_seurat.rds")
)
saveRDS(
  CgG_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "CgG_neuronal_markers.rds")
)
saveRDS(
  CgG_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "CgG_neuronal_EX_seurat.rds")
)
saveRDS(
  CgG_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "CgG_neuronal_EX_markers.rds")
)
saveRDS(
  CgG_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "CgG_neuronal_IN_seurat.rds")
)
saveRDS(
  CgG_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "CgG_neuronal_IN_markers.rds")
)

saveRDS(M1lm_seurat, file.path(results_dir, "data", paste0("0", analysis_no), "M1lm_seurat.rds"))
saveRDS(M1lm_markers, file.path(results_dir, "data", paste0("0", analysis_no), "M1lm_markers.rds"))
saveRDS(
  M1lm_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1lm_neuronal_seurat.rds")
)
saveRDS(
  M1lm_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1lm_neuronal_markers.rds")
)
saveRDS(
  M1lm_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1lm_neuronal_EX_seurat.rds")
)
saveRDS(
  M1lm_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1lm_neuronal_EX_markers.rds")
)
saveRDS(
  M1lm_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1lm_neuronal_IN_seurat.rds")
)
saveRDS(
  M1lm_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1lm_neuronal_IN_markers.rds")
)

saveRDS(S1ul_seurat, file.path(results_dir, "data", paste0("0", analysis_no), "S1ul_seurat.rds"))
saveRDS(S1ul_markers, file.path(results_dir, "data", paste0("0", analysis_no), "S1ul_markers.rds"))
saveRDS(
  S1ul_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1ul_neuronal_seurat.rds")
)
saveRDS(
  S1ul_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1ul_neuronal_markers.rds")
)
saveRDS(
  S1ul_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1ul_neuronal_EX_seurat.rds")
)
saveRDS(
  S1ul_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1ul_neuronal_EX_markers.rds")
)
saveRDS(
  S1ul_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1ul_neuronal_IN_seurat.rds")
)
saveRDS(
  S1ul_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1ul_neuronal_IN_markers.rds")
)

saveRDS(S1lm_seurat, file.path(results_dir, "data", paste0("0", analysis_no), "S1lm_seurat.rds"))
saveRDS(S1lm_markers, file.path(results_dir, "data", paste0("0", analysis_no), "S1lm_markers.rds"))
saveRDS(
  S1lm_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1lm_neuronal_seurat.rds")
)
saveRDS(
  S1lm_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1lm_neuronal_markers.rds")
)
saveRDS(
  S1lm_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1lm_neuronal_EX_seurat.rds")
)
saveRDS(
  S1lm_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1lm_neuronal_EX_markers.rds")
)
saveRDS(
  S1lm_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1lm_neuronal_IN_seurat.rds")
)
saveRDS(
  S1lm_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "S1lm_neuronal_IN_markers.rds")
)

saveRDS(M1ul_seurat, file.path(results_dir, "data", paste0("0", analysis_no), "M1ul_seurat.rds"))
saveRDS(M1ul_markers, file.path(results_dir, "data", paste0("0", analysis_no), "M1ul_markers.rds"))
saveRDS(
  M1ul_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1ul_neuronal_seurat.rds")
)
saveRDS(
  M1ul_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1ul_neuronal_markers.rds")
)
saveRDS(
  M1ul_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1ul_neuronal_EX_seurat.rds")
)
saveRDS(
  M1ul_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1ul_neuronal_EX_markers.rds")
)
saveRDS(
  M1ul_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1ul_neuronal_IN_seurat.rds")
)
saveRDS(
  M1ul_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "M1ul_neuronal_IN_markers.rds")
)

saveRDS(A1C_seurat, file.path(results_dir, "data", paste0("0", analysis_no), "A1C_seurat.rds"))
saveRDS(A1C_markers, file.path(results_dir, "data", paste0("0", analysis_no), "A1C_markers.rds"))
saveRDS(
  A1C_neuronal_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "A1C_neuronal_seurat.rds")
)
saveRDS(
  A1C_neuronal_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "A1C_neuronal_markers.rds")
)
saveRDS(
  A1C_neuronal_EX_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "A1C_neuronal_EX_seurat.rds")
)
saveRDS(
  A1C_neuronal_EX_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "A1C_neuronal_EX_markers.rds")
)
saveRDS(
  A1C_neuronal_IN_seurat,
  file.path(results_dir, "data", paste0("0", analysis_no), "A1C_neuronal_IN_seurat.rds")
)
saveRDS(
  A1C_neuronal_IN_markers,
  file.path(results_dir, "data", paste0("0", analysis_no), "A1C_neuronal_IN_markers.rds")
)

sessionInfo()
```
