---
title: '03 Vulnerable Types - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "03-vulnerable-types.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).

In this particular document we probe for vulnerable cell-types.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

#    This file is part of tau-vulnerability.
#    Copyright (C) 2019  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# ------------------------------------------

packages <- c("conflicted", "Seurat", "dplyr", "ggplot2", "ggrepel", "DT", "data.table")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 3
protocol_type <- "smart-seq"
download_name <- "transcrip.tome" # Name of download file corresponding to dataset.

options(stringsAsFactors = FALSE)

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
data_name_stem <- unlist(strsplit(data_name, "-"))[1] # `data_name` up to the first hyphen.
data_dir <- file.path(getwd(), "..", "..", "data", data_name) # Publicly available data.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Misc binaries and temp files.
results_dir <- file.path(getwd(), "..", "..", "results", data_name)

# Unique cache and results directory for each analysis number.
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}
if (!dir.exists(file.path(results_dir, "data", paste0("0", analysis_no)))) {
  dir.create(file.path(results_dir, "data", paste0("0", analysis_no)), recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 10)
```

We use data saved from previous analyses.

```{r}
MTG_neuronal_EX_seurat <- readRDS(
  file.path(results_dir, "data", paste0("0", analysis_no - 1), "MTG_neuronal_EX_seurat.rds")
)
MTG_neuronal_IN_seurat <- readRDS(
  file.path(results_dir, "data", paste0("0", analysis_no - 1), "MTG_neuronal_IN_seurat.rds")
)
```

# Datasets

## MTG Neuronal EX

```{r}
markers <- readRDS(
  file.path(results_dir, "data", paste0("0", analysis_no - 1), "MTG_neuronal_EX_markers.rds")
)
seurat <- MTG_neuronal_EX_seurat

red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

top <- markers %>% group_by(cluster) %>% top_n(n = 3, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:9], ncol = 3)
VlnPlot(seurat, features = top$gene[10:17], ncol = 3)

FeaturePlot(
  seurat, features = top$gene[1:9], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 3
)
FeaturePlot(
  seurat, features = top$gene[10:17], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 3
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 30, wt = avg_logFC))
```

## MTG Neuronal IN

```{r}
markers <- readRDS(
  file.path(results_dir, "data", paste0("0", analysis_no - 1), "MTG_neuronal_IN_markers.rds")
)
seurat <- MTG_neuronal_IN_seurat

red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat")

top <- markers %>% group_by(cluster) %>% top_n(n = 5, wt = avg_logFC)
DoHeatmap(seurat, features = top$gene, size = 3) + # TODO: Customize color.
  NoLegend()

top <- markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
datatable_custom(top)

VlnPlot(seurat, features = top$gene[1:6], ncol = 3)
VlnPlot(seurat, features = top$gene[7:10], ncol = 3)

FeaturePlot(
  seurat, features = top$gene[1:6], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 3
)
FeaturePlot(
  seurat, features = top$gene[7:10], order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 3
)

DotPlot(seurat, features = unique(top$gene), cols = c("blue", "red")) + RotatedAxis()

datatable_custom(markers %>% group_by(cluster) %>% top_n(n = 50, wt = avg_logFC))
```

# Throughout EX

```{r}
genes <- c("GRIN1", "SLC17A7")
seurat <- MTG_neuronal_EX_seurat

if (all(genes %in% rownames(seurat))) { # `blend` requires all features present.
  plots <- FeaturePlot(
    seurat, features = genes, order = TRUE,
    reduction = "umap", blend = TRUE, combine = FALSE
  )
  CombinePlots(plots[3:4], legend = "none")
} else {
  FeaturePlot(
    seurat, features = genes, order = TRUE, reduction = "umap",
    cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
  )
}
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 1)
```

# Throughout BA9 EX

```{r}
genes <- c("TBR1", "SATB2", "WFS1")
seurat <- MTG_neuronal_EX_seurat

FeaturePlot(
  seurat, features = genes, order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 2)

# Get top 10 highest expressed genes.
sub_seurat <- subset(seurat, idents = c(5, 6, 7, 2, 9, 12, 16))
dim(sub_seurat)
mat <- as.matrix(GetAssayData(sub_seurat, slot = "scale.data"))
expr_genes <- rowSums(mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))
```

# Neocortex L2-3 EX

```{r}
genes <- c("RELN", "NEFH")
seurat <- MTG_neuronal_EX_seurat

if (all(genes %in% rownames(seurat))) { # `blend` requires all features present.
  plots <- FeaturePlot(
    seurat, features = genes, order = TRUE,
    reduction = "umap", blend = TRUE, combine = FALSE
  )
  CombinePlots(plots[3:4], legend = "none")
} else {
  FeaturePlot(
    seurat, features = genes, order = TRUE, reduction = "umap",
    cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
  )
}
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 1)
```

# BA9 L2-3 EX

## Superficial

```{r}
genes <- c("CUX2", "LAMP5", "SERPINE2")
seurat <- MTG_neuronal_EX_seurat

sub_clusters <- c(
  "Exc L2-3 LINC00507 RPL9P17", "Exc L2-4 RORB GRIK1", "Exc L3-5 FEZF2 ONECUT1",
  "Exc L3 RORB CARTPT", "Exc L3-4 RORB FOLH1B", "Exc L3-4 RORB SEMA6D",
  "Exc L3-4 RORB PRSS12"
)
for (i in seq_along(sub_clusters)) {
  print(sub_clusters[i])

  # TODO: Remove hardcoded metadata call.
  print(DimPlot(
    seurat,
    order = TRUE,
    reduction = "umap",
    cells.highlight = WhichCells(seurat, expression = allen_clusters_name == sub_clusters[i])) +
      NoLegend()
  )
}

FeaturePlot(
  seurat, features = genes, order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 2)

# Get top 10 highest expressed genes.
sub_seurat <- subset(seurat, idents = c(5, 6, 7, 8))
dim(sub_seurat)
mat <- as.matrix(GetAssayData(sub_seurat, slot = "scale.data"))
expr_genes <- rowSums(mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))
```

## Deep

```{r}
genes <- c("CUX2", "COL5A2")
seurat <- MTG_neuronal_EX_seurat

if (all(genes %in% rownames(seurat))) { # `blend` requires all features present.
  plots <- FeaturePlot(
    seurat, features = genes, order = TRUE,
    reduction = "umap", blend = TRUE, combine = FALSE
  )
  CombinePlots(plots[3:4], legend = "none")
} else {
  FeaturePlot(
    seurat, features = genes, order = TRUE, reduction = "umap",
    cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
  )
}
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 1)
```

# BA9 L4-5 EX

## 1

```{r}
genes <- c("RORB", "GAL1")
seurat <- MTG_neuronal_EX_seurat

if (all(genes %in% rownames(seurat))) { # `blend` requires all features present.
  plots <- FeaturePlot(
    seurat, features = genes, order = TRUE,
    reduction = "umap", blend = TRUE, combine = FALSE
  )
  CombinePlots(plots[3:4], legend = "none")
} else {
  FeaturePlot(
    seurat, features = genes, order = TRUE, reduction = "umap",
    cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
  )
}
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 1)
```

## 2

```{r}
genes <- c("RORB", "PCP4")
seurat <- MTG_neuronal_EX_seurat

if (all(genes %in% rownames(seurat))) { # `blend` requires all features present.
  plots <- FeaturePlot(
    seurat, features = genes, order = TRUE,
    reduction = "umap", blend = TRUE, combine = FALSE
  )
  CombinePlots(plots[3:4], legend = "none")
} else {
  FeaturePlot(
    seurat, features = genes, order = TRUE, reduction = "umap",
    cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
  )
}
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 1)

# Get top 10 highest expressed genes.
sub_seurat <- subset(seurat, idents = c(3, 4))
dim(sub_seurat)
mat <- as.matrix(GetAssayData(sub_seurat, slot = "scale.data"))
expr_genes <- rowSums(mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))
```

# BA9 L5-6 EX

```{r}
genes <- c("NFIA", "THEMIS")
seurat <- MTG_neuronal_EX_seurat

if (all(genes %in% rownames(seurat))) { # `blend` requires all features present.
  plots <- FeaturePlot(
    seurat, features = genes, order = TRUE,
    reduction = "umap", blend = TRUE, combine = FALSE
  )
  CombinePlots(plots[3:4], legend = "none")
} else {
  FeaturePlot(
    seurat, features = genes, order = TRUE, reduction = "umap",
    cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
  )
}
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 1)

# Get top 10 highest expressed genes.
sub_seurat <- subset(seurat, idents = 0)
dim(sub_seurat)
mat <- as.matrix(GetAssayData(sub_seurat, slot = "scale.data"))
expr_genes <- rowSums(mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))
```

# BA10 L4 EX

```{r}
genes <- c("LINGO1", "RASGEF1B", "SLC26A3")
seurat <- MTG_neuronal_EX_seurat

FeaturePlot(
  seurat, features = genes, order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 2)
```

# BA9 Chandelier

GAD2 expresssion should be low!

```{r}
genes <- c("PVALB", "GAD1", "GAD2", "SCUBE3")
seurat <- MTG_neuronal_IN_seurat

sub_clusters <- c("Inh L1-6 PVALB SCUBE3")
for (i in seq_along(sub_clusters)) {
  print(sub_clusters[i])

  # TODO: Remove hardcoded metadata call.
  print(DimPlot(
    seurat,
    order = TRUE,
    reduction = "umap",
    cells.highlight = WhichCells(seurat, expression = allen_clusters_name == sub_clusters[i])) +
      NoLegend()
  )
}

FeaturePlot(
  seurat, features = genes, order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 2)

# Get top 10 highest expressed genes.
sub_seurat <- subset(seurat, allen_clusters_name == "Inh L1-6 PVALB SCUBE3")
dim(sub_seurat)
mat <- as.matrix(GetAssayData(sub_seurat, slot = "scale.data"))
expr_genes <- rowSums(mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))
```

# Hippocampus CA1 EX

```{r}
genes <- c("RELN", "NEFH", "SSTR4", "NOV")
seurat <- MTG_neuronal_EX_seurat

FeaturePlot(
  seurat, features = genes, order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 2)
```

# Hippocampus Subiculum

```{r}
genes <- c("RELN", "NEFH", "CARTPT")
seurat <- MTG_neuronal_EX_seurat

FeaturePlot(
  seurat, features = genes, order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 2)
```

# EC L2

```{r}
genes <- c("RELN", "RASGRP2", "SH3BGRL2")
seurat <- MTG_neuronal_EX_seurat

FeaturePlot(
  seurat, features = genes, order = TRUE, reduction = "umap",
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 2
)
DotPlot(seurat, features = genes, cols = c("blue", "red")) + RotatedAxis()
DoHeatmap(seurat, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill"
  )
VlnPlot(seurat, features = genes, ncol = 2)
```

# References

This is the concluding section of the document. Here we write relevant results to disk, output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
