---
title: '05 DE - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "05-DE.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).

In this particular document we probe for vulnerable cell-types.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

#    This file is part of tau-vulnerability.
#    Copyright (C) 2019  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# ------------------------------------------

packages <- c("conflicted", "Seurat", "dplyr", "ggplot2", "ggrepel", "DT", "data.table")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 5
protocol_type <- "smart-seq"
download_name <- "transcrip.tome" # Name of download file corresponding to dataset.

options(stringsAsFactors = FALSE)

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
data_name_stem <- unlist(strsplit(data_name, "-"))[1] # `data_name` up to the first hyphen.
data_dir <- file.path(getwd(), "..", "..", "data", data_name) # Publicly available data.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Misc binaries and temp files.
results_dir <- file.path(getwd(), "..", "..", "results", data_name)

# Unique cache and results directory for each analysis number.
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}
if (!dir.exists(file.path(results_dir, "data", paste0("0", analysis_no)))) {
  dir.create(file.path(results_dir, "data", paste0("0", analysis_no)), recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 10)
```

We use data saved from previous analyses.

```{r}
MTG_neuronal_seurat <- readRDS(
  file.path(results_dir, "data", paste0("0", analysis_no - 3), "MTG_neuronal_seurat.rds")
)
MTG_neuronal_seurat
seurat <- MTG_neuronal_seurat
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "class_label", "cat")

MTG_neuronal_EX_seurat <- readRDS(
  file.path(results_dir, "data", paste0("0", analysis_no - 3), "MTG_neuronal_EX_seurat.rds")
)
MTG_neuronal_EX_seurat
seurat <- MTG_neuronal_EX_seurat
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "class_label", "cat")

MTG_neuronal_IN_seurat <- readRDS(
  file.path(results_dir, "data", paste0("0", analysis_no - 3), "MTG_neuronal_IN_seurat.rds")
)
MTG_neuronal_IN_seurat
seurat <- MTG_neuronal_IN_seurat
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "class_label", "cat")
```

# MTG EX vs. IN

```{r}
# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_vs_IN_markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindMarkers(
    MTG_neuronal_seurat, ident.1 = c(6, 4, 0, 9, 2, 3, 20, 1, 10, 12, 15),
    ident.2 = c(8, 5, 16, 11, 14, 17, 19, 13, 7)
  )
  saveRDS(markers, rds)
}
datatable_custom(markers)
```

# Throughout BA9 EX/BA9 L2-3 EX Superficial vs. MTG IN

```{r}
# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "BA9_23_sup_vs_IN_markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindMarkers(
    MTG_neuronal_seurat, ident.1 = c(0, 4), ident.2 = c(8, 5, 16, 11, 14, 17, 19, 13, 7)
  )
  saveRDS(markers, rds)
}
datatable_custom(markers)
```

# Chandelier vs. MTG IN

```{r}
# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "chand_vs_IN_markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindMarkers(MTG_neuronal_IN_seurat, ident.1 = 9, ident.2 = c(1:8))
  saveRDS(markers, rds)
}
datatable_custom(markers)
```

# Throughout BA9 EX/BA9 L2-3 EX Superficial vs. MTG EX

```{r}
# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "BA9_23_sup_vs_EX_markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindMarkers(
    MTG_neuronal_EX_seurat, ident.1 = c(5, 6, 7), ident.2 = c(1:4, 8:16)
  )
  saveRDS(markers, rds)
}
datatable_custom(markers)
```

# Non-throughout BA9 EX/BA9 L2-3 EX Superficial vs. MTG IN

```{r}
# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "non_BA9_23_sup_vs_IN_markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindMarkers(
    MTG_neuronal_seurat, ident.1 = c(6, 9, 2, 3, 20, 1, 10, 12, 15),
    ident.2 = c(8, 5, 16, 11, 14, 17, 19, 13, 7)
  )
  saveRDS(markers, rds)
}
datatable_custom(markers)
```

# Throughout BA9 EX/BA9 L2-3 EX Superficial vs. L5-6 EX

```{r}
# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "BA9_23_sup_vs_56_EX.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindMarkers(MTG_neuronal_EX_seurat, ident.1 = c(5, 6, 7), ident.2 = 15)
  saveRDS(markers, rds)
}
datatable_custom(markers)
```

# L4-5 EX 2 vs. MTG IN

```{r}
# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "45_EX_2_vs_IN_markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindMarkers(
    MTG_neuronal_seurat, ident.1 = 1, ident.2 = c(8, 5, 16, 11, 14, 17, 19, 13, 7)
  )
  saveRDS(markers, rds)
}
datatable_custom(markers)
```

# L4-5 EX 2 vs. Throughout BA9 EX/BA9 L2-3 EX Superficial

```{r}
# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "45_EX_2_vs_BA9_23_sup.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindMarkers(MTG_neuronal_seurat, ident.1 = 1, ident.2 = c(0, 4))
  saveRDS(markers, rds)
}
datatable_custom(markers)
```

# Non-chandelier vs. MTG IN

```{r}
# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "non_chand_vs_IN_markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindMarkers(MTG_neuronal_IN_seurat, ident.1 = 8, ident.2 = c(1:7, 9))
  saveRDS(markers, rds)
}
datatable_custom(markers)
```

# References

This is the concluding section of the document. Here we write relevant results to disk, output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
