---
title: '07 GSVA - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "07-gsva.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).

In this particular document we probe for vulnerable cell-types.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[6]``.

```{r}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

#    This file is part of tau-vulnerability.
#    Copyright (C) 2019  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# ------------------------------------------

packages <- c(
  "conflicted", "Seurat", "dplyr", "ggplot2", "ggrepel", "GSEABase", "data.table", "DT", "GSVA",
  "parallel", "limma", "RColorBrewer"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 7
protocol_type <- "smart-seq"
download_name <- "transcrip.tome" # Name of download file corresponding to dataset.

options(stringsAsFactors = FALSE)

# Heatmap color variables.
color_legend <- c("lightgray", "darkgray")
color <- rev(colorRampPalette(brewer.pal(n = 7, name = "RdBu"))(100))

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
data_name_stem <- unlist(strsplit(data_name, "-"))[1] # `data_name` up to the first hyphen.
data_dir <- file.path(getwd(), "..", "..", "data", data_name) # Publicly available data.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Misc binaries and temp files.
results_dir <- file.path(getwd(), "..", "..", "results", data_name)

# Unique cache and results directory for each analysis number.
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}
if (!dir.exists(file.path(results_dir, "data", paste0("0", analysis_no)))) {
  dir.create(file.path(results_dir, "data", paste0("0", analysis_no)), recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 10)
```

We use data saved from previous analyses.

```{r}
MTG_neuronal_seurat <- readRDS(
  file.path(results_dir, "data", paste0("0", analysis_no - 5), "MTG_neuronal_seurat.rds")
)
MTG_neuronal_seurat
seurat <- MTG_neuronal_seurat
red_dim_plot(seurat, "umap1", "umap2", "seurat_clusters", "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", paste0(data_name_stem, "_clusters_name"), "cat") +
  NoLegend()
red_dim_plot(seurat, "umap1", "umap2", "class_label", "cat")
```

# EX vs. IN

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "Glutamatergic"
sub[[2]] <- "GABAergic"
conditions <- c("EX", "IN")
seurat <- MTG_neuronal_seurat
expr <- FetchData(seurat, vars = "class_label")

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets
```

This limma-based design mainly follows the GSVA vignette on BioConductor.

```{r}
# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr == sub[[1]])]
condition1_seurat <- subset(condition1_seurat, downsample = 10) # TODO: Run without downsampling.
condition1_mat <- as.matrix(GetAssayData(object = condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr == sub[[2]])]
condition2_seurat <- subset(condition2_seurat, downsample = 10) # TODO: Run without downsampling.
condition2_mat <- as.matrix(GetAssayData(object = condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]
```

```{r}
# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_vs_IN_gsva.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)
}
```

```{r}
# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[1])

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[1])

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(fit, coef = conditions[1], number = 25, adjust.method = "holm")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 40),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# References

This is the concluding section of the document. Here we write relevant results to disk, output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
