---
title: '07 GSVA - `r unlist(strsplit(getwd(), "/"))[6]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: '../../`r unlist(strsplit(getwd(), "/"))[4]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(
    "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "07-gsva.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).

In this particular document we probe for vulnerable cell-types.
The data here is derived from @`r unlist(strsplit(getwd(), "/"))[6]` and will be referenced using the name `r unlist(strsplit(getwd(), "/"))[6]`.

```{r}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

#    This file is part of tau-vulnerability.
#    Copyright (C) 2019  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# These should be checked per document.
# ------------------------------------------

packages <- c(
  "conflicted", "Seurat", "GSEABase", "data.table", "DT", "GSVA", "parallel", "limma",
  "RColorBrewer", "pheatmap"
)
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 7
protocol_type <- "smart-seq"
download_name <- "transcrip.tome" # Name of download file corresponding to dataset.

options(stringsAsFactors = FALSE)

# Heatmap color variables.
color_legend <- c("lightgray", "darkgray")
color <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100)
color2 <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100)
color2[50] <- "#DDDDDD"

gsva_min_size <- 3 # Minimum size of gene sets.

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset.
data_name_stem <- unlist(strsplit(data_name, "-"))[1] # `data_name` up to the first hyphen.
data_dir <- file.path(getwd(), "..", "..", "data", data_name) # Publicly available data.
assets_dir <- file.path(getwd(), "..", "..", "assets", data_name) # Misc binaries and temp files.
results_dir <- file.path(getwd(), "..", "..", "results", data_name)

# Unique cache and results directory for each analysis number.
if (!dir.exists(file.path(assets_dir, "cache", paste0("0", analysis_no)))) {
  dir.create(file.path(assets_dir, "cache", paste0("0", analysis_no)), recursive = TRUE)
}
if (!dir.exists(file.path(results_dir, "data", paste0("0", analysis_no)))) {
  dir.create(file.path(results_dir, "data", paste0("0", analysis_no)), recursive = TRUE)
}

knitr::opts_chunk$set(fig.width = 10, fig.height = 10)
```

We use data saved from previous analyses.

```{r}
MTG_neuronal_seurat <- readRDS(
  file.path(results_dir, "data", paste0("0", analysis_no - 5), "MTG_neuronal_seurat.rds")
)
MTG_neuronal_seurat
```

# MTG EX vs. MTG IN

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "GABAergic"
sub[[2]] <- "Glutamatergic"
conditions <- c("MTG IN", "MTG EX")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = "class_label")

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_vs_MTG_IN.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX vs. IN (Subproteomes)

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "GABAergic"
sub[[2]] <- "Glutamatergic"
conditions <- c("MTG IN", "MTG EX")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = "class_label")

set_name <- "subproteomes"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_vs_MTG_IN_subpro.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm")

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 20),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L2-3 Superficial vs. MTG IN

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "GABAergic"
sub[[2]] <- "Exc L2-3 LINC00507 RPL9P17"
conditions <- c("MTG IN", "MTG EX L2-3 Superficial")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_23_sup_vs_MTG_IN.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

comb1 <- fit # Prepare for combined analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L2-3 Superficial vs. MTG IN (Subproteomes)

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "GABAergic"
sub[[2]] <- "Exc L2-3 LINC00507 RPL9P17"
conditions <- c("MTG IN", "MTG EX L2-3 Superficial")
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

set_name <- "subproteomes"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_23_sup_vs_MTG_IN_subpro.rds"
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm")

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 20),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L2-3 Superficial vs. MTG EX Remaining

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "Glutamatergic"
sub[[2]] <- "Exc L2-3 LINC00507 RPL9P17"
conditions <- c("MTG EX Remaining", "MTG EX L2-3 Superficial")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
expr1 <- FetchData(condition1_seurat, vars = paste0(data_name_stem, "_clusters_name"))
condition1_seurat <- condition1_seurat[ , with(expr1, which(!(allen_clusters_name %in% sub[[2]])))]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_23_sup_vs_MTG_EX_remain.rds"
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L2-3 Superficial vs. 3 MTG IN

```{r}
names <- c("MTG PVALB", "MTG CALB2", "MTG SST")
```

## `r names[1]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 5
sub[[2]] <- "Exc L2-3 LINC00507 RPL9P17"
conditions <- c(names[1], "MTG EX L2-3 Superficial")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_23_sup_vs_", names[1], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit1 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## `r names[2]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 7
sub[[2]] <- "Exc L2-3 LINC00507 RPL9P17"
conditions <- c(names[2], "MTG EX L2-3 Superficial")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_23_sup_vs_", names[2], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit2 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## `r names[3]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 8
sub[[2]] <- "Exc L2-3 LINC00507 RPL9P17"
conditions <- c(names[3], "MTG EX L2-3 Superficial")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_23_sup_vs_", names[3], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit3 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Intersections

```{r}
sets <- list(
  topTable(fit1, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05),
  topTable(fit2, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05),
  topTable(fit3, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05)
)
sets <- lapply(sets, function(x) cbind(x , rownames(x)))
sets <- lapply(sets, function(x) x[ , c(dim(x)[2], 1)])
colnames(sets[[1]])[2] <- paste0(names[1], " log2foldFC")
colnames(sets[[2]])[2] <- paste0(names[2], " log2foldFC")
colnames(sets[[3]])[2] <- paste0(names[3], " log2foldFC")
merged_set <- Reduce(function(...) merge(..., all = TRUE), list(
  sets[[1]][ , c(1, 2)], sets[[2]][ , c(1, 2)], sets[[3]][ , c(1, 2)]
))

set1 <- head(merged_set[order(-abs(merged_set[[paste0(names[1], " log2foldFC")]])), ], n = 100)
set2 <- head(merged_set[order(-abs(merged_set[[paste0(names[2], " log2foldFC")]])), ], n = 100)
set3 <- head(merged_set[order(-abs(merged_set[[paste0(names[3], " log2foldFC")]])), ], n = 100)

merged_top <- Reduce(function(...) merge(..., all = TRUE), list(
  set1[ , c(1, 2)], set2[ , c(1, 3)], set3[ , c(1, 4)]
))
top_genes <- merged_top$`rownames(x)`
merged_top <- merged_top[ , 2:dim(merged_top)[2]]
rownames(merged_top) <- top_genes
datatable_custom(merged_top) %>%
  formatRound(columns = c(1:dim(merged_top)[2]), digits = 2)

# Generate heatmap, using grey for 0 in color palette to indicate `NA` values.
merged_top <- as.matrix(merged_top)
rownames(merged_top) <- top_genes
merged_top <- replace(merged_top, is.na(merged_top), 0)
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 45, fontsize = 9, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 4
)

# Generate heatmap without `NA` values.
merged_top[merged_top == 0] <- NA
merged_top <- merged_top[complete.cases(merged_top), ]
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 45, fontsize = 9, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 6
)
```

# MTG EX L2-3 Deep vs. MTG IN

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "GABAergic"
sub[[2]] <- c("Exc L3 RORB CARTPT", "Exc L3-4 RORB FOLH1B")
conditions <- c("MTG IN", "MTG EX L2-3 Deep")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_23_deep_vs_MTG_IN.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

comb2 <- fit # Prepare for combined analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L2-3 Deep vs. MTG IN (Subproteomes)

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "GABAergic"
sub[[2]] <- c("Exc L3 RORB CARTPT", "Exc L3-4 RORB FOLH1B")
conditions <- c("MTG IN", "MTG EX L2-3 Deep")
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

set_name <- "subproteomes"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_23_deep_vs_MTG_IN_subpro.rds"
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm")

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 20),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L2-3 Deep vs. MTG EX Remaining

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "Glutamatergic"
sub[[2]] <- c("Exc L3 RORB CARTPT", "Exc L3-4 RORB FOLH1B")
conditions <- c("MTG EX Remaining", "MTG EX L2-3 Deep")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
expr1 <- FetchData(condition1_seurat, vars = paste0(data_name_stem, "_clusters_name"))
condition1_seurat <- condition1_seurat[ , with(expr1, which(!(allen_clusters_name %in% sub[[2]])))]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 != sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_23_deep_vs_MTG_EX_remain.rds"
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L2-3 Deep vs. 3 MTG IN

```{r}
names <- c("MTG PVALB", "MTG CALB2", "MTG SST")
```

## `r names[1]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 5
sub[[2]] <- c("Exc L3 RORB CARTPT", "Exc L3-4 RORB FOLH1B")
conditions <- c(names[1], "MTG EX L2-3 Deep")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_23_deep_vs_", names[1], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit1 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## `r names[2]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 7
sub[[2]] <- c("Exc L3 RORB CARTPT", "Exc L3-4 RORB FOLH1B")
conditions <- c(names[2], "MTG EX L2-3 Deep")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_23_deep_vs_", names[2], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit2 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## `r names[3]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 8
sub[[2]] <- c("Exc L3 RORB CARTPT", "Exc L3-4 RORB FOLH1B")
conditions <- c(names[3], "MTG EX L2-3 Deep")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_23_deep_vs_", names[3], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit3 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Intersections

```{r}
sets <- list(
  topTable(fit1, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05),
  topTable(fit2, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05),
  topTable(fit3, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05)
)
sets <- lapply(sets, function(x) cbind(x , rownames(x)))
sets <- lapply(sets, function(x) x[ , c(dim(x)[2], 1)])
colnames(sets[[1]])[2] <- paste0(names[1], " log2foldFC")
colnames(sets[[2]])[2] <- paste0(names[2], " log2foldFC")
colnames(sets[[3]])[2] <- paste0(names[3], " log2foldFC")
merged_set <- Reduce(function(...) merge(..., all = TRUE), list(
  sets[[1]][ , c(1, 2)], sets[[2]][ , c(1, 2)], sets[[3]][ , c(1, 2)]
))

set1 <- head(merged_set[order(-abs(merged_set[[paste0(names[1], " log2foldFC")]])), ], n = 100)
set2 <- head(merged_set[order(-abs(merged_set[[paste0(names[2], " log2foldFC")]])), ], n = 100)
set3 <- head(merged_set[order(-abs(merged_set[[paste0(names[3], " log2foldFC")]])), ], n = 100)

merged_top <- Reduce(function(...) merge(..., all = TRUE), list(
  set1[ , c(1, 2)], set2[ , c(1, 3)], set3[ , c(1, 4)]
))
top_genes <- merged_top$`rownames(x)`
merged_top <- merged_top[ , 2:dim(merged_top)[2]]
rownames(merged_top) <- top_genes
datatable_custom(merged_top) %>%
  formatRound(columns = c(1:dim(merged_top)[2]), digits = 2)

# Generate heatmap, using grey for 0 in color palette to indicate `NA` values.
merged_top <- as.matrix(merged_top)
rownames(merged_top) <- top_genes
merged_top <- replace(merged_top, is.na(merged_top), 0)
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 45, fontsize = 9, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 4
)

# Generate heatmap without `NA` values.
merged_top[merged_top == 0] <- NA
merged_top <- merged_top[complete.cases(merged_top), ]
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 45, fontsize = 9, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 6
)
```

# MTG EX L4-5 vs. MTG IN

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "GABAergic"
sub[[2]] <- c("Exc L4-5 RORB HNRNPA1P46", "Exc L4-5 RORB LCN15")
conditions <- c("MTG IN", "MTG EX L4-5")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_45_vs_MTG_IN.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

comb3 <- fit # Prepare for combined analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L4-5 vs. MTG IN (Subproteomes)

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "GABAergic"
sub[[2]] <- c("Exc L4-5 RORB HNRNPA1P46", "Exc L4-5 RORB LCN15")
conditions <- c("MTG IN", "MTG EX L4-5")
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

set_name <- "subproteomes"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_45_vs_MTG_IN_subpro.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm")

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 20),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L4-5 vs. MTG EX Remaining

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "Glutamatergic"
sub[[2]] <- c("Exc L4-5 RORB HNRNPA1P46", "Exc L4-5 RORB LCN15")
conditions <- c("MTG EX Remaining", "MTG EX L4-5")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
expr1 <- FetchData(condition1_seurat, vars = paste0(data_name_stem, "_clusters_name"))
condition1_seurat <- condition1_seurat[ , with(expr1, which(!(allen_clusters_name %in% sub[[2]])))]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_45_vs_MTG_EX_remain.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L4-5 vs. 3 MTG IN

```{r}
names <- c("MTG PVALB", "MTG CALB2", "MTG SST")
```

## `r names[1]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 5
sub[[2]] <- c("Exc L4-5 RORB HNRNPA1P46", "Exc L4-5 RORB LCN15")
conditions <- c(names[1], "MTG EX L4-5")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_45_vs_", names[1], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit1 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## `r names[2]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 7
sub[[2]] <- c("Exc L4-5 RORB HNRNPA1P46", "Exc L4-5 RORB LCN15")
conditions <- c(names[2], "MTG EX L4-5")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_45_vs_", names[2], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit2 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## `r names[3]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 8
sub[[2]] <- c("Exc L4-5 RORB HNRNPA1P46", "Exc L4-5 RORB LCN15")
conditions <- c(names[3], "MTG EX L4-5")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , with(expr2, which(allen_clusters_name %in% sub[[2]]))]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_45_vs_", names[3], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit3 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Intersections

```{r}
sets <- list(
  topTable(fit1, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05),
  topTable(fit2, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05),
  topTable(fit3, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05)
)
sets <- lapply(sets, function(x) cbind(x , rownames(x)))
sets <- lapply(sets, function(x) x[ , c(dim(x)[2], 1)])
colnames(sets[[1]])[2] <- paste0(names[1], " log2foldFC")
colnames(sets[[2]])[2] <- paste0(names[2], " log2foldFC")
colnames(sets[[3]])[2] <- paste0(names[3], " log2foldFC")
merged_set <- Reduce(function(...) merge(..., all = TRUE), list(
  sets[[1]][ , c(1, 2)], sets[[2]][ , c(1, 2)], sets[[3]][ , c(1, 2)]
))

set1 <- head(merged_set[order(-abs(merged_set[[paste0(names[1], " log2foldFC")]])), ], n = 100)
set2 <- head(merged_set[order(-abs(merged_set[[paste0(names[2], " log2foldFC")]])), ], n = 100)
set3 <- head(merged_set[order(-abs(merged_set[[paste0(names[3], " log2foldFC")]])), ], n = 100)

merged_top <- Reduce(function(...) merge(..., all = TRUE), list(
  set1[ , c(1, 2)], set2[ , c(1, 3)], set3[ , c(1, 4)]
))
top_genes <- merged_top$`rownames(x)`
merged_top <- merged_top[ , 2:dim(merged_top)[2]]
rownames(merged_top) <- top_genes
datatable_custom(merged_top) %>%
  formatRound(columns = c(1:dim(merged_top)[2]), digits = 2)

# Generate heatmap, using grey for 0 in color palette to indicate `NA` values.
merged_top <- as.matrix(merged_top)
rownames(merged_top) <- top_genes
merged_top <- replace(merged_top, is.na(merged_top), 0)
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 45, fontsize = 7, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 4
)

# Generate heatmap without `NA` values.
merged_top[merged_top == 0] <- NA
merged_top <- merged_top[complete.cases(merged_top), ]
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 45, fontsize = 9, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 6
)
```

# MTG EX L5-6 vs. MTG IN

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "GABAergic"
sub[[2]] <- "Exc L6 THEMIS LINC00343"
conditions <- c("MTG IN", "MTG EX L5-6")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_56_vs_MTG_IN.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

comb4 <- fit # Prepare for combined analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L5-6 vs. MTG IN (Subproteomes)

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "GABAergic"
sub[[2]] <- "Exc L6 THEMIS LINC00343"
conditions <- c("MTG IN", "MTG EX L5-6")
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

set_name <- "subproteomes"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_56_vs_MTG_IN_subpro.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm")

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 20),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L5-6 vs. MTG EX Remaining

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- "Glutamatergic"
sub[[2]] <- "Exc L6 THEMIS LINC00343"
conditions <- c("MTG EX Remaining", "MTG EX L5-6")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "class_label")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- seurat[ , which(expr1 == sub[[1]])]
expr1 <- FetchData(condition1_seurat, vars = paste0(data_name_stem, "_clusters_name"))
condition1_seurat <- condition1_seurat[ , with(expr1, which(!(allen_clusters_name %in% sub[[2]])))]
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), "MTG_EX_56_vs_MTG_EX_remain.rds"
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

# MTG EX L5-6 vs. 3 MTG IN

```{r}
names <- c("MTG PVALB", "MTG CALB2", "MTG SST")
```

## `r names[1]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 5
sub[[2]] <- "Exc L6 THEMIS LINC00343"
conditions <- c(names[1], "MTG EX L5-6")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_56_vs_", names[1], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit1 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## `r names[2]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 7
sub[[2]] <- "Exc L6 THEMIS LINC00343"
conditions <- c(names[2], "MTG EX L5-6")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_56_vs_", names[2], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit2 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## `r names[3]`

```{r}
sub <- vector(mode = "list", length = 2)
sub[[1]] <- 8
sub[[2]] <- "Exc L6 THEMIS LINC00343"
conditions <- c(names[3], "MTG EX L5-6")
seurat <- MTG_neuronal_seurat
expr1 <- FetchData(seurat, vars = "ident")
expr2 <- FetchData(seurat, vars = paste0(data_name_stem, "_clusters_name"))

# The gene set to be used, comprised of GO biological pathways.
# Downloaded from MSigDB http://software.broadinstitute.org/gsea/msigdb/collections.jsp.
set_name <- "c5.bp.v7.0.symbols"
gene_sets <- getGmt(file.path(assets_dir, "..", "gene-sets", paste0(set_name, ".gmt")))
gene_sets

# Subset populations, note that the "data" slot is used.
# This is the same default as when doing DEG analysis with Seurat.
condition1_seurat <- subset(seurat, idents = sub[[1]]) # TODO: Get FetchData working with idents.
condition1_mat <- as.matrix(GetAssayData(condition1_seurat))
dim(condition1_mat)

# Sanity check by getting rowsums for the subset.
comb_mat <- condition1_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Now for condition 2.
condition2_seurat <- seurat[ , which(expr2 == sub[[2]])]
condition2_mat <- as.matrix(GetAssayData(condition2_seurat))
dim(condition2_mat)

# Sanity check again.
comb_mat <- condition2_mat
expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

# Bind together the matrices and setup the design matrix and heatmap colors.
comb_mat <- cbind(condition1_mat, condition2_mat)
groups <- factor(rep(c(0, 1), c(dim(condition1_mat)[2], dim(condition2_mat)[2])))
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Cache the results.
rds <- file.path(
  assets_dir, "cache", paste0("0", analysis_no), paste0("MTG_EX_56_vs_", names[3], ".rds")
)
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = gsva_min_size)
  saveRDS(gsva, rds)
}

# Fit a linear model to the matrix of genes, not GSVA output, for a basic DEG analysis.
# This step is not neccessary, but is useful to view alongside differential pathway anaysis.
# Positive logFC values are those upreguled in the coef argument.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

# Fit another linear model, this time performing differential pathway analysis on GSVA output.
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = conditions[2], adjust.method = "holm", p.value = 0.05)

fit3 <- fit # Prepare for intersection analysis.

# Set a cutoff and generate a heatmap from the data.
DE_gene_sets <- topTable(
  fit, coef = conditions[2], number = 25, adjust.method = "holm", p.value = 0.05
)
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]
names(sample_color_map) <- colnames(gsva_cutoff)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = data_name, margins = c(2, 50),
  col = color, ylab = set_name,
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol = " ",
  Rowv = as.dendrogram(gene_set_clustering), Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Intersections

```{r}
sets <- list(
  topTable(fit1, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05),
  topTable(fit2, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05),
  topTable(fit3, number = Inf, coef = conditions[2], adjust.method = "holm", p.value = 0.05)
)
sets <- lapply(sets, function(x) cbind(x , rownames(x)))
sets <- lapply(sets, function(x) x[ , c(dim(x)[2], 1)])
colnames(sets[[1]])[2] <- paste0(names[1], " log2foldFC")
colnames(sets[[2]])[2] <- paste0(names[2], " log2foldFC")
colnames(sets[[3]])[2] <- paste0(names[3], " log2foldFC")
merged_set <- Reduce(function(...) merge(..., all = TRUE), list(
  sets[[1]][ , c(1, 2)], sets[[2]][ , c(1, 2)], sets[[3]][ , c(1, 2)]
))

set1 <- head(merged_set[order(-abs(merged_set[[paste0(names[1], " log2foldFC")]])), ], n = 100)
set2 <- head(merged_set[order(-abs(merged_set[[paste0(names[2], " log2foldFC")]])), ], n = 100)
set3 <- head(merged_set[order(-abs(merged_set[[paste0(names[3], " log2foldFC")]])), ], n = 100)

merged_top <- Reduce(function(...) merge(..., all = TRUE), list(
  set1[ , c(1, 2)], set2[ , c(1, 3)], set3[ , c(1, 4)]
))
top_genes <- merged_top$`rownames(x)`
merged_top <- merged_top[ , 2:dim(merged_top)[2]]
rownames(merged_top) <- top_genes
datatable_custom(merged_top) %>%
  formatRound(columns = c(1:dim(merged_top)[2]), digits = 2)

# Generate heatmap, using grey for 0 in color palette to indicate `NA` values.
merged_top <- as.matrix(merged_top)
rownames(merged_top) <- top_genes
merged_top <- replace(merged_top, is.na(merged_top), 0)
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 45, fontsize = 9, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 4
)

# Generate heatmap without `NA` values.
merged_top[merged_top == 0] <- NA
merged_top <- merged_top[complete.cases(merged_top), ]
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 45, fontsize = 9, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 6
)
```

# Vulnerable Clusters Combined

```{r}
names <- c("MTG EX L2-3 Superficial", "MTG EX L2-3 Deep", "MTG EX L4-5", "MTG EX L5-6")

sets <- list(
  topTable(comb1, number = Inf, coef = names[1], adjust.method = "holm", p.value = 0.05),
  topTable(comb2, number = Inf, coef = names[2], adjust.method = "holm", p.value = 0.05),
  topTable(comb3, number = Inf, coef = names[3], adjust.method = "holm", p.value = 0.05),
  topTable(comb4, number = Inf, coef = names[4], adjust.method = "holm", p.value = 0.05)
)
sets <- lapply(sets, function(x) cbind(x , rownames(x)))
sets <- lapply(sets, function(x) x[ , c(dim(x)[2], 1)])
colnames(sets[[1]])[2] <- paste0(names[1], " log2foldFC")
colnames(sets[[2]])[2] <- paste0(names[2], " log2foldFC")
colnames(sets[[3]])[2] <- paste0(names[3], " log2foldFC")
colnames(sets[[4]])[2] <- paste0(names[4], " log2foldFC")
merged_set <- Reduce(function(...) merge(..., all = TRUE), list(
  sets[[1]][ , c(1, 2)], sets[[2]][ , c(1, 2)], sets[[3]][ , c(1, 2)], sets[[4]][ , c(1, 2)]
))

set1 <- head(merged_set[order(-abs(merged_set[[paste0(names[1], " log2foldFC")]])), ], n = 50)
set2 <- head(merged_set[order(-abs(merged_set[[paste0(names[2], " log2foldFC")]])), ], n = 50)
set3 <- head(merged_set[order(-abs(merged_set[[paste0(names[3], " log2foldFC")]])), ], n = 50)
set4 <- head(merged_set[order(-abs(merged_set[[paste0(names[4], " log2foldFC")]])), ], n = 50)

merged_top <- Reduce(function(...) merge(..., all = TRUE), list(
  set1[ , c(1, 2)], set2[ , c(1, 3)], set3[ , c(1, 4)], set4[ , c(1, 5)]
))
top_genes <- merged_top$`rownames(x)`
merged_top <- merged_top[ , 2:dim(merged_top)[2]]
rownames(merged_top) <- top_genes
datatable_custom(merged_top) %>%
  formatRound(columns = c(1:dim(merged_top)[2]), digits = 2)

# Generate heatmap, using grey for 0 in color palette to indicate `NA` values.
merged_top <- as.matrix(merged_top)
rownames(merged_top) <- top_genes
merged_top <- replace(merged_top, is.na(merged_top), 0)
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 45, fontsize = 8, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 4
)

# Generate heatmap without `NA` values.
merged_top[merged_top == 0] <- NA
merged_top <- merged_top[complete.cases(merged_top), ]
breaks <- c(
  seq(min(merged_top), -0.001, by = abs(min(merged_top)) / 50), 0,
  seq(0.001, max(merged_top), by = abs(max(merged_top)) / 50)
)
pheatmap(
  merged_top, angle_col = 45, fontsize = 7, color = color2,
  cluster_cols = FALSE, breaks = breaks, fontsize_row = 6
)
```

# References

This is the concluding section of the document. Here we write relevant results to disk, output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
