---
title: "10 Lake 2016 snRNAseq Prep"
author:
  - name: "Emir Turkes [et2628@cumc.columbia.edu]"
  - name: "Columbia University"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: "../tau-vulnerability.bib"
biblio-style: apalike
link-citations: true
output:
  html_document:
    code_folding: show
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile, encoding = encoding, output_file = "../results/10-lake-2016-snRNAseq-prep.html")})
---

```{r, include = FALSE}
#    This file is part of tau-vulnerability.
#    Copyright (C) 2019  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

knitr::opts_chunk$set(fig.width = 8.5, fig.height = 7)
```

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

In this document we apply additional processing on the cleaned dataset derived from the analysis in [github.com/eturkes/lake-2016-snRNAseq](https://github.com/eturkes/lake-2016-snRNAseq).
We start by setting some global variables and loading in any required packages.

```{r}
assets_dir <- file.path(getwd(), "..", "assets")
results_dir <- file.path(getwd(), "..", "results")

packages <- c(
  "conflicted", "SingleCellExperiment", "magrittr", "dplyr", "ggplot2", "ggrepel", "S4Vectors",
  "SummarizedExperiment", "DropletUtils", "scran", "BiocSingular", "scater", "Rtsne", "svd",
  "SC3", "DT", "data.table", "Seurat", "uwot", "viridis")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))

conflict_prefer("which", "BiocGenerics")
options(stringsAsFactors = FALSE)

# Create a unique cache and results data directory for each iterated section.
if (!dir.exists(file.path(assets_dir, "cache", "10"))) {
  dir.create(file.path(assets_dir, "cache", "10"), recursive = TRUE)}
if (!dir.exists(file.path(results_dir, "data", "10"))) {
  dir.create(file.path(results_dir, "data", "10"), recursive = TRUE)}

# ggplot2 function providing custom aesthetics and automatic placement of categorical labels.
# For continuous data, a colorbar is implemented.
dim_red_plot <- function(data, x, y, col, type) {
  gg <- ggplot(data, aes_string(x = x, y = y, color = col)) +
    geom_point(alpha = 0.35, stroke = 0.05, shape = 21, aes_string(fill = col)) +
    theme_classic() +
    theme(
      legend.position = "right", plot.title = element_text(hjust = 0.5),
      legend.title = element_blank()) +
    guides(color = guide_legend(override.aes = list(alpha = 1)))
    if (type == "cat") {
      gg <- gg + geom_label_repel(data = label_df2, aes(label = label), show.legend = FALSE)
    } else if (type == "cont") {
      gg <- ggplot(data, aes_string(x = x, y = y)) +
        geom_point(alpha = 0.35, stroke = 0.05, aes_string(color = col)) +
        theme_classic() +
        theme(
          legend.position = "right", plot.title = element_text(hjust = 0.5),
          legend.title = element_blank()) +
        scale_colour_viridis()}
  gg}
```

# Read in Normalized Data

We do not perform QC as only normalized TPMs were distributed and not raw counts.

```{r}
sce <- readRDS(file.path(assets_dir, "lake-2016-snRNAseq", "sce_orig.rds"))
sce
```

# Dimensionality Reduction

## PPK SVD + UMAP

We calculate a truncated SVD (singular value decomposition) from which PCs can be extracted from it.
In our testing, this approach, particularly the PROPACK implementation, was found to have high power in distinguishing subpopulations of common cell types.

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "10", "ppk.rds")
if (file.exists(rds)) {
  ppk <- readRDS(rds)
} else {
  set.seed(1)
  ppk <- propack.svd(scale(t(as.matrix(logcounts(sce)))), neig = 50)
  saveRDS(ppk, rds)}
```

```{r}
pca <- t(ppk$d * t(ppk$u))
retain_pcs <- 27
retain_pcs
par(mfrow = c(1, 1))
plot(
  log10(ppk$d), xlab = "PC", ylab = "log10(Proportion of Variance Explained)",
  pch = 20, cex = 0.6, col = rgb(0.8, 0.2, 0.2, 0.5))
abline(v = retain_pcs, lty = 2, col = "red")
```

```{r}
rds <- file.path(assets_dir, "cache", "10", "ppk_umap.rds")
if (file.exists(rds)) {
  umap <- readRDS(rds)
} else {
  set.seed(1)
  umap <- umap(pca[ , 1:retain_pcs], min_dist = 0.5)
  saveRDS(umap, rds)}
```

```{r}
reducedDims(sce) <- SimpleList(PCA = pca, UMAP = umap)
add_df <- data.frame(reducedDim(sce, "UMAP"))
names(add_df) <- paste0("ppk_umap", seq(ncol(add_df)))
colData(sce) <- cbind(
  colData(sce), ppk_umap1 = add_df$ppk_umap1, ppk_umap2 = add_df$ppk_umap2)

# Set up data frame for ggplot2.
gg_df <- data.frame(colData(sce)[ , c("lake_cluster_name", paste0("ppk_umap", 1:2))])
rownames(gg_df) <- NULL
gg_df$lake_cluster_name <- factor(gg_df$lake_cluster_name)

# Setup for automatic placement of cluster labels.
label_df <- data.frame(
  lake_cluster_name = levels(gg_df$lake_cluster_name),
  label = levels(gg_df$lake_cluster_name))
label_df2 <- gg_df %>%
  group_by(lake_cluster_name) %>%
  summarize(ppk_umap1 = median(ppk_umap1), ppk_umap2 = median(ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "ppk_umap1", y = "ppk_umap2", col = "lake_cluster_name", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") + ggtitle("Lake Clusters PPK UMAP")
```
