---
title: "15 Habib Neuronal Pathway Analysis Normalization Benchmarks"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: "../tau-vulnerability.bib"
biblio-style: apalike
link-citations: true
output:
  html_document:
    code_folding: show
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile, encoding = encoding,
      output_file = "../results/15-habib-neuronal-path-norm-benchmark.html")})
---

```{r, include = FALSE}
#    This file is part of tau-vulnerability.
#    Copyright (C) 2019  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

knitr::opts_chunk$set(fig.width = 15, fig.height = 13)
```

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

This document follows up its preparation file to provide and in-depth analysis of neuronal populations in the dataset from @habib_massively_2017.
We start by setting some global variables and loading in any required packages.

```{r}
assets_dir <- file.path(getwd(), "..", "assets")
results_dir <- file.path(getwd(), "..", "results")

packages <- c(
  "conflicted", "SingleCellExperiment", "magrittr", "dplyr", "ggplot2", "ggrepel", "S4Vectors",
  "SummarizedExperiment", "DropletUtils", "scran", "BiocSingular", "scater", "Rtsne", "svd", "SC3",
  "DT", "data.table", "Seurat", "uwot", "viridis", "pheatmap", "RColorBrewer", "GSVA", "GSEABase",
  "limma")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))

conflict_prefer("which", "BiocGenerics")
options(stringsAsFactors = FALSE)

# Create a unique cache and results data directory for each iterated section.
if (!dir.exists(file.path(assets_dir, "cache", "15"))) {
  dir.create(file.path(assets_dir, "cache", "15"), recursive = TRUE)}
if (!dir.exists(file.path(results_dir, "data", "15"))) {
  dir.create(file.path(results_dir, "data", "15"), recursive = TRUE)}

# Adds download buttons.
datatable_custom <- function(dt) {
  datatable(
    dt,
    extensions = "Buttons", options = list(dom = "Blfrtip", buttons = list(
      "copy", "print",
      list(extend = "collection", buttons = c("csv", "excel", "pdf"), text = "Download"))))}

gene_sets <- getGmt(file.path(assets_dir, "gene-sets", "c5.bp.v7.0.symbols.gmt"))
color_legend <- c("lightgray", "darkgray")
color <- rev(colorRampPalette(brewer.pal(n = 7, name = "RdBu"))(100))
IN_p_value <- 1e-25
EX_p_value <- 1e-25
comb_p_value <- 1e-100
```

# Whole-set Normalization

```{r}
sce_hvg <- readRDS(file.path(results_dir, "data", "13", "sce_hvg.rds"))
seurat_hvg <- readRDS(file.path(results_dir, "data", "13", "seurat_hvg.rds"))
sce_orig <- readRDS(file.path(results_dir, "data", "08", "sce_orig.rds"))
sce_orig <- calculateQCMetrics(sce_orig)
seurat_orig <- as.Seurat(sce_orig)
seurat_orig <- NormalizeData(seurat_orig)
seurat_orig <- ScaleData(seurat_orig, features = rownames(seurat_orig))

keep <- readRDS(file.path(results_dir, "data", "08", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "08", "keep2.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "13", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]

umap <- as.matrix(data.frame(sce_hvg$IN_hvg_ppk_umap1, sce_hvg$IN_hvg_ppk_umap2))
colnames(umap) <- paste0("UMAP_", 1:2)
rownames(umap) <- colnames(sce_orig)
seurat_orig[["UMAP"]] <- CreateDimReducObject(
  embeddings = umap, key = "UMAP_", assay = DefaultAssay(seurat_orig))
Idents(object = seurat_orig) <- Idents(object = seurat_hvg)
IN_seurat <- seurat_orig
IN_seurat

sce_hvg <- readRDS(file.path(results_dir, "data", "12", "sce_hvg.rds"))
seurat_hvg <- readRDS(file.path(results_dir, "data", "12", "seurat_hvg.rds"))
sce_orig <- readRDS(file.path(results_dir, "data", "08", "sce_orig.rds"))
sce_orig <- calculateQCMetrics(sce_orig)
seurat_orig <- as.Seurat(sce_orig)
seurat_orig <- NormalizeData(seurat_orig)
seurat_orig <- ScaleData(seurat_orig, features = rownames(seurat_orig))

keep <- readRDS(file.path(results_dir, "data", "08", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
sce_orig <- calculateQCMetrics(sce_orig)
seurat_orig <- seurat_orig[ , keep]
seurat_orig <- NormalizeData(seurat_orig)
seurat_orig <- ScaleData(seurat_orig, features = rownames(seurat_orig))
keep <- readRDS(file.path(results_dir, "data", "08", "keep2.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "12", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]

umap <- as.matrix(data.frame(sce_hvg$EX_hvg_ppk_umap1, sce_hvg$EX_hvg_ppk_umap2))
colnames(umap) <- paste0("UMAP_", 1:2)
rownames(umap) <- colnames(sce_orig)
seurat_orig[["UMAP"]] <- CreateDimReducObject(
  embeddings = umap, key = "UMAP_", assay = DefaultAssay(seurat_orig))
Idents(object = seurat_orig) <- Idents(object = seurat_hvg)
EX_seurat <- seurat_orig
rm(sce_hvg, sce_orig, seurat_hvg, seurat_orig)
EX_seurat
```

## IN

```{r}
PVALB_seurat <- subset(x = IN_seurat, idents = "0")
PVALB_mat <- as.matrix(GetAssayData(object = PVALB_seurat, slot = "scale.data"))
dim(PVALB_mat)
CALB2_seurat <- subset(x = IN_seurat, idents = "1")
CALB2_mat <- as.matrix(GetAssayData(object = CALB2_seurat, slot = "scale.data"))
dim(CALB2_mat)
comb_mat <- cbind(PVALB_mat, CALB2_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(PVALB_mat)[2], dim(CALB2_mat)[2])))
levels(groups) <- c("PVALB", "CALB2")
design <- model.matrix(~ groups)
colnames(design) <- c("PVALB", "CALB2")
names(color_legend) <- c("PVALB", "CALB2")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "CALB2")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_whole_norm_IN_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "CALB2")

IN_fit <- fit

DE_gene_sets <- topTable(fit, coef = "CALB2", number = Inf, p.value = 1e-25, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## EX

```{r}
TBR1_seurat <- subset(x = EX_seurat, idents = "1")
TBR1_mat <- as.matrix(GetAssayData(object = TBR1_seurat, slot = "scale.data"))
dim(TBR1_mat)
SATB2_seurat <- subset(x = EX_seurat, idents = "5")
SATB2_mat <- as.matrix(GetAssayData(object = SATB2_seurat, slot = "scale.data"))
dim(SATB2_mat)
comb_mat <- cbind(TBR1_mat, SATB2_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(TBR1_mat)[2], dim(SATB2_mat)[2])))
levels(groups) <- c("TBR1", "SATB2")
design <- model.matrix(~ groups)
colnames(design) <- c("TBR1", "SATB2")
names(color_legend) <- c("TBR1", "SATB2")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "SATB2")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_whole_norm_EX_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "SATB2")

EX_fit <- fit

DE_gene_sets <- topTable(fit, coef = "SATB2", number = Inf, p.value = 1e-25, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Combined, EX Coef

```{r}
IN_sub_seurat <- subset(x = IN_seurat, idents = c("0", "1"))
IN_mat <- as.matrix(GetAssayData(object = IN_sub_seurat, slot = "scale.data"))
dim(IN_mat)
EX_sub_seurat <- subset(x = EX_seurat, idents = c("1", "5"))
EX_mat <- as.matrix(GetAssayData(object = EX_sub_seurat, slot = "scale.data"))
dim(EX_mat)
comb_mat <- cbind(IN_mat, EX_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(EX_mat)[2], dim(IN_mat)[2])))
levels(groups) <- c("IN", "EX")
design <- model.matrix(~ groups)
colnames(design) <- c("IN", "EX")
names(color_legend) <- c("IN", "EX")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "EX")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_whole_norm_comb_EX_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "EX")

EX_comb_fit <- fit

DE_gene_sets <- topTable(fit, coef = "EX", number = Inf, p.value = 1e-100, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Combined, IN Coef

```{r}
comb_mat <- cbind(EX_mat, IN_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(EX_mat)[2], dim(IN_mat)[2])))
levels(groups) <- c("EX", "IN")
design <- model.matrix(~ groups)
colnames(design) <- c("EX", "IN")
names(color_legend) <- c("EX", "IN")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "IN")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_whole_norm_comb_IN_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "IN")

IN_comb_fit <- fit

DE_gene_sets <- topTable(fit, coef = "IN", number = Inf, p.value = 1e-100, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Between Vulnerability Markers

```{r}
head(
  topTable(
    IN_comb_fit, coef = "IN", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
head(
  topTable(
    EX_comb_fit, coef = "EX", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
```

## Within Vulnerability Markers

```{r}
tail(
  topTable(
    EX_fit, coef = "SATB2", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
tail(
  topTable(
    IN_fit, coef = "CALB2", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
```

# Neuronal Normalization

```{r}
sce_hvg <- readRDS(file.path(results_dir, "data", "13", "sce_hvg.rds"))
seurat_hvg <- readRDS(file.path(results_dir, "data", "13", "seurat_hvg.rds"))
sce_orig <- readRDS(file.path(results_dir, "data", "08", "sce_orig.rds"))
seurat_orig <- as.Seurat(sce_orig)

keep <- readRDS(file.path(results_dir, "data", "08", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
sce_orig <- calculateQCMetrics(sce_orig)
seurat_orig <- seurat_orig[ , keep]
seurat_orig <- NormalizeData(seurat_orig)
seurat_orig <- ScaleData(seurat_orig, features = rownames(seurat_orig))
keep <- readRDS(file.path(results_dir, "data", "08", "keep2.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "13", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]

umap <- as.matrix(data.frame(sce_hvg$IN_hvg_ppk_umap1, sce_hvg$IN_hvg_ppk_umap2))
colnames(umap) <- paste0("UMAP_", 1:2)
rownames(umap) <- colnames(sce_orig)
seurat_orig[["UMAP"]] <- CreateDimReducObject(
  embeddings = umap, key = "UMAP_", assay = DefaultAssay(seurat_orig))
Idents(object = seurat_orig) <- Idents(object = seurat_hvg)
IN_seurat <- seurat_orig
IN_seurat

sce_hvg <- readRDS(file.path(results_dir, "data", "12", "sce_hvg.rds"))
seurat_hvg <- readRDS(file.path(results_dir, "data", "12", "seurat_hvg.rds"))
sce_orig <- readRDS(file.path(results_dir, "data", "08", "sce_orig.rds"))
seurat_orig <- as.Seurat(sce_orig)

keep <- readRDS(file.path(results_dir, "data", "08", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
sce_orig <- calculateQCMetrics(sce_orig)
seurat_orig <- seurat_orig[ , keep]
seurat_orig <- NormalizeData(seurat_orig)
seurat_orig <- ScaleData(seurat_orig, features = rownames(seurat_orig))
keep <- readRDS(file.path(results_dir, "data", "08", "keep2.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "12", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]

umap <- as.matrix(data.frame(sce_hvg$EX_hvg_ppk_umap1, sce_hvg$EX_hvg_ppk_umap2))
colnames(umap) <- paste0("UMAP_", 1:2)
rownames(umap) <- colnames(sce_orig)
seurat_orig[["UMAP"]] <- CreateDimReducObject(
  embeddings = umap, key = "UMAP_", assay = DefaultAssay(seurat_orig))
Idents(object = seurat_orig) <- Idents(object = seurat_hvg)
EX_seurat <- seurat_orig
rm(sce_hvg, sce_orig, seurat_hvg, seurat_orig)
EX_seurat
```

## IN

```{r}
PVALB_seurat <- subset(x = IN_seurat, idents = "0")
PVALB_mat <- as.matrix(GetAssayData(object = PVALB_seurat, slot = "scale.data"))
dim(PVALB_mat)
CALB2_seurat <- subset(x = IN_seurat, idents = "1")
CALB2_mat <- as.matrix(GetAssayData(object = CALB2_seurat, slot = "scale.data"))
dim(CALB2_mat)
comb_mat <- cbind(PVALB_mat, CALB2_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(PVALB_mat)[2], dim(CALB2_mat)[2])))
levels(groups) <- c("PVALB", "CALB2")
design <- model.matrix(~ groups)
colnames(design) <- c("PVALB", "CALB2")
names(color_legend) <- c("PVALB", "CALB2")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "CALB2")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_norm_IN_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "CALB2")

IN_fit <- fit

DE_gene_sets <- topTable(fit, coef = "CALB2", number = Inf, p.value = 1e-25, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## EX

```{r}
TBR1_seurat <- subset(x = EX_seurat, idents = "1")
TBR1_mat <- as.matrix(GetAssayData(object = TBR1_seurat, slot = "scale.data"))
dim(TBR1_mat)
SATB2_seurat <- subset(x = EX_seurat, idents = "5")
SATB2_mat <- as.matrix(GetAssayData(object = SATB2_seurat, slot = "scale.data"))
dim(SATB2_mat)
comb_mat <- cbind(TBR1_mat, SATB2_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(TBR1_mat)[2], dim(SATB2_mat)[2])))
levels(groups) <- c("TBR1", "SATB2")
design <- model.matrix(~ groups)
colnames(design) <- c("TBR1", "SATB2")
names(color_legend) <- c("TBR1", "SATB2")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "SATB2")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_norm_EX_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "SATB2")

EX_fit <- fit

DE_gene_sets <- topTable(fit, coef = "SATB2", number = Inf, p.value = 1e-25, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Combined, EX Coef

```{r}
IN_sub_seurat <- subset(x = IN_seurat, idents = c("0", "1"))
IN_mat <- as.matrix(GetAssayData(object = IN_sub_seurat, slot = "scale.data"))
dim(IN_mat)
EX_sub_seurat <- subset(x = EX_seurat, idents = c("1", "5"))
EX_mat <- as.matrix(GetAssayData(object = EX_sub_seurat, slot = "scale.data"))
dim(EX_mat)
comb_mat <- cbind(IN_mat, EX_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(EX_mat)[2], dim(IN_mat)[2])))
levels(groups) <- c("IN", "EX")
design <- model.matrix(~ groups)
colnames(design) <- c("IN", "EX")
names(color_legend) <- c("IN", "EX")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "EX")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_norm_comb_EX_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "EX")

EX_comb_fit <- fit

DE_gene_sets <- topTable(fit, coef = "EX", number = Inf, p.value = 1e-50, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Combined, IN Coef

```{r}
comb_mat <- cbind(EX_mat, IN_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(EX_mat)[2], dim(IN_mat)[2])))
levels(groups) <- c("EX", "IN")
design <- model.matrix(~ groups)
colnames(design) <- c("EX", "IN")
names(color_legend) <- c("EX", "IN")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "IN")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_norm_comb_IN_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "IN")

IN_comb_fit <- fit

DE_gene_sets <- topTable(fit, coef = "IN", number = Inf, p.value = 1e-50, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Between Vulnerability Markers

```{r}
head(
  topTable(
    IN_comb_fit, coef = "IN", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
head(
  topTable(
    EX_comb_fit, coef = "EX", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
```

## Within Vulnerability Markers

```{r}
tail(
  topTable(
    EX_fit, coef = "SATB2", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
tail(
  topTable(
    IN_fit, coef = "CALB2", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
```

# Neuronal Subset Normalization

```{r}
sce_hvg <- readRDS(file.path(results_dir, "data", "13", "sce_hvg.rds"))
seurat_hvg <- readRDS(file.path(results_dir, "data", "13", "seurat_hvg.rds"))
sce_orig <- readRDS(file.path(results_dir, "data", "08", "sce_orig.rds"))
seurat_orig <- as.Seurat(sce_orig)

keep <- readRDS(file.path(results_dir, "data", "08", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "08", "keep2.rds"))
sce_orig <- sce_orig[ , keep]
sce_orig <- calculateQCMetrics(sce_orig)
seurat_orig <- seurat_orig[ , keep]
seurat_orig <- NormalizeData(seurat_orig)
seurat_orig <- ScaleData(seurat_orig, features = rownames(seurat_orig))
keep <- readRDS(file.path(results_dir, "data", "13", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]

umap <- as.matrix(data.frame(sce_hvg$IN_hvg_ppk_umap1, sce_hvg$IN_hvg_ppk_umap2))
colnames(umap) <- paste0("UMAP_", 1:2)
rownames(umap) <- colnames(sce_orig)
seurat_orig[["UMAP"]] <- CreateDimReducObject(
  embeddings = umap, key = "UMAP_", assay = DefaultAssay(seurat_orig))
Idents(object = seurat_orig) <- Idents(object = seurat_hvg)
IN_seurat <- seurat_orig
IN_seurat

sce_hvg <- readRDS(file.path(results_dir, "data", "12", "sce_hvg.rds"))
seurat_hvg <- readRDS(file.path(results_dir, "data", "12", "seurat_hvg.rds"))
sce_orig <- readRDS(file.path(results_dir, "data", "08", "sce_orig.rds"))
seurat_orig <- as.Seurat(sce_orig)

keep <- readRDS(file.path(results_dir, "data", "08", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "08", "keep2.rds"))
sce_orig <- sce_orig[ , keep]
sce_orig <- calculateQCMetrics(sce_orig)
seurat_orig <- seurat_orig[ , keep]
seurat_orig <- NormalizeData(seurat_orig)
seurat_orig <- ScaleData(seurat_orig, features = rownames(seurat_orig))
keep <- readRDS(file.path(results_dir, "data", "12", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]

umap <- as.matrix(data.frame(sce_hvg$EX_hvg_ppk_umap1, sce_hvg$EX_hvg_ppk_umap2))
colnames(umap) <- paste0("UMAP_", 1:2)
rownames(umap) <- colnames(sce_orig)
seurat_orig[["UMAP"]] <- CreateDimReducObject(
  embeddings = umap, key = "UMAP_", assay = DefaultAssay(seurat_orig))
Idents(object = seurat_orig) <- Idents(object = seurat_hvg)
EX_seurat <- seurat_orig
rm(sce_hvg, sce_orig, seurat_hvg, seurat_orig)
EX_seurat
```

## IN

```{r}
PVALB_seurat <- subset(x = IN_seurat, idents = "0")
PVALB_mat <- as.matrix(GetAssayData(object = PVALB_seurat, slot = "scale.data"))
dim(PVALB_mat)
CALB2_seurat <- subset(x = IN_seurat, idents = "1")
CALB2_mat <- as.matrix(GetAssayData(object = CALB2_seurat, slot = "scale.data"))
dim(CALB2_mat)
comb_mat <- cbind(PVALB_mat, CALB2_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(PVALB_mat)[2], dim(CALB2_mat)[2])))
levels(groups) <- c("PVALB", "CALB2")
design <- model.matrix(~ groups)
colnames(design) <- c("PVALB", "CALB2")
names(color_legend) <- c("PVALB", "CALB2")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "CALB2")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_subset_norm_IN_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "CALB2")

IN_fit <- fit

DE_gene_sets <- topTable(fit, coef = "CALB2", number = Inf, p.value = 1e-25, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## EX

```{r}
TBR1_seurat <- subset(x = EX_seurat, idents = "1")
TBR1_mat <- as.matrix(GetAssayData(object = TBR1_seurat, slot = "scale.data"))
dim(TBR1_mat)
SATB2_seurat <- subset(x = EX_seurat, idents = "5")
SATB2_mat <- as.matrix(GetAssayData(object = SATB2_seurat, slot = "scale.data"))
dim(SATB2_mat)
comb_mat <- cbind(TBR1_mat, SATB2_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(TBR1_mat)[2], dim(SATB2_mat)[2])))
levels(groups) <- c("TBR1", "SATB2")
design <- model.matrix(~ groups)
colnames(design) <- c("TBR1", "SATB2")
names(color_legend) <- c("TBR1", "SATB2")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "SATB2")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_subset_norm_EX_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "SATB2")

EX_fit <- fit

DE_gene_sets <- topTable(fit, coef = "SATB2", number = Inf, p.value = 1e-25, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Combined, EX Coef

```{r}
IN_sub_seurat <- subset(x = IN_seurat, idents = c("0", "1"))
IN_mat <- as.matrix(GetAssayData(object = IN_sub_seurat, slot = "scale.data"))
dim(IN_mat)
EX_sub_seurat <- subset(x = EX_seurat, idents = c("1", "5"))
EX_mat <- as.matrix(GetAssayData(object = EX_sub_seurat, slot = "scale.data"))
dim(EX_mat)
comb_mat <- cbind(IN_mat, EX_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(EX_mat)[2], dim(IN_mat)[2])))
levels(groups) <- c("IN", "EX")
design <- model.matrix(~ groups)
colnames(design) <- c("IN", "EX")
names(color_legend) <- c("IN", "EX")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "EX")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_subset_norm_comb_EX_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "EX")

EX_comb_fit <- fit

DE_gene_sets <- topTable(fit, coef = "EX", number = Inf, p.value = 1e-50, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Combined, IN Coef

```{r}
comb_mat <- cbind(EX_mat, IN_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(EX_mat)[2], dim(IN_mat)[2])))
levels(groups) <- c("EX", "IN")
design <- model.matrix(~ groups)
colnames(design) <- c("EX", "IN")
names(color_legend) <- c("EX", "IN")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "IN")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_subset_norm_comb_IN_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "IN")

IN_comb_fit <- fit

DE_gene_sets <- topTable(fit, coef = "IN", number = Inf, p.value = 1e-50, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Between Vulnerability Markers

```{r}
head(
  topTable(
    IN_comb_fit, coef = "IN", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
head(
  topTable(
    EX_comb_fit, coef = "EX", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
```

## Within Vulnerability Markers

```{r}
tail(
  topTable(
    EX_fit, coef = "SATB2", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
tail(
  topTable(
    IN_fit, coef = "CALB2", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
```

# Neuronal Subset Subpopulation Normalization

```{r}
sce_hvg <- readRDS(file.path(results_dir, "data", "13", "sce_hvg.rds"))
seurat_hvg <- readRDS(file.path(results_dir, "data", "13", "seurat_hvg.rds"))
sce_orig <- readRDS(file.path(results_dir, "data", "08", "sce_orig.rds"))
seurat_orig <- as.Seurat(sce_orig)

keep <- readRDS(file.path(results_dir, "data", "08", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "08", "keep2.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "13", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
sce_orig <- calculateQCMetrics(sce_orig)
seurat_orig <- seurat_orig[ , keep]
seurat_orig <- NormalizeData(seurat_orig)
seurat_orig <- ScaleData(seurat_orig, features = rownames(seurat_orig))

umap <- as.matrix(data.frame(sce_hvg$IN_hvg_ppk_umap1, sce_hvg$IN_hvg_ppk_umap2))
colnames(umap) <- paste0("UMAP_", 1:2)
rownames(umap) <- colnames(sce_orig)
seurat_orig[["UMAP"]] <- CreateDimReducObject(
  embeddings = umap, key = "UMAP_", assay = DefaultAssay(seurat_orig))
Idents(object = seurat_orig) <- Idents(object = seurat_hvg)
IN_seurat <- seurat_orig
IN_seurat

sce_hvg <- readRDS(file.path(results_dir, "data", "12", "sce_hvg.rds"))
seurat_hvg <- readRDS(file.path(results_dir, "data", "12", "seurat_hvg.rds"))
sce_orig <- readRDS(file.path(results_dir, "data", "08", "sce_orig.rds"))
seurat_orig <- as.Seurat(sce_orig)

keep <- readRDS(file.path(results_dir, "data", "08", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "08", "keep2.rds"))
sce_orig <- sce_orig[ , keep]
seurat_orig <- seurat_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "12", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
sce_orig <- calculateQCMetrics(sce_orig)
seurat_orig <- seurat_orig[ , keep]
seurat_orig <- NormalizeData(seurat_orig)
seurat_orig <- ScaleData(seurat_orig, features = rownames(seurat_orig))

umap <- as.matrix(data.frame(sce_hvg$EX_hvg_ppk_umap1, sce_hvg$EX_hvg_ppk_umap2))
colnames(umap) <- paste0("UMAP_", 1:2)
rownames(umap) <- colnames(sce_orig)
seurat_orig[["UMAP"]] <- CreateDimReducObject(
  embeddings = umap, key = "UMAP_", assay = DefaultAssay(seurat_orig))
Idents(object = seurat_orig) <- Idents(object = seurat_hvg)
EX_seurat <- seurat_orig
rm(sce_hvg, sce_orig, seurat_hvg, seurat_orig)
EX_seurat
```

## IN

```{r}
PVALB_seurat <- subset(x = IN_seurat, idents = "0")
PVALB_mat <- as.matrix(GetAssayData(object = PVALB_seurat, slot = "scale.data"))
dim(PVALB_mat)
CALB2_seurat <- subset(x = IN_seurat, idents = "1")
CALB2_mat <- as.matrix(GetAssayData(object = CALB2_seurat, slot = "scale.data"))
dim(CALB2_mat)
comb_mat <- cbind(PVALB_mat, CALB2_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(PVALB_mat)[2], dim(CALB2_mat)[2])))
levels(groups) <- c("PVALB", "CALB2")
design <- model.matrix(~ groups)
colnames(design) <- c("PVALB", "CALB2")
names(color_legend) <- c("PVALB", "CALB2")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "CALB2")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_subset_subpop_norm_IN_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "CALB2")

IN_fit <- fit

DE_gene_sets <- topTable(fit, coef = "CALB2", number = Inf, p.value = 1e-25, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## EX

```{r}
TBR1_seurat <- subset(x = EX_seurat, idents = "1")
TBR1_mat <- as.matrix(GetAssayData(object = TBR1_seurat, slot = "scale.data"))
dim(TBR1_mat)
SATB2_seurat <- subset(x = EX_seurat, idents = "5")
SATB2_mat <- as.matrix(GetAssayData(object = SATB2_seurat, slot = "scale.data"))
dim(SATB2_mat)
comb_mat <- cbind(TBR1_mat, SATB2_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(TBR1_mat)[2], dim(SATB2_mat)[2])))
levels(groups) <- c("TBR1", "SATB2")
design <- model.matrix(~ groups)
colnames(design) <- c("TBR1", "SATB2")
names(color_legend) <- c("TBR1", "SATB2")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "SATB2")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_subset_subpop_norm_EX_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "SATB2")

EX_fit <- fit

DE_gene_sets <- topTable(fit, coef = "SATB2", number = Inf, p.value = 1e-25, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Combined, EX Coef

```{r}
IN_sub_seurat <- subset(x = IN_seurat, idents = c("0", "1"))
IN_mat <- as.matrix(GetAssayData(object = IN_sub_seurat, slot = "scale.data"))
dim(IN_mat)
EX_sub_seurat <- subset(x = EX_seurat, idents = c("1", "5"))
EX_mat <- as.matrix(GetAssayData(object = EX_sub_seurat, slot = "scale.data"))
dim(EX_mat)
comb_mat <- cbind(IN_mat, EX_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(EX_mat)[2], dim(IN_mat)[2])))
levels(groups) <- c("IN", "EX")
design <- model.matrix(~ groups)
colnames(design) <- c("IN", "EX")
names(color_legend) <- c("IN", "EX")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "EX")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_subset_subpop_norm_comb_EX_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "EX")

EX_comb_fit <- fit

DE_gene_sets <- topTable(fit, coef = "EX", number = Inf, p.value = 1e-100, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Combined, IN Coef

```{r}
comb_mat <- cbind(EX_mat, IN_mat)

expr_genes <- rowSums(comb_mat)
expr_genes <- expr_genes[order(-expr_genes)]
expr_genes_table <- as.data.table(expr_genes)
expr_genes_table$genes <- names(expr_genes)
setcolorder(expr_genes_table, c("genes", "expr_genes"))
datatable_custom(head(expr_genes_table, n = 10))

groups <- groups <- factor(rep(c(0, 1), c(dim(EX_mat)[2], dim(IN_mat)[2])))
levels(groups) <- c("EX", "IN")
design <- model.matrix(~ groups)
colnames(design) <- c("EX", "IN")
names(color_legend) <- c("EX", "IN")
sample_color_map <- color_legend[groups]

fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
topTable(fit, coef = "IN")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "15", "gsva_neuronal_subset_subpop_norm_comb_IN_c5.bp.v7.0.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(comb_mat, gene_sets, min.sz = 3)
  saveRDS(gsva, rds)}
```

```{r}
fit <- lmFit(gsva, design)
fit <- eBayes(fit)
topTable(fit, coef = "IN")

IN_comb_fit <- fit

DE_gene_sets <- topTable(fit, coef = "IN", number = Inf, p.value = 1e-100, adjust = "BH")
gsva_cutoff <- gsva[rownames(DE_gene_sets), ]

names(sample_color_map) <- colnames(gsva_cutoff)
sample_clustering <- hclust(as.dist(1 - cor(gsva_cutoff, method = "spearman")), method = "complete")
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(gsva_cutoff), method = "pearson")), method = "complete")
heatmap(
  gsva_cutoff, ColSideColors = sample_color_map,
  xlab = "Neurons", margins = c(2, 40),
  col = color, ylab = "c5.bp.v7.0",
  labRow = substr(gsub("_", " ", gsub("^GO_", "", rownames(gsva_cutoff))), 1, 100), labCol=" ",
  scale = "row", Rowv = as.dendrogram(gene_set_clustering),
  Colv = as.dendrogram(sample_clustering))
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")
```

## Between Vulnerability Markers

```{r}
head(
  topTable(
    IN_comb_fit, coef = "IN", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
head(
  topTable(
    EX_comb_fit, coef = "EX", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
```

## Within Vulnerability Markers

```{r}
tail(
  topTable(
    EX_fit, coef = "SATB2", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
tail(
  topTable(
    IN_fit, coef = "CALB2", number = Inf,
    adjust.method = "holm", sort.by = "logFC", p.value = 0.05),
  n = 20)
```

# References

This is the concluding section of the document. Here we write relevant results to disk, output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
