---
title: "23 Habib Neuronal EX All Dimensions"
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography: "../tau-vulnerability.bib"
biblio-style: apalike
link-citations: true
output:
  html_document:
    code_folding: show
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile, encoding = encoding, output_file = "../results/23-habib-neuronal-EX-alldim.html")})
---

```{r, include = FALSE}
#    This file is part of tau-vulnerability.
#    Copyright (C) 2019  Emir Turkes
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

knitr::opts_chunk$set(fig.width = 15, fig.height = 13)
```

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of the [Tau Vulnerability Project](https://github.com/eturkes/tau-vulnerability).*

This document follows up its preparation file to provide and in-depth analysis of neuronal populations in the dataset from @habib_massively_2017.
We start by setting some global variables and loading in any required packages.

```{r}
assets_dir <- file.path(getwd(), "..", "assets")
results_dir <- file.path(getwd(), "..", "results")

packages <- c(
  "conflicted", "SingleCellExperiment", "magrittr", "dplyr", "ggplot2", "ggrepel", "S4Vectors",
  "SummarizedExperiment", "DropletUtils", "scran", "BiocSingular", "scater", "Rtsne", "svd", "SC3",
  "DT", "data.table", "Seurat", "uwot", "viridis", "WGCNA", "pheatmap", "RColorBrewer")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))

conflict_prefer("which", "BiocGenerics")
options(stringsAsFactors = FALSE)

# Create a unique cache and results data directory for each iterated section.
if (!dir.exists(file.path(assets_dir, "cache", "23"))) {
  dir.create(file.path(assets_dir, "cache", "23"), recursive = TRUE)}
if (!dir.exists(file.path(results_dir, "data", "23"))) {
  dir.create(file.path(results_dir, "data", "23"), recursive = TRUE)}

# ggplot2 function providing custom aesthetics and automatic placement of categorical labels.
# For continuous data, a colorbar is implemented.
dim_red_plot <- function(data, x, y, col, type) {
  gg <- ggplot(data, aes_string(x = x, y = y, color = col)) +
    geom_point(alpha = 0.35, stroke = 0.05, shape = 21, aes_string(fill = col)) +
    theme_classic() +
    theme(
      legend.position = "right", plot.title = element_text(hjust = 0.5),
      legend.title = element_blank()) +
    guides(color = guide_legend(override.aes = list(alpha = 1)))
    if (type == "cat") {
      gg <- gg + geom_label_repel(
        data = label_df2, aes(label = label), show.legend = FALSE, size = 3)
    } else if (type == "cont") {
      gg <- ggplot(data, aes_string(x = x, y = y)) +
        geom_point(alpha = 0.35, stroke = 0.05, aes_string(color = col)) +
        theme_classic() +
        theme(
          legend.position = "right", plot.title = element_text(hjust = 0.5),
          legend.title = element_blank()) +
        scale_colour_viridis()}
  gg}

# From SC3 source, needed to look at more marker genes than the top 10.
organise_marker_genes <- function(object, k, p_val, auroc) {
  dat <- rowData(object)[ , c(
    paste0("sc3_", k, "_markers_clusts"), paste0("sc3_", k, "_markers_auroc"),
    paste0("sc3_", k, "_markers_padj"), "feature_symbol")]
  dat <- dat[dat[ , paste0("sc3_", k, "_markers_padj")] < p_val & !is.na(
    dat[, paste0("sc3_", k, "_markers_padj")]), ]
  dat <- dat[dat[ , paste0("sc3_", k, "_markers_auroc")] > auroc, ]
  d <- NULL
  for (i in sort(unique(dat[, paste0("sc3_", k, "_markers_clusts")]))) {
    tmp <- dat[dat[, paste0("sc3_", k, "_markers_clusts")] == i, ]
    tmp <- tmp[order(tmp[, paste0("sc3_", k, "_markers_auroc")], decreasing = TRUE), ]
    d <- rbind(d, tmp)}
  if (nrow(dat) > 0) {
    return(d)
  } else {
    return(NULL)}}

# From SC3 source, needed to look at more DE genes than the top 50.
organise_de_genes <- function(object, k, p_val) {
  de_genes <- rowData(object)[, paste0("sc3_", k, "_de_padj")]
  names(de_genes) <- rowData(object)$feature_symbol
  de_genes <- de_genes[!is.na(de_genes)]
  de_genes <- de_genes[de_genes < p_val]
  de_genes <- de_genes[order(de_genes)]
  return(de_genes)}

# Adds download buttons.
datatable_custom <- function(dt) {
  datatable(
    dt,
    extensions = "Buttons", options = list(dom = "Blfrtip", buttons = list(
      "copy", "print",
      list(extend = "collection", buttons = c("csv", "excel", "pdf"), text = "Download"))))}
```

We load in the data from the previous analysis.

```{r}
sce <- readRDS(file.path(results_dir, "data", "08", "sce.rds"))
sce
```

# Subsetting to EX

```{r}
# Set up data frame for ggplot2.
gg_df <- data.frame(
  colData(sce)[ , c(
    "cell_id_stem", paste0("habib_tsne", 1:2), "log10_total_features_by_counts",
    "habib_cluster_name", "hvg_seurat_clusters_res_name", paste0("sub_hvg_ppk_umap", 1:2))])
rownames(gg_df) <- NULL
gg_df$habib_cluster_name <- factor(gg_df$habib_cluster_name)

# Setup for automatic placement of cluster labels.
label_df <- data.frame(
  hvg_seurat_clusters_res_name = levels(gg_df$hvg_seurat_clusters_res_name),
  label = levels(gg_df$hvg_seurat_clusters_res_name))
label_df2 <- gg_df %>%
  group_by(hvg_seurat_clusters_res_name) %>%
  summarize(
    sub_hvg_ppk_umap1 = median(sub_hvg_ppk_umap1),
    sub_hvg_ppk_umap2 = median(sub_hvg_ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "sub_hvg_ppk_umap1", y = "sub_hvg_ppk_umap2",
  col = "hvg_seurat_clusters_res_name", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") +
  ggtitle("Subset HVG Seurat ~1.3 Resolution Clusters Subset HVG PPK UMAP") + theme_linedraw()
```

```{r}
keep <- sce$sub_hvg_ppk_umap1 < 2.5
saveRDS(keep, file.path(results_dir, "data", "23", "keep1.rds"))
sce <- sce[ , keep]
sce
```

We remove genes that are now lowly expressed after the subset.

```{r}
sce <- calculateQCMetrics(sce)
detected_genes <- rowSums(counts(sce) >= 2)
sce <- sce[which(detected_genes >= 10), ]
dim(sce)[1]
```

Due to the change in dimensions from subsetting, we renormalize.

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "sub_quickCluster.rds")
if (file.exists(rds)) {
  quickCluster <- readRDS(rds)
} else {
  set.seed(1)
  quickCluster <- quickCluster(sce, use.ranks = FALSE, BSPARAM = IrlbaParam())
  saveRDS(quickCluster, rds)}
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "sub_size_factors.rds")
if (file.exists(rds)) {
  size_factors <- readRDS(rds)
} else {
  size_factors <- computeSumFactors(sce, cluster = quickCluster, min.mean = 0.1)
  saveRDS(size_factors, rds)}
```

```{r}
sce <- normalize(sce)
```

And we view the subsetted UMAP figure.

```{r}
# Set up data frame for ggplot2.
gg_df <- data.frame(
  colData(sce)[ , c(
    "cell_id_stem", paste0("habib_tsne", 1:2), "log10_total_features_by_counts",
    "habib_cluster_name", "hvg_seurat_clusters_res_name", paste0("sub_hvg_ppk_umap", 1:2))])
rownames(gg_df) <- NULL
gg_df$habib_cluster_name <- factor(gg_df$habib_cluster_name)

# Setup for automatic placement of cluster labels.
label_df <- data.frame(
  hvg_seurat_clusters_res_name = levels(gg_df$hvg_seurat_clusters_res_name),
  label = levels(gg_df$hvg_seurat_clusters_res_name))
label_df2 <- gg_df %>%
  group_by(hvg_seurat_clusters_res_name) %>%
  summarize(
    sub_hvg_ppk_umap1 = median(sub_hvg_ppk_umap1),
    sub_hvg_ppk_umap2 = median(sub_hvg_ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "sub_hvg_ppk_umap1", y = "sub_hvg_ppk_umap2",
  col = "hvg_seurat_clusters_res_name", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") +
  ggtitle("Subset HVG Seurat ~1.3 Resolution Clusters Subset HVG PPK UMAP") + theme_linedraw()
```

## PPK SVD + UMAP

We repeat our preferred dimensionality reduction method from earlier.

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "ppk.rds")
if (file.exists(rds)) {
  ppk <- readRDS(rds)
} else {
  set.seed(1)
  ppk <- propack.svd(scale(t(as.matrix(logcounts(sce)))), neig = 50)
  saveRDS(ppk, rds)}
```

```{r}
pca <- t(ppk$d * t(ppk$u))
retain_pcs <- 21
retain_pcs
par(mfrow = c(1, 1))
plot(
  log10(ppk$d), xlab = "PC", ylab = "log10(Proportion of Variance Explained)",
  pch = 20, cex = 0.6, col = rgb(0.8, 0.2, 0.2, 0.5))
abline(v = retain_pcs, lty = 2, col = "red")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "ppk_umap.rds")
if (file.exists(rds)) {
  umap <- readRDS(rds)
} else {
  set.seed(0)
  umap <- umap(pca[ , 1:retain_pcs], min_dist = 0.2)
  saveRDS(umap, rds)}
```

```{r}
reducedDims(sce) <- SimpleList(PCA = pca, UMAP = umap)
add_df <- data.frame(reducedDim(sce, "UMAP"))
names(add_df) <- paste0("EX_ppk_umap", seq(ncol(add_df)))
colData(sce) <- cbind(
  colData(sce), EX_ppk_umap1 = add_df$EX_ppk_umap1,
  EX_ppk_umap2 = add_df$EX_ppk_umap2)
gg_df <- data.frame(gg_df, add_df)
rm(pca, add_df)

label_df2 <- gg_df %>%
  group_by(hvg_seurat_clusters_res_name) %>%
  summarize(
    EX_ppk_umap1 = median(EX_ppk_umap1), EX_ppk_umap2 = median(EX_ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
  col = "hvg_seurat_clusters_res_name", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") + ggtitle("Whole Set Seurat Clusters Subset PPK UMAP")
dim_red_plot(
  data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
  col = "cell_id_stem", type = "other") +
  xlab("UMAP 1") + ylab("UMAP 2") + ggtitle("Sample Distribution Subset PPK UMAP")
dim_red_plot(
  data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
  col = "log10_total_features_by_counts", type = "cont") +
  xlab("UMAP 1") + ylab("UMAP 2") + ggtitle("log10(Total Features) Subset PPK UMAP")
```

# Clustering

## Seurat

Seurat is a nearest neighbor graph clustering method with a `resolution` parameter that indirectly influences the number of clusters generated.
Concerning `resolution`, the authors write:

> setting this parameter between 0.4-1.2 typically returns good results for single-cell datasets of around 3K cells.
Optimal resolution often increases for larger datasets.

We try scaling the median of these recommended settings (0.8, which is also the default) to the size of our dataset.
As Seurat typically uses HVGs, we also use `sce` as input.

```{r}
resolution <- (dim(sce)[2] / 3000) * 0.8
resolution
```

We cluster in the range of $\pm 0.2$ from this resolution.

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "seurat_res_minus0.2.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- as.Seurat(sce, counts = "counts", data = NULL)
  seurat <- NormalizeData(seurat)
  seurat <- FindVariableFeatures(seurat, nfeatures = dim(sce)[1])
  seurat <- ScaleData(seurat, features = rownames(seurat))
  seurat <- FindNeighbors(seurat, reduction = "PCA", dims = 1:retain_pcs)
  seurat <- FindClusters(seurat, resolution = resolution - 0.2)
  saveRDS(seurat, rds)}
```

```{r}
gg_df$seurat_clusters_res_minus0.2 <- seurat$seurat_clusters

# Setup for automatic placement of cluster labels.
label_df <- data.frame(
  seurat_clusters_res_minus0.2 = levels(gg_df$seurat_clusters_res_minus0.2),
  label = levels(gg_df$seurat_clusters_res_minus0.2))
label_df2 <- gg_df %>%
  group_by(seurat_clusters_res_minus0.2) %>%
  summarize(
    EX_ppk_umap1 = median(EX_ppk_umap1), EX_ppk_umap2 = median(EX_ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
  col = "seurat_clusters_res_minus0.2", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") +
  ggtitle("Subset Seurat ~0.7 Resolution Clusters Subset PPK UMAP")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "seurat_res_minus0.1.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- as.Seurat(sce, counts = "counts", data = NULL)
  seurat <- NormalizeData(seurat)
  seurat <- FindVariableFeatures(seurat, nfeatures = dim(sce)[1])
  seurat <- ScaleData(seurat, features = rownames(seurat))
  seurat <- FindNeighbors(seurat, reduction = "PCA", dims = 1:retain_pcs)
  seurat <- FindClusters(seurat, resolution = resolution - 0.1)
  saveRDS(seurat, rds)}
```

```{r}
gg_df$seurat_clusters_res_minus0.1 <- seurat$seurat_clusters

# Setup for automatic placement of cluster labels.
label_df <- data.frame(
  seurat_clusters_res_minus0.1 = levels(gg_df$seurat_clusters_res_minus0.1),
  label = levels(gg_df$seurat_clusters_res_minus0.1))
label_df2 <- gg_df %>%
  group_by(seurat_clusters_res_minus0.1) %>%
  summarize(
    EX_ppk_umap1 = median(EX_ppk_umap1), EX_ppk_umap2 = median(EX_ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
  col = "seurat_clusters_res_minus0.1", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") +
  ggtitle("Subset Seurat ~0.8 Resolution Clusters Subset PPK UMAP")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "seurat_res.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- as.Seurat(sce, counts = "counts", data = NULL)
  seurat <- NormalizeData(seurat)
  seurat <- FindVariableFeatures(seurat, nfeatures = dim(sce)[1])
  seurat <- ScaleData(seurat, features = rownames(seurat))
  seurat <- FindNeighbors(seurat, reduction = "PCA", dims = 1:retain_pcs)
  seurat <- FindClusters(seurat, resolution = resolution)
  saveRDS(seurat, rds)}
```

```{r}
gg_df$seurat_clusters_res <- seurat$seurat_clusters

# Setup for automatic placement of cluster labels.
label_df <- data.frame(
  seurat_clusters_res = levels(gg_df$seurat_clusters_res),
  label = levels(gg_df$seurat_clusters_res))
label_df2 <- gg_df %>%
  group_by(seurat_clusters_res) %>%
  summarize(
    EX_ppk_umap1 = median(EX_ppk_umap1), EX_ppk_umap2 = median(EX_ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
  col = "seurat_clusters_res", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") +
  ggtitle("Subset Seurat ~0.9 Resolution Clusters Subset PPK UMAP")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "seurat_res_plus0.1.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- as.Seurat(sce, counts = "counts", data = NULL)
  seurat <- NormalizeData(seurat)
  seurat <- FindVariableFeatures(seurat, nfeatures = dim(sce)[1])
  seurat <- ScaleData(seurat, features = rownames(seurat))
  seurat <- FindNeighbors(seurat, reduction = "PCA", dims = 1:retain_pcs)
  seurat <- FindClusters(seurat, resolution = resolution + 0.1)
  saveRDS(seurat, rds)}
```

```{r}
gg_df$seurat_clusters_res_plus0.1 <- seurat$seurat_clusters

# Setup for automatic placement of cluster labels.
label_df <- data.frame(
  seurat_clusters_res_plus0.1 = levels(gg_df$seurat_clusters_res_plus0.1),
  label = levels(gg_df$seurat_clusters_res_plus0.1))
label_df2 <- gg_df %>%
  group_by(seurat_clusters_res_plus0.1) %>%
  summarize(
    EX_ppk_umap1 = median(EX_ppk_umap1), EX_ppk_umap2 = median(EX_ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
  col = "seurat_clusters_res_plus0.1", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") +
  ggtitle("Subset Seurat ~1.0 Resolution Clusters Subset PPK UMAP")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "seurat_res_plus0.2.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- as.Seurat(sce, counts = "counts", data = NULL)
  seurat <- NormalizeData(seurat)
  seurat <- FindVariableFeatures(seurat, nfeatures = dim(sce)[1])
  seurat <- ScaleData(seurat, features = rownames(seurat))
  seurat <- FindNeighbors(seurat, reduction = "PCA", dims = 1:retain_pcs)
  seurat <- FindClusters(seurat, resolution = resolution + 0.2)
  saveRDS(seurat, rds)}
```

```{r}
gg_df$seurat_clusters_res_plus0.2 <- seurat$seurat_clusters

# Setup for automatic placement of cluster labels.
label_df <- data.frame(
  seurat_clusters_res_plus0.2 = levels(gg_df$seurat_clusters_res_plus0.2),
  label = levels(gg_df$seurat_clusters_res_plus0.2))
label_df2 <- gg_df %>%
  group_by(seurat_clusters_res_plus0.2) %>%
  summarize(
    EX_ppk_umap1 = median(EX_ppk_umap1), EX_ppk_umap2 = median(EX_ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
  col = "seurat_clusters_res_plus0.2", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") +
  ggtitle("Subset Seurat ~1.1 Resolution Clusters Subset PPK UMAP")
```

We also try at 2 and 3 times the median of the recommended settings.

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "seurat_res_times2.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- as.Seurat(sce, counts = "counts", data = NULL)
  seurat <- NormalizeData(seurat)
  seurat <- FindVariableFeatures(seurat, nfeatures = dim(sce)[1])
  seurat <- ScaleData(seurat, features = rownames(seurat))
  seurat <- FindNeighbors(seurat, reduction = "PCA", dims = 1:retain_pcs)
  seurat <- FindClusters(seurat, resolution = resolution * 2)
  saveRDS(seurat, rds)}
```

```{r}
gg_df$seurat_clusters_res_times2 <- seurat$seurat_clusters

# Setup for automatic placement of cluster labels.
label_df <- data.frame(
  seurat_clusters_res_times2 = levels(gg_df$seurat_clusters_res_times2),
  label = levels(gg_df$seurat_clusters_res_times2))
label_df2 <- gg_df %>%
  group_by(seurat_clusters_res_times2) %>%
  summarize(
    EX_ppk_umap1 = median(EX_ppk_umap1), EX_ppk_umap2 = median(EX_ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
  col = "seurat_clusters_res_times2", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") +
  ggtitle("Subset Seurat ~1.8 Resolution Clusters Subset PPK UMAP")
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "seurat_res_times3.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- as.Seurat(sce, counts = "counts", data = NULL)
  seurat <- NormalizeData(seurat)
  seurat <- FindVariableFeatures(seurat, nfeatures = dim(sce)[1])
  seurat <- ScaleData(seurat, features = rownames(seurat))
  seurat <- FindNeighbors(seurat, reduction = "PCA", dims = 1:retain_pcs)
  seurat <- FindClusters(seurat, resolution = resolution * 3)
  saveRDS(seurat, rds)}
```

```{r}
gg_df$seurat_clusters_res_times3 <- seurat$seurat_clusters

# Setup for automatic placement of cluster labels.
label_df <- data.frame(
  seurat_clusters_res_times3 = levels(gg_df$seurat_clusters_res_times3),
  label = levels(gg_df$seurat_clusters_res_times3))
label_df2 <- gg_df %>%
  group_by(seurat_clusters_res_times3) %>%
  summarize(
    EX_ppk_umap1 = median(EX_ppk_umap1), EX_ppk_umap2 = median(EX_ppk_umap2)) %>%
  left_join(label_df)

dim_red_plot(
  data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
  col = "seurat_clusters_res_times3", type = "cat") +
  xlab("UMAP 1") + ylab("UMAP 2") +
  ggtitle("Subset Seurat ~2.7 Resolution Clusters Subset PPK UMAP")
```

We see that unlike inhibitory neurons, cluster count continues to increase with increasing resolution.
We also observe cluster stability across resolutions at 9 clusters.
Therefore, we choose to go forward with clusters at the recommended resolution, as is done with the inhibitory neuron set.

```{r}
seurat <- readRDS(file.path(assets_dir, "cache", "23", "seurat_res.rds"))
```

## SC3

As an alternative, we try clustering using SC3.
First, we estimate the optimal $k$ for our dataset.

```{r}
rowData(sce)$feature_symbol = rowData(sce)$external_gene_name
```

```{r, cache = TRUE}
rds <- file.path(assets_dir, "cache", "23", "sc3_estimate_k.rds")
if (file.exists(rds)) {
  sc3_estimate_k <- readRDS(rds)
} else {
  set.seed(1)
  sc3_estimate_k <- sc3_estimate_k(sce)
  saveRDS(sc3_estimate_k, rds)}
```

```{r}
sce <- sc3_estimate_k
rm(sc3_estimate_k)
metadata(sce)$sc3$k_estimation
```

SC3's optimal $k$ seems large compared to Seurat's suggestions, so we generate 7-12 clusters to capture those with Seurat with a margin of error.

```{r}
num_clust <- c(7:12)
counts(sce) <- as.matrix(counts(sce))
logcounts(sce) <- as.matrix(logcounts(sce))
```

```{r}
# Knitr caching turned off due to:
# Error in lazyLoadDBinsertVariable(vars[i], from, datafile, ascii, compress,  :
# long vectors not supported yet
rds <- file.path(assets_dir, "cache", "23", "sc3.rds")
if (file.exists(rds)) {
  sc3 <- readRDS(rds)
} else {
  sc3 <- sc3(sce, ks = num_clust, gene_filter = FALSE, biology = TRUE, n_cores = 8)
  saveRDS(sc3, rds)}
```

```{r}
sce <- sc3
rm(sc3)

for (i in num_clust) {
  sc3_label <- paste0("sc3_", i, "_clusters")
  sc3_label <- paste0("sc3_", i, "_clusters")
  gg_df[[sc3_label]] <- factor(colData(sce)[ , sc3_label])}
clusts <- c(paste0("sc3_", num_clust, "_clusters"))
clusts_nice <- c(paste0("SC3 ", num_clust, " Clusters"))

for (i in 1:length(clusts)) {
  # Setup for automatic placement of cluster labels.
  label_df <- data.frame(tmp = levels(gg_df[[clusts[i]]]), label = levels(gg_df[[clusts[i]]]))
  names(label_df)[names(label_df) == "tmp"] <- clusts[i]
  label_df2 <- gg_df %>%
    group_by_at(clusts[i]) %>%
    summarize(
      EX_ppk_umap1 = median(EX_ppk_umap1),
      EX_ppk_umap2 = median(EX_ppk_umap2)) %>%
    left_join(label_df)

  print(dim_red_plot(
    data = gg_df, x = "EX_ppk_umap1", y = "EX_ppk_umap2",
    col = clusts[i], type = "cat") +
      xlab("UMAP 1") + ylab("UMAP 2") +
      ggtitle(paste("Subset", clusts_nice[i], "Subset PPK UMAP")))}
```

# IHC Genes

Next, we examine the expression of a set genes for which we have antibodies to conduct immunohistochemistry.

```{r}
sce_orig <- readRDS(file.path(results_dir, "data", "08", "sce_orig.rds"))
keep <- readRDS(file.path(results_dir, "data", "08", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "08", "keep2.rds"))
sce_orig <- sce_orig[ , keep]
keep <- readRDS(file.path(results_dir, "data", "23", "keep1.rds"))
sce_orig <- sce_orig[ , keep]
sce_orig <- calculateQCMetrics(sce_orig)
detected_genes <- rowSums(counts(sce_orig) >= 1)
sce_orig <- sce_orig[which(detected_genes >= 1), ]

seurat_orig <- as.Seurat(sce_orig, counts = "counts", data = NULL)
umap <- as.matrix(data.frame(sce$EX_ppk_umap1, sce$EX_ppk_umap2))
colnames(umap) <- paste0("UMAP_", 1:2)
rownames(umap) <- colnames(sce_orig)
seurat_orig[["UMAP"]] <- CreateDimReducObject(
  embeddings = umap, key = "UMAP_", assay = DefaultAssay(seurat_orig))
Idents(object = seurat_orig) <- Idents(object = seurat)
seurat_orig <- NormalizeData(seurat_orig)
seurat_orig <- ScaleData(seurat_orig, features = rownames(seurat_orig))

genes <- c("TBR1", "SATB2", "SATB1", "WFS1", "GAD1", "CALB1", "SST", "PVALB", "CALB2")
rm(umap)
DoHeatmap(seurat_orig, slot = "data", features = genes, size = 3) +
  NoLegend() +
  scale_fill_gradient2(
    low = rev(c('#D1E5F0','#67A9CF','#2166AC')), mid = "white",
    high = rev(c('#B2182B','#EF8A62','#FDDBC7')), midpoint = 0,
    guide = "colourbar", aesthetics = "fill")
VlnPlot(seurat_orig, features = genes, ncol = 3)
FeaturePlot(
  seurat_orig, reduction = "UMAP", features = genes, order = TRUE,
  cols = c("lightgrey", "red"), max.cutoff = 20, ncol = 3)
DotPlot(seurat_orig, features = genes, cols = c("blue", "red")) + RotatedAxis()
```

# References

This is the concluding section of the document. Here we write relevant results to disk, output the `sessionInfo`, and create a bibliography for works cited.

```{r}
saveRDS(sce, file.path(results_dir, "data", "23", "sce.rds"))

sessionInfo()
```
